### 2013-07-26 paper.tex
Initial.

%% This is file `elsarticle-template-1-num.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%%
%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
%%

%\documentclass[preprint,authoryear,review,12pt]{elsarticle}
\documentclass[final,5p,times,twocolumn]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,two column; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{color}
\usepackage{multirow,booktabs,ctable,array}
\usepackage{lscape}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{ulem}
\usepackage{setspace}
\usepackage{listings}
\usepackage{float}

\floatstyle{plain}
\newfloat{command}{thp}{lop}
\floatname{command}{Command}

%\usepackage[nomarkers,notablist]{endfloat}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
% \usepackage{amsthm}

 \usepackage{makecell}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\providecommand{\OO}[1]{\operatorname{O}\bigl(#1\bigr)}

\graphicspath{{./Figures/}
                          }

\long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

    \usepackage{color}

    \definecolor{listcomment}{rgb}{0.0,0.5,0.0}
    \definecolor{listkeyword}{rgb}{0.0,0.0,0.5}
    \definecolor{listnumbers}{gray}{0.65}
    \definecolor{listlightgray}{gray}{0.955}
    \definecolor{listwhite}{gray}{1.0}

\newcommand{\lstsetcpp}
{
\lstset{frame = tb,
        framerule = 0.25pt,
        float,
        fontadjust,
        backgroundcolor={\color{listlightgray}},
        basicstyle = {\ttfamily\scriptsize},
        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
        identifierstyle = {\ttfamily},
        commentstyle = {\ttfamily\color{listcomment}\textit},
        stringstyle = {\ttfamily},
        showstringspaces = false,
        showtabs = false,
        numbers = none,
        numbersep = 6pt,
        numberstyle={\ttfamily\color{listnumbers}},
        tabsize = 2,
        language=[ANSI]C++,
        floatplacement=!h,
        caption={},
        captionpos=b,
        }
}

\journal{Medical Image Analysis}

\begin{document}

\begin{frontmatter}

\title{Explicit B-spline Regularization in Diffeomorphic Image Registration}

\author[label1]{Nicholas J.~Tustison\fnref{label0}}
%  \ead{ntustison@virginia.edu}
  \fntext[label0]{\scriptsize Corresponding author:  PO Box 801339, Charlottesville, VA 22908; T:  434-924-7730; email address:  ntustison@virginia.edu }
\author[label2]{Brian B.~Avants}

\address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
\address[label2]{Penn Image Computing and Science Laboratory, Department of Radiology, University of Pennsylvania, Philadelphia, PA}

%\maketitle

%\linenumbers

\begin{abstract}
Important contributions in the evolution of image registration algorithms include those 
in which the correspondence relationship is characterized by diffeomorphic constraints.
The  popularity of these diffeomorphic image registration methods is due largely to their 
topological properties and success in providing solutions to small and large deformation 
estimation problems. Variants of diffeomorphic algorithms include those characterized by 
time-varying velocity fields, constant velocity fields, and symmetrical considerations. 
To enforce transform plausibility, regularization of the velocity field is usually enforced 
either implicitly through penalization of the Laplacian or through explicit Gaussian smoothing.
In this work, we demonstrate that B-splines provide a suitable alternative for regularization 
and provide a new “flavor” of diffeomorphic image registration solutions.  This is showcased
by an open source, well-vetted implementation of three popular diffeomorphic variants available
through the Insight Toolkit ITK and the Advanced Normalization Tools (ANTs) repository.  
Evaluation is also performed using publicly available brain data used in the MICCAI 
2012 brain atlas labeling challenge.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

\end{frontmatter}
%
%
\newpage

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text

\section{Introduction}
%State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results.

Significant algorithmic developments characterizing modern intensity-based
image registration research include the B-spline parameterized approach
(so called {\em free-form deformation}) with early contributions 
including \cite{szeliski1997,thevenaz1998,rueckert1999}.  Amongst the 
variant extensions, the 
{\em directly manipulated free-form deformation} approach  \cite{tustison2009} 
 addressed the hemstitching issue associated with steepest descent traversal of
problematic energy topographies during the course of optimization.
%caused by the distribution of the uniform 
%B-spline shape functions over the transformation domain

Other important image registration research
reflected increased emphasis on topological transformation considerations
in modeling biological/physical systems where topology is 
consistent throughout the course of deformation or a 
homeomorphic relationship is assumed between image domains.
Methods such as LDDMM \cite{beg2005} optimize time-varying velocity field 
flows to yield diffeomorphic transformations.  Alternatively, the FFD 
variant reported in \cite{rueckert2006} enforced diffeomorphic transforms
by concatenating multiple FFD transforms, each of which is constrained
to describe a one-to-one mapping.  Another FFD registration
incorporated the recent log-Euclidean framework for enforcing diffeomorphic
transformations \cite{Modat2011}.
Recently, the work of 
\cite{de-craene2011} combined these registration concepts into a single
framework called {\em temporal free-form deformation} in which the 
time-varying velocity field characteristic of LDDMM-style algorithms
is modeled using a 4-D B-spline object (3-D + time).  Integration of 
the velocity field yields the mapping between parameterized time points.

In this work, we describe our extension to these methods.  Similar to 
\cite{de-craene2011}, we also use an $N$-D + time B-spline object to 
represent the characteristic velocity field.  However, we use the 
directly manipulated free-form deformation optimization formulation to improve 
convergence during the course of optimization.  This also facilitates
modeling temporal periodicity and
the enforcement of stationary boundaries consistent with diffeomorphic
transforms.
We also incorporate B-spline mesh multi-resolution capabilities
for increased control during registration progression.  
Most importantly, we also describe
the parallelized algorithmic implementation as open source available through the Insight Toolkit.%
\footnote{
http://www.itk.org/
}

We first describe the methodology by laying out a mathematical description 
of the various algorithmic elements coupled with implementation details
where appropriate.  This is followed by an evaluation on publicly available
brain data.

\section{Material and Methods}
%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.

{\color{red}{\em Remember Brian's suggestion of adding the work where we integrated the velocity field (via the control point lattice) as opposed to integrating the path through the velocity field.}}

Three popular algorithms [1,2,3] (and their B-spline analogs) were implemented as part of the recent refactoring of the Insight Toolkit (ITK) [5] and included as part of the Advanced Normalization Tools (ANTs) repository [6].  Thus, all code used for this work is available as open source.   For comparison, the ITK-based, well-vetted SyN algorithm [3], which uses explicit Gaussian regularization, was used to provide the canonical mappings in the MICCAI 2012 multi-atlas labeling challenge [7] between 20 training subjects and 15 testing subjects (140 labeled structures).  Using the exact same parameters, except those specific to the regularization, we produced the analogous B-spline SyN mappings for the same data and compared transformations using the Dice overlap.

\section{Results}
%Results should be clear and concise.

Sample label results (with ground truth) are showin in Fig. 1.  Box plots comparing the Dice metrics between the two diffeomorphic registration algorithms are pro-vided in Fig. 1.  A paired Student’s t-test demonstrated that B-spline SyN produced slightly better average overlap ($p < 1e-4$).

\section{Discussion}
%This should explore the significance of the results of the work, not repeat them. A combined Results and Discussion section is often appropriate. Avoid extensive citations and discussion of published literature.

B-spline regularization is easily adapted into the diffeomorphic registration framework and performs comparably to analogous algorithms which we demonstrated in the case of SyN.  Even more importantly, all source code and data used in this work is publicly available for the interested researcher to explore.

\section{Conclusions}
%The main conclusions of the study may be presented in a short Conclusions section, which may stand alone or form a subsection of a Discussion or Results and Discussion section.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \citep{key}          ==>>  [#]
%%   \cite[chap. 2]{key} ==>>  [#, chap. 2]
%%   \citet{key}         ==>>  Author [#]

%% References with bibTeX database:

\section*{Acknowledgments}

\section*{References}

\bibliographystyle{elsarticle-harv}
\bibliography{references}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model1-num-names.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file `elsarticle-template-1-num.tex'.

### 2013-08-19 paper.tex
ENH:  Adding more edits.

@@
-%% This is file `elsarticle-template-1-num.tex',
-%%
-%% Copyright 2009 Elsevier Ltd
-%%
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%%
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%%
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%%
-%% $Id: elsarticle-template-1-num.tex 149 2009-10-08 05:01:15Z rishi $
-%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-1-num.tex $
-%%
-
 %\documentclass[preprint,authoryear,review,12pt]{elsarticle}
-\documentclass[final,5p,times,twocolumn]{elsarticle}
+\documentclass{frontiersSCNS}
 
@@
 
-\journal{Medical Image Analysis}
+\copyrightyear{}
+\pubyear{}
+
+\def\journal{Neuroscience}%%% write here for which journal %%%
+\def\DOI{}
+\def\articleType{Methods}
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+\def\firstAuthorLast{Tustison {et~al.}} %use et al only if is more than 1 author
+\def\Authors{Nicholas J. Tustison\,$^{1,*}$ and Brian A. Avants\,$^{2}$
+ }
+% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
+% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
+\def\Address{$^{1}$University of Virginia, Department of Radiology and Medical Imaging, Charlottesville, VA, USA \\
+$^{2}$Penn Image Computing and Science Laboratory, University of Pennsylvania, Department of Radiology, Philadelphia, PA, USA }
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
+\def\corrAuthor{Nick Tustison}
+\def\corrAddress{University of Virginia, Department of Radiology and Medical Imaging,  480 Ray C Hunt Drive, Charlottesville, VA, 22903}
+\def\corrEmail{ntustison@virginia.edu}
+
+% \color{FrontiersColor} Is the color used in the Journal name, in the title, and the names of the sections
 
 \begin{document}
+\onecolumn
+\firstpage{1}
 
-\begin{frontmatter}
+\title[Explicit B-spline Regularization]{Explicit B-spline Regularization in Diffeomorphic Image Registration}
+\author[\firstAuthorLast ]{\Authors}
+\address{}
+\correspondance{}
+\editor{}
+\topic{}
 
-\title{Explicit B-spline Regularization in Diffeomorphic Image Registration}
-
-\author[label1]{Nicholas J.~Tustison\fnref{label0}}
-%  \ead{ntustison@virginia.edu}
-  \fntext[label0]{\scriptsize Corresponding author:  PO Box 801339, Charlottesville, VA 22908; T:  434-924-7730; email address:  ntustison@virginia.edu }
-\author[label2]{Brian B.~Avants}
-
-\address[label1]{Department of Radiology and Medical Imaging, University of Virginia, Charlottesville, VA}
-\address[label2]{Penn Image Computing and Science Laboratory, Department of Radiology, University of Pennsylvania, Philadelphia, PA}
-
-%\maketitle
+\maketitle
 
@@
 \begin{abstract}
-Important contributions in the evolution of image registration algorithms include those 
-in which the correspondence relationship is characterized by diffeomorphic constraints.
-The  popularity of these diffeomorphic image registration methods is due largely to their 
-topological properties and success in providing solutions to small and large deformation 
+Important methodological developments in the evolution of image registration algorithms include those 
+in which the correspondence relationship is characterized by diffeomorphisms.
+The  popularity of these approaches is due largely to their 
+topological properties and success in providing biologically plausible 
+solutions to small and large deformation 
 estimation problems. Variants of diffeomorphic algorithms include those characterized by 
-time-varying velocity fields, constant velocity fields, and symmetrical considerations. 
-To enforce transform plausibility, regularization of the velocity field is usually enforced 
-either implicitly through penalization of the Laplacian or through explicit Gaussian smoothing.
-In this work, we demonstrate that B-splines provide a suitable alternative for regularization 
-and provide a new “flavor” of diffeomorphic image registration solutions.  This is showcased
-by an open source, well-vetted implementation of three popular diffeomorphic variants available
-through the Insight Toolkit ITK and the Advanced Normalization Tools (ANTs) repository.  
-Evaluation is also performed using publicly available brain data used in the MICCAI 
-2012 brain atlas labeling challenge.
+time-varying and constant velocity fields, and symmetrical considerations. 
+Prior information in the form of regularization is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
+Gaussian smoothing (a la Thirion's Demons \citep{thirion1998}).
+In the context of the original Demons' framework, the traditional free-form deformation method can be viewed as a variant in which explicit regularization is achieved through the B-spline basis functions.
+This characterization can be used to provide alternative ``flavors'' of diffeomorphic image registration solutions with several advantages which we describe in this work.
+Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
+A thorough evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open data and open source evaluation tools.
+\tiny
+\keyFont{ \section{Keywords:} 
+open science best practices comparison reproducibility software }
 \end{abstract}
 
-\begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-\end{keyword}
-
-\end{frontmatter}
+%\begin{itemize}
+%  \item We should trace the use of explicit regularization from 
+%  \item The advantage is not the parameterization per se (contra Tom) but the other 
+%  benefits listed (Vercuraten, Non-parametric Diffeomorphic Image Registration with the Demons Algorithm).
+%  \item This paper provides the link between traditional B-spline approaches and other
+%  approaches, i.e. Gaussian smoothing, Demons, 
+%  \item The ability to weight the boundaries provides a natural way of enforcing Dirichlet boundary conditions which is consistent with the geometric (i.e. B-spline) modeling. (Cahill, SPIE 2012).
+%  \item Fluid vs. Elastic B-spline registration algorithms
+%  \item Fitting a continuous object (as opposed to the discrete gaussian which has no
+%        continuity nor does convolution using FFT provide continuity constraints).
+%  \item Fitting routine is parallelized for fast sampling as is sampling to go from continuous object to sampled b-spline object.
+%  \item Everything is open-source in ANTs and ITK.
+%  \item Also, we should talk about how B-spline SyN works really well for large smoothing  but Gaussian SyN does not.
+%\end{itemize}
 %
 %
-\newpage
+%Also talk about fixed boundary conditions in the context of Nathan Cahill, SPIE 2012.
+%
 
@@
 
-Significant algorithmic developments characterizing modern intensity-based
-image registration research include the B-spline parameterized approach
-(so called {\em free-form deformation}) with early contributions 
-including \cite{szeliski1997,thevenaz1998,rueckert1999}.  Amongst the 
-variant extensions, the 
-{\em directly manipulated free-form deformation} approach  \cite{tustison2009} 
- addressed the hemstitching issue associated with steepest descent traversal of
-problematic energy topographies during the course of optimization.
-%caused by the distribution of the uniform 
-%B-spline shape functions over the transformation domain
+Establishment of anatomical and functional correspondence
+is a crucial step towards insight into certain biological 
+sciences.  Neuroscience research efforts, such as characterizing 
+brain morphology, require accurate and robust methods for
+producing such mappings.  The 
+extensive literature detailing methodology is evidence of the rich history of 
+algorithmic development which continues contemporaneously.
+We highlight several key contributions which are particularly relevant to the work presented.
 
-Other important image registration research
-reflected increased emphasis on topological transformation considerations
-in modeling biological/physical systems where topology is 
-consistent throughout the course of deformation or a 
-homeomorphic relationship is assumed between image domains.
-Methods such as LDDMM \cite{beg2005} optimize time-varying velocity field 
-flows to yield diffeomorphic transformations.  Alternatively, the FFD 
+Free-form deformation (FFD) image registration, preceded by related work for
+geometric modeling \citep{sederberg1986}, originated with such important
+contributions as \cite{szeliski1997,thevenaz1998,rueckert1999}.  Continued 
+development within this early spline-based paradigm produced additional innovations
+such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation
+constraints \citep[e.g.][]{rohlfing2003}, and open source contributions \citep[e.g.][]{ibanez2005,klein2010,shackleford2010}.
+
+Parallel to this branch of algorithmic progress are the informally 
+denoted ``dense transforms''
+perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
+Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are elucidated in 
+the works of \cite{bro-nielsen1996,pennec1999} who observe that
+Gaussian smoothing, characteristic of Demons, of the update or displacement
+field is a greedy approximation for solving the partial differential equations governing
+the physics of an elastic or fluid deformation.  However, the use of such 
+approximations comes with the caveat that physical properties, such as topological
+regularity,
+are no longer guaranteed.
+
+It is interesting to note that within this context, traditional FFD algorithms
+can be viewed as a type of fluid-like Demons approach 
+where, rather than produced by Gaussian convolution, smooth fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.
+This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD
+optimization.  Further details can be gleaned from the ITK implementation of this work%
+\footnote{
+http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineSmoothingOnUpdateDisplacementFieldTransform.html
+}
+which permits both B-spline smoothing on the update (``viscous'') and total (``elastic'') 
+displacement fields at each iteration (cf
+analogous Gaussian, i.e. Demons, implementation%
+\footnote{
+http://www.itk.org/Doxygen/html/classitk\_1\_1GaussianSmoothingOnUpdateDisplacementFieldTransform.html
+}).
+
+Continuing from the work of \cite{christensen1996} and subsequent exploration into the mathematical formalisms of diffeomorphisms \citep[e.g.][]{dupuis1998}, 
+the well-known Large Deformation Diffeomorphic Metric Mapping (LDDMM) algorithm  
+was proposed in \cite{Beg2005}.  In contrast to the mapping produced
+by \cite{christensen1996}, LDDMM yields the geodesic solution in the space of diffeomorphisms between two images.
+
+%Other important image registration research
+%reflected increased emphasis on topological transformation considerations
+%in modeling biological/physical systems where topology is 
+%consistent throughout the course of deformation or a 
+%homeomorphic relationship is assumed between image domains.
+%Methods such as LDDMM \citep{beg2005} optimize time-varying velocity field 
+%flows to yield diffeomorphic transformations.
+
+Alternatively, the FFD 
 variant reported in \cite{rueckert2006} enforced diffeomorphic transforms
@@
 
-In this work, we describe our extension to these methods.  Similar to 
-\cite{de-craene2011}, we also use an $N$-D + time B-spline object to 
-represent the characteristic velocity field.  However, we use the 
-directly manipulated free-form deformation optimization formulation to improve 
-convergence during the course of optimization.  This also facilitates
-modeling temporal periodicity and
-the enforcement of stationary boundaries consistent with diffeomorphic
-transforms.
-We also incorporate B-spline mesh multi-resolution capabilities
-for increased control during registration progression.  
-Most importantly, we also describe
-the parallelized algorithmic implementation as open source available through the Insight Toolkit.%
-\footnote{
-http://www.itk.org/
-}
+%
+%
+%framework developed independently of Demons and LDDMM/DARTEL.
+%But then the FFD people saw how useful the diff. properties were and
+%tried to integrate B-splines into the diff. framework.
+%
+%
+%The early deformable work of 
+%\begin{itemize}
+%  \item xx Early deformable work \cite{broit1981,bajcsy1989}.
+%  \item FFD \cite{Sleziski,Rueckert1989}
+%  \item Christensen's fluid work
+%  \item Optical flow?
+%  \item Fast fluid registration of medical images (Bro-Nielsen and Gramkow).
+%  \item Consistent image registration
+%  \item Demons and Diff demons
+%  \item LDDMM 
+%  \item DARTEL
+%  \item Arno Klein's study
+%\end{itemize}
+%
+%
+%
+%Early work included the elastic deformation
 
-We first describe the methodology by laying out a mathematical description 
-of the various algorithmic elements coupled with implementation details
-where appropriate.  This is followed by an evaluation on publicly available
-brain data.
+%Significant algorithmic developments characterizing modern intensity-based
+%image registration research include the B-spline parameterized approach
+%(so called {\em free-form deformation}) with early contributions 
+%including \cite{szeliski1997,thevenaz1998,rueckert1999}.  Amongst the 
+%variant extensions, the 
+%{\em directly manipulated free-form deformation} approach  \cite{tustison2009} 
+% addressed the hemstitching issue associated with steepest descent traversal of
+%problematic energy topographies during the course of optimization.
+%%caused by the distribution of the uniform 
+%%B-spline shape functions over the transformation domain 
+%
+%
+%In this work, we describe our extension to these methods.  Similar to 
+%\cite{de-craene2011}, we also use an $N$-D + time B-spline object to 
+%represent the characteristic velocity field.  However, we use the 
+%directly manipulated free-form deformation optimization formulation to improve 
+%convergence during the course of optimization.  This also facilitates
+%modeling temporal periodicity and
+%the enforcement of stationary boundaries consistent with diffeomorphic
+%transforms.
+%We also incorporate B-spline mesh multi-resolution capabilities
+%for increased control during registration progression.  
+%Most importantly, we also describe
+%the parallelized algorithmic implementation as open source available through the Insight Toolkit.%
+%\footnote{
+%http://www.itk.org/
+%}
+%
+%We first describe the methodology by laying out a mathematical description 
+%of the various algorithmic elements coupled with implementation details
+%where appropriate.  This is followed by an evaluation on publicly available
+%brain data.
 
@@
 %Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.
-
-{\color{red}{\em Remember Brian's suggestion of adding the work where we integrated the velocity field (via the control point lattice) as opposed to integrating the path through the velocity field.}}
 
@@
 
+\subsection{Public Data}
+
+\subsection{Implementation}
+
 \section{Results}
@@
 
-Sample label results (with ground truth) are showin in Fig.
-1.
-Box plots comparing the Dice metrics between the two diffeomorphic registration algorithms are pro-vided in Fig.
-1.
-A paired Student’s t-test demonstrated that B-spline SyN produced slightly better average overlap ($p < 1e-4$).
-
-\section{Discussion}
+\section{Discussion and Conclusions}
 %This should explore the significance of the results of the work, not repeat them.
@@
 
-\section{Conclusions}
+%\section{Conclusions}
 %The main conclusions of the study may be presented in a short Conclusions section, which may stand alone or form a subsection of a Discussion or Results and Discussion section.
@@
 
-\section*{Acknowledgments}
-
-\section*{References}
-
-\bibliographystyle{elsarticle-harv}
+\bibliographystyle{frontiersinSCNS}
+%\bibliographystyle{plain}
 \bibliography{references}

### 2013-09-07 paper.tex
ENH:  Adding files.

@@
 perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
-Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are elucidated in 
-the works of \cite{bro-nielsen1996,pennec1999} who observe that
+Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are detailed in 
+the works of \cite{bro-nielsen1996} and \cite{pennec1999} who observe that
 Gaussian smoothing, characteristic of Demons, of the update or displacement
@@
 approximations comes with the caveat that physical properties, such as topological
-regularity,
-are no longer guaranteed.
+regularity, are no longer guaranteed.
 
@@
 where, rather than produced by Gaussian convolution, smooth fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.
-This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD
-optimization.  Further details can be gleaned from the ITK implementation of this work%
+This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
+This preconditioning used to reduce the
+
+Further details can be gleaned from the ITK implementation of this work%
 \footnote{

### 2013-09-07 Results/compileLogFiles.pl

@@ -1,0 +1,137 @@
+#! /usr/bin/perl -w
+
+use strict;
+
+use Cwd 'realpath';
+use File::Find;
+use File::Basename;
+use File::Path;
+use File::Spec;
+use FindBin qw($Bin);
+
+my $usage = qq{
+  Usage: compileLogFiles.pl <inputDir> <outputFile>
+ };
+
+my ( $inputDirectory, $outputFile ) = @ARGV;
+
+my @logFiles = <${inputDirectory}/*syn.o*>;
+
+my @numberOfIterationsPerLevel = ( 100, 100, 70, 20 );
+
+open( CSV_FILE, ">${outputFile}" );
+print CSV_FILE "WHICH,REGULARIZATION,";
+for( my $i = 0; $i < @numberOfIterationsPerLevel; $i++ )
+  {
+  for( my $j = 0; $j < $numberOfIterationsPerLevel[$i]; $j++ )
+    {
+    print CSV_FILE "LEVEL${i}_ITERATION${j}";
+    if( $i == @numberOfIterationsPerLevel-1 && $j == $numberOfIterationsPerLevel[$i]-1 )
+      {
+      print CSV_FILE "\n";
+      }
+    else
+      {
+      print CSV_FILE ",";
+      }
+    }
+  }
+
+for( my $i = 0; $i < @logFiles; $i++ )
+  {
+  open( FILE, "<${logFiles[$i]}" );
+  my @fileContents = <FILE>;
+  close( FILE );
+
+  # Check to see if this particular registration ran to
+  # completion successfully.
+
+  my @lastLineTokens = split( ': ', $fileContents[-1] );
+  if( ${lastLineTokens[0]} !~ m/Output\ warped\ image/ )
+    {
+    print "Bad file:  ${logFiles[$i]}\n";
+    }
+
+  chomp( ${lastLineTokens[1]} );
+  my ( $filename, $directories, $suffix ) = fileparse( ${lastLineTokens[1]}, ".nii.gz" );
+
+  my $regularization = 'syn';
+  if( $directories =~ /BSyN_DOUBLE/ )
+    {
+    $regularization = 'bsyn';
+    }
+  print CSV_FILE "${filename},${regularization},";
+
+  # get the iteration index where each level begins
+
+  my $keepTrack = 0;
+  my @iterationIndices = ();
+  for( my $j = 0; $j < @fileContents; $j++ )
+    {
+    if( $fileContents[$j] =~ /^Stage\ 2/ )
+      {
+      $keepTrack = 1;
[…]

### 2013-09-07 Results/overlapMeasures.R

@@ -1,0 +1,88 @@
+library( ggplot2 )
+
+stats <- list()
+volumes <- list()
+
+cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MICCAI' )
+
+stats[[1]] <- read.csv( "OverlapMeasures/overlapMeasuresCUMC.csv" )
+volumes[[1]] <- read.csv( "OverlapMeasures/labelVolumesCUMC.csv" )
+stats[[2]] <- read.csv( "OverlapMeasures/overlapMeasuresIBSR.csv" )
+volumes[[2]] <- read.csv( "OverlapMeasures/labelVolumesIBSR.csv" )
+stats[[3]] <- read.csv( "OverlapMeasures/overlapMeasuresLPBA.csv" )
+volumes[[3]] <- read.csv( "OverlapMeasures/labelVolumesLPBA.csv" )
+stats[[4]] <- read.csv( "OverlapMeasures/overlapMeasuresMGH.csv" )
+volumes[[4]] <- read.csv( "OverlapMeasures/labelVolumesMGH.csv" )
+stats[[5]] <- read.csv( "OverlapMeasures/overlapMeasuresMICCAI.csv" )
+volumes[[5]] <- read.csv( "OverlapMeasures/labelVolumesMICCAI.csv" )
+
+
+#
+# Render box plots (syn vs. bsyn) for each of the 5 cohorts
+#
+
+for( i in 1:length( cohortNames ) )
+  {
+  diceSyN <- stats[[i]]$AllLabels[stats[[i]]$Regularization == 'syn']
+  diceBSplineSyN <- stats[[i]]$AllLabels[stats[[i]]$Regularization == 'bsyn']
+
+  plotLabels <- c( rep.int( "SyN", length( diceSyN ) ), rep.int( "BSplineSyN", length( diceBSplineSyN ) ) )
+
+  plotData <- data.frame( Type = plotLabels, Dice = c( diceSyN, diceBSplineSyN ) )
+
+  myPlot <- ggplot( plotData,  aes( x = factor( Type ), y = Dice ) ) +
+            geom_boxplot( aes( fill = Type ) ) +
+            ggtitle( cohortNames[i] ) +
+            scale_x_discrete( "" )
+
+  ggsave( filename = paste0( "synComparison", cohortNames[i], ".pdf" ), plot = myPlot, width = 8, height = 5, units = 'in' )
+  }
+
+
+#
[…]

### 2013-09-07 Results/plotConvergence.R

@@ -1,0 +1,8 @@
+convergenceTable <- read.csv( "LogFilesMGH.csv" )
+
+numberOfIterationsPerLevel <- c( 100, 100, 70, 20 )
+
+for( i in 1:length( numberOfIterationsPerLevel ) )
+  {
+  averageConvergence <- colMeans( convergenceTable[3:( 2+numberOfIterationsPerLevel[i] )] )
+  }

### 2013-09-09 Results/overlapMeasures.R
ENH:  Plotting convergence.

@@ -31,7 +31,8 @@ for( i in 1:length( cohortNames ) )
   plotData <- data.frame( Type = plotLabels, Dice = c( diceSyN, diceBSplineSyN ) )
 
   myPlot <- ggplot( plotData,  aes( x = factor( Type ), y = Dice ) ) +
-            geom_boxplot( aes( fill = Type ) ) +
+            geom_boxplot( aes( fill = Type ), alpha = 0.5 ) +
+  										scale_fill_manual( values = c( "darkred", "navyblue" ), labels = c( "BSplineSyN", "SyN" ) ) +
             ggtitle( cohortNames[i] ) +
             scale_x_discrete( "" )
 

### 2013-09-09 Results/plotConvergence.R

@@ -1,8 +1,114 @@
-convergenceTable <- read.csv( "LogFilesMGH.csv" )
+library( ggplot2 )
+library( gridExtra )
+
+logFiles <- c( "CUMC", "IBSR", "LPBA", "MGH", "MICCAI" )
+cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MICCAI' )
+
+stats <- list()
+volumes <- list()
+
+stats[[1]] <- read.csv( "OverlapMeasures/overlapMeasuresCUMC.csv" )
+volumes[[1]] <- read.csv( "OverlapMeasures/labelVolumesCUMC.csv" )
+stats[[2]] <- read.csv( "OverlapMeasures/overlapMeasuresIBSR.csv" )
+volumes[[2]] <- read.csv( "OverlapMeasures/labelVolumesIBSR.csv" )
+stats[[3]] <- read.csv( "OverlapMeasures/overlapMeasuresLPBA.csv" )
+volumes[[3]] <- read.csv( "OverlapMeasures/labelVolumesLPBA.csv" )
+stats[[4]] <- read.csv( "OverlapMeasures/overlapMeasuresMGH.csv" )
+volumes[[4]] <- read.csv( "OverlapMeasures/labelVolumesMGH.csv" )
+stats[[5]] <- read.csv( "OverlapMeasures/overlapMeasuresMICCAI.csv" )
+volumes[[5]] <- read.csv( "OverlapMeasures/labelVolumesMICCAI.csv" )
+
 
 numberOfIterationsPerLevel <- c( 100, 100, 70, 20 )
+iterationCumSum <- cumsum( numberOfIterationsPerLevel )
 
-for( i in 1:length( numberOfIterationsPerLevel ) )
+for( f in 1:length( logFiles ) )
   {
-  averageConvergence <- colMeans( convergenceTable[3:( 2+numberOfIterationsPerLevel[i] )] )
+  cat( "Processing ", logFiles[f], "\n", sep = "" )
+  convergenceTable <- read.csv( paste0( "LogFiles", logFiles[f], ".csv" ) )
+
+  convergencePlots <- list();
+		for( i in 1:length( numberOfIterationsPerLevel ) )
+				{
+				colIndices <- 4:( 3+numberOfIterationsPerLevel[i] )
+				if( i > 1 )
+						{
+						colIndices <- ( 4 + iterationCumSum[i-1] ):( 3 + iterationCumSum[i] )
+						}
+
[…]

### 2013-10-03 paper.tex
ENH:  Adding Log jacobians.

--- 2013-08-19 paper.tex
@@
 solutions to small and large deformation 
-estimation problems. Variants of diffeomorphic algorithms include those characterized by 
+estimation problems. Popular variants of diffeomorphic algorithms include those characterized by 
 time-varying and constant velocity fields, and symmetrical considerations. 
-Prior information in the form of regularization is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
-Gaussian smoothing (a la Thirion's Demons \citep{thirion1998}).
+Prior information (i.e.
+regularization) is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
+Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
 In the context of the original Demons' framework, the traditional free-form deformation method can be viewed as a variant in which explicit regularization is achieved through the B-spline basis functions.
-This characterization can be used to provide alternative ``flavors'' of diffeomorphic image registration solutions with several advantages which we describe in this work.
+This characterization can be used to provide alternative B-spline ``flavored'' diffeomorphic image registration solutions with several advantages which we describe in this work.
 Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
-A thorough evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open data and open source evaluation tools.
+A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open data and open source evaluation tools.
 \tiny
 \keyFont{ \section{Keywords:} 
-open science best practices comparison reproducibility software }
+open science, best practices, reproducibility, software }
 \end{abstract}
@@
 Establishment of anatomical and functional correspondence
-is a crucial step towards insight into certain biological 
+is a crucial step towards gaining insight into certain biological 
 sciences.  Neuroscience research efforts, such as characterizing 
@@
 
-Free-form deformation (FFD) image registration, preceded by related work for
-geometric modeling \citep{sederberg1986}, originated with such important
-contributions as \cite{szeliski1997,thevenaz1998,rueckert1999}.  Continued 
-development within this early spline-based paradigm produced additional innovations
-such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation
-constraints \citep[e.g.][]{rohlfing2003}, and open source contributions \citep[e.g.][]{ibanez2005,klein2010,shackleford2010}.
+Free-form deformation (FFD) image registration, characterized by  regularization based on the B-spline basis functions, has several 
+advantages including ease of algorithmic understanding, 
+good performance, desirable geometric properties, and several available 
+implementations.  Current research was preceded by related work for
+geometric modeling \citep{sederberg1986} and originated with such important
+contributions as \cite{szeliski1997,thevenaz1998,rueckert1999}.  
+Continued development within this early spline-based paradigm produced additional innovations such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation constraints \citep[e.g.][]{rohlfing2003}, and notable open source implementations \citep[e.g.][]{ibanez2005,klein2010,shackleford2010}.
 
@@
 perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
-Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are elucidated in 
-the works of \cite{bro-nielsen1996,pennec1999} who observe that
+Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are detailed in 
+the works of \cite{bro-nielsen1996} and \cite{pennec1999} who observe that
 Gaussian smoothing, characteristic of Demons, of the update or displacement
 field is a greedy approximation for solving the partial differential equations governing
-the physics of an elastic or fluid deformation.  However, the use of such 
-approximations comes with the caveat that physical properties, such as topological
-regularity,
-are no longer guaranteed.
+the physics of an elastic or fluid deformation, respectively.  However, the use of such 
+approximations entails that physical properties, such as topological
+regularity, are no longer guaranteed.
 
@@
 can be viewed as a type of fluid-like Demons approach 
-where, rather than produced by Gaussian convolution, smooth fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.
-This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD
-optimization.  Further details can be gleaned from the ITK implementation of this work%
+where, rather than produced by Gaussian convolution, regularized fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.
+This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
+As mentioned in \cite{tustison2009}, this preconditioning is used to reduce the hemstitching induced by the traditional gradient-based FFD formulation.
+Further details can be gleaned from the ITK implementation of this 
+earlier work%
 \footnote{
@@
 was proposed in \cite{Beg2005}.  In contrast to the mapping produced
-by \cite{christensen1996}, LDDMM yields the geodesic solution in the space of diffeomorphisms between two images.
+by \cite{christensen1996}, LDDMM yields the geodesic solution in the space of diffeomorphisms between two images. Since its introduction, LDDMM has
+inspired much innovation in the image registration literature.  
+Symmetrical considerations in the velocity field parameterization, for example, are discussed in \cite{avants2008} in the context of a cross correlation similarity metric which minimizes the bias of the resulting transformation in selecting the identity of the fixed and moving images.
+A greedy version of this algorithm which has proven successful in general application \citep{Klein2009}.
 

### 2013-10-03 Results/compileLogFiles.pl

@@ -20,7 +20,7 @@ my @logFiles = <${inputDirectory}/*syn.o*>;
 my @numberOfIterationsPerLevel = ( 100, 100, 70, 20 );
 
 open( CSV_FILE, ">${outputFile}" );
-print CSV_FILE "WHICH,REGULARIZATION,";
+print CSV_FILE "WHICH,REGULARIZATION,ELAPSED_TIME,";
 for( my $i = 0; $i < @numberOfIterationsPerLevel; $i++ )
   {
   for( my $j = 0; $j < $numberOfIterationsPerLevel[$i]; $j++ )
@@ -36,20 +36,36 @@ for( my $i = 0; $i < @numberOfIterationsPerLevel; $i++ )
       }
     }
   }
+  
+my @whichRegistrations = ();  
 
-for( my $i = 0; $i < @logFiles; $i++ )
+# go backwards to get the latest duplicate (if exists)
+
+for( my $i = @logFiles - 1; $i >= 0; $i-- )
   {
   open( FILE, "<${logFiles[$i]}" );
   my @fileContents = <FILE>;
   close( FILE );
 
+  if( @fileContents == 0 )
+    {
+    print "Bad file:  ${logFiles[$i]}\n";
+    next;
+    }
+
   # Check to see if this particular registration ran to
   # completion successfully.
 
   my @lastLineTokens = split( ': ', $fileContents[-1] );
+  if( @lastLineTokens < 2 )
+    {
+    print "Bad file:  ${logFiles[$i]}\n";
+    next;
+    }
   if( ${lastLineTokens[0]} !~ m/Output\ warped\ image/ )
     {
     print "Bad file:  ${logFiles[$i]}\n";
+    next;
     }
 
   chomp( ${lastLineTokens[1]} );
@@ -60,7 +76,20 @@ for( my $i = 0; $i < @logFiles; $i++ )
     {
     $regularization = 'bsyn';
     }
-  print CSV_FILE "${filename},${regularization},";
+
+  # ensure that there are no duplicates 
+  
+  my $which = "${filename}_${regularization}";
+  
+  if( grep { $_ eq $which } @whichRegistrations )
+    {
+    print "Duplicate entry: $which\n";
+    next;
+    }
+  else
+    {  
+    push( @whichRegistrations, $which );
+    }
 
   # get the iteration index where each level begins
 
@@ -88,7 +117,22 @@ for( my $i = 0; $i < @logFiles; $i++ )
   else
     {
[…]

### 2013-10-12 paper.tex
ENH:  Adding more stuff.

@@
 \usepackage{float}
+
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \graphicspath{{./Figures/}
-                          }
+              {./Results/}}
 
@@
 \def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Tustison {et~al.}} %use et al only if is more than 1 author
-\def\Authors{Nicholas J. Tustison\,$^{1,*}$ and Brian A. Avants\,$^{2}$
+%\def\firstAuthorLast{Tustison {et~al.}} %use et al only if is more than 1 author
+\def\firstAuthorLast{Tustison and Avants} %use et al only if is more than 1 author
+\def\Authors{Nicholas J. Tustison\,$^{1,*}$ and Brian B. Avants\,$^{2}$
  }
@@
 
-\title[Explicit B-spline Regularization]{Explicit B-spline Regularization in Diffeomorphic Image Registration}
+\title[Explicit B-spline regularization]{Explicit B-spline regularization in diffeomorphic image registration}
 \author[\firstAuthorLast ]{\Authors}
@@
 Free-form deformation (FFD) image registration, characterized by  regularization based on the B-spline basis functions, has several
-advantages including ease of algorithmic understanding, 
-good performance, desirable geometric properties, and several available 
-implementations.  Current research was preceded by related work for
+advantages including algorithmic simplicity,
+good performance, desirable properties, and several available
+implementations.  Current research was
+preceded by related work for
 geometric modeling \citep{sederberg1986} and originated with such important
@@
 the works of \cite{bro-nielsen1996} and \cite{pennec1999} who observe that
-Gaussian smoothing, characteristic of Demons, of the update or displacement
+smoothing via Gaussian convolution, a defining characteristic of Demons,
+of the update or displacement
 field is a greedy approximation for solving the partial differential equations governing
@@
 can be viewed as a type of fluid-like Demons approach
-where, rather than produced by Gaussian convolution, regularized fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.
-This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
-As mentioned in \cite{tustison2009}, this preconditioning is used to reduce the hemstitching induced by the traditional gradient-based FFD formulation.
+where, rather than projecting the update field to the space
+of regularized fields using Gaussian convolution, gradient fields
+are projected to a smooth space characterized by the B-spline
+basis functions.
+This analogy was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
+This preconditioning is used to mitigate the hemstitching effect induced by the ill-conditioned nature of the traditional gradient-based FFD formulation.
 Further details can be gleaned from the ITK implementation of this
@@
 the well-known Large Deformation Diffeomorphic Metric Mapping (LDDMM) algorithm
-was proposed in \cite{Beg2005}.  In contrast to the mapping produced
+was proposed in \cite{beg2005}.  In contrast to the mapping produced
 by \cite{christensen1996}, LDDMM yields the geodesic solution in the space of diffeomorphisms between two images. Since its introduction, LDDMM has
-inspired much innovation in the image registration literature.  
-Symmetrical considerations in the velocity field parameterization, for example, are discussed in \cite{avants2008} in the context of a cross correlation similarity metric which minimizes the bias of the resulting transformation in selecting the identity of the fixed and moving images.
-A greedy version of this algorithm which has proven successful in general application \citep{Klein2009}.
+inspired much innovation in the image registration literature.  Applying
+the log-Euclidean framework of \cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using Exponential Lie algebra) uses a constant velocity field parameterization to provide a fast, diffeomorphic alternative \citep{ashburner2007}.
+Additionally, symmetrical considerations in the velocity field parameterization are discussed in \cite{avants2008} in the context of a cross correlation similarity metric which minimizes the bias of the resulting transformation when selecting the ``fixed'' and ``moving'' images.
+A greedy version of this algorithm has proven successful in neuroimaging \citep{klein2009} and pulmonary \citep{murphy2011} applications.
 
@@
 
-Alternatively, the FFD 
-variant reported in \cite{rueckert2006} enforced diffeomorphic transforms
+Although many extensions of LDDMM rely
+on some form of Gaussian convolution for
+regularization \citep[e.g.][]{risser2011}, there has been significant interest in constraining
+FFD approaches to the space of diffeomorphisms.
+An early attempt reported in \cite{rueckert2006} enforced
+diffeomorphic transforms
 by concatenating multiple FFD transforms, each of which is constrained
-to describe a one-to-one mapping.  Another FFD registration
-incorporated the recent log-Euclidean framework for enforcing diffeomorphic
-transformations \cite{Modat2011}.
-Recently, the work of 
-\cite{de-craene2011} combined these registration concepts into a single
-framework called {\em temporal free-form deformation} in which the 
-time-varying velocity field characteristic of LDDMM-style algorithms
-is modeled using a 4-D B-spline object (3-D + time).  Integration of 
-the velocity field yields the mapping between parameterized time points.
+to describe a one-to-one mapping.
+\cite{modat2011} incorporated the log-Euclidean
+framework for enforcing diffeomorphic transformations and ensuring invertibility.  Similarly, the work of \cite{de-craene2011} provided
+a full LDDMM-style algorithm based on B-splines called
+{\em temporal free-form deformation} in which the
+time-varying velocity field
+is modeled using a 4-D B-spline object (3-D + time).  Numerical integration of the mapping propagated within the velocity field yields the transform between parameterized time points.
 
-%
-%
-%framework developed independently of Demons and LDDMM/DARTEL.
-%But then the FFD people saw how useful the diff. properties were and
-%tried to integrate B-splines into the diff. framework.
-%
-%
-%The early deformable work of 
-%\begin{itemize}
-%  \item xx Early deformable work \cite{broit1981,bajcsy1989}.
-%  \item FFD \cite{Sleziski,Rueckert1989}
-%  \item Christensen's fluid work
-%  \item Optical flow?
-%  \item Fast fluid registration of medical images (Bro-Nielsen and Gramkow).
-%  \item Consistent image registration
-%  \item Demons and Diff demons
-%  \item LDDMM 
-%  \item DARTEL
-%  \item Arno Klein's study
-%\end{itemize}
-%
-%
-%
-%Early work included the elastic deformation
+As alluded to earlier, B-spline approximation can also be used
+for regularizing vector fields in an analogous fashion as
+Gaussian convolution.  In this vein and similar to
+\cite{de-craene2011}, we reported in \cite{tustison2012a,tustison2012}
+the use of an $n$-D + time B-spline object to
+represent the characteristic velocity fields.  However, we use the
+directly manipulated free-form deformation formulation to improve the solution convergence and actual results.  This also facilitates modeling
+temporal periodicity and the enforcement of stationary boundaries.
 
-%Significant algorithmic developments characterizing modern intensity-based
-%image registration research include the B-spline parameterized approach
-%(so called {\em free-form deformation}) with early contributions 
-%including \cite{szeliski1997,thevenaz1998,rueckert1999}.  Amongst the 
-%variant extensions, the 
-%{\em directly manipulated free-form deformation} approach  \cite{tustison2009} 
-% addressed the hemstitching issue associated with steepest descent traversal of
-%problematic energy topographies during the course of optimization.
-%%caused by the distribution of the uniform 
-%%B-spline shape functions over the transformation domain 
-%
-%
-%In this work, we describe our extension to these methods.  Similar to 
-%\cite{de-craene2011}, we also use an $N$-D + time B-spline object to 
-%represent the characteristic velocity field.  However, we use the 
-%directly manipulated free-form deformation optimization formulation to improve 
-%convergence during the course of optimization.  This also facilitates
-%modeling temporal periodicity and
-%the enforcement of stationary boundaries consistent with diffeomorphic
-%transforms.
-%We also incorporate B-spline mesh multi-resolution capabilities
-%for increased control during registration progression.  
-%Most importantly, we also describe
-%the parallelized algorithmic implementation as open source available through the Insight Toolkit.%
-%\footnote{
-%http://www.itk.org/
-%}
-%
-%We first describe the methodology by laying out a mathematical description 
-%of the various algorithmic elements coupled with implementation details
-%where appropriate.  This is followed by an evaluation on publicly available
-%brain data.
+Both this latter work \citep{tustison2012a} and our earlier work
+\citep{tustison2009} demonstrate that our DMFFD framework is potentially
+applicable to the
+entire gamut of diffeomorphic registration algorithms and provides
+unique ``flavors'' of smoothing possibilities.  However, differences
+between the two approaches produce characteristically different
+solutions.  In addition to
+smoothing kernel differences, Gaussian
+convolution tends to ``flatten'' the signal in contrast to an
+approximation or fitting of the signal provided by the DMFFD
+approach.  Also, whereas
+convolution operates entirely within the discrete space,
+the fitting process constructs a continuous object prior
+to any discrete sampling.  Further elaboration on these differences
+are discussed in subsequent sections.
+
+Variants of three popular diffeomorphic algorithms and their DMFFD analogs
+(LDDMM%
+\footnote{
+\begin{itemize}
+\item http://www.itk.org/Doxygen/html/classitk\_1\_1TimeVaryingVelocityFieldTransform.html
+\item http://www.itk.org/Doxygen/html/classitk\_1\_1TimeVaryingBSplineVelocityFieldTransform.html
+\end{itemize}
+},
+DARTEL%
+\footnote{
+\begin{itemize}
+\item http://www.itk.org/Doxygen/html/classitk\_1\_1GaussianExponentialDiffeomorphicTransform.html
+\item http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineExponentialDiffeomorphicTransform.html
+\end{itemize}
+},
+ and SyN%
+\footnote{
+\begin{itemize}
+\item http://www.itk.org/Doxygen/html/classitk\_1\_1SyNImageRegistrationMethod.html
+\item http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineSyNImageRegistrationMethod.html
+\end{itemize}
+})
+were implemented by the authors
+as part of the recent refactoring of the open source
+Insight Toolkit (ITK) although related work had been
+previously implemented within the popular Advanced
+Normalization Tools (ANTs).%
+\footnote{
+http://stnava.github.io/ANTs
+}
+Given the popularity and excellent performance of the symmetric normalization (SyN) algorithm, our focus in this work is
+its B-spline equivalent which we denote as ``B-spline SyN.''
+Evaluation and comparisons of the respective algorithmic instantiations
+are performed using the \verb#antsRegistration# program found
+in the ANTs repository (also originally developed by the authors \citep{tustison2012}).  This permits a direct algorithmic comparison
+as potential sources for implementation bias have been removed \citep{tustison2013}.  Additionally, in the spirit of open science,
+all text, figures, and
+scripts to reproduce the results contained in this work are publicly available online.%
+\footnote{
+https://github.com/ntustison/BSplineMorphisms
+}
 
 \section{Material and Methods}
-%Provide sufficient detail to allow the work to be reproduced. Methods already published should be indicated by a reference: only relevant modifications should be described.
 
-Three popular algorithms [1,2,3] (and their B-spline analogs) were implemented as part of the recent refactoring of the Insight Toolkit (ITK) [5] and included as part of the Advanced Normalization Tools (ANTs) repository [6].
-Thus, all code used for this work is available as open source.
-For comparison, the ITK-based, well-vetted SyN algorithm [3], which uses explicit Gaussian regularization, was used to provide the canonical mappings in the MICCAI 2012 multi-atlas labeling challenge [7] between 20 training subjects and 15 testing subjects (140 labeled structures).
-Using the exact same parameters, except those specific to the regularization, we produced the analogous B-spline SyN mappings for the same data and compared transformations using the Dice overlap.
+\subsection{Theoretical Overview}
 
-\subsection{Public Data}
+Given the spatial domain $\Omega$ of $d-$dimensionality
+defined over image $I$, a diffeomorphic
+mapping, $\phi$, parameterized over $t \in [0,1]$ transforms the image
+$I$ to the target image $J$ using $I\circ\phi(\mathbf{x},1)$ where the
+geodesic path $\phi(\mathbf{x},t)$ is described by \citep{beg2005}
+\begin{align}
+  \inf_{\phi} \left( \int_0^1 \|v(t)\|_L^2 dt +
+                     \int_{\Omega} | I \circ \phi^{-1}(\mathbf{x},1) - J |^2 d\Omega
+              \right).
+  \label{eq:lddmm}
+\end{align}
+$\phi$ is generated as the solution of the ordinary differential equation
+\begin{align}
+  \frac{d \phi(\mathbf{x},t)}{dt} = v( \phi(\mathbf{x},t), t ),\,\, \phi(\mathbf{x},0) = \mathbf{Id}
+\label{eq:ode}
+\end{align}
+where $v$ is a time-dependent smooth field (as dictated by the functional norm $L$), $v : \Omega \times t
+\rightarrow \mathrm{R}^d$ parameterized by $t \in [0,1]$.  Diffeomorphic mappings
+between parameterized time points $\{t_a,t_b\} \in [0,1]$
+are obtained from  Eq. (\ref{eq:ode}) through integration of the transport
+equation, viz.
+\begin{align}
+  \label{eq:integral}
+\phi(\mathbf{x},t_b) &= \phi(\mathbf{x},t_a) + \int_{t_a}^{t_b} v(\phi(\mathbf{x}), t) dt.
+\end{align}
+%and $L$ is the functional norm which enforces velocity field regularity.
+
+However,
+as pointed out in \cite{avants2008}, implementations of this standard
+LDDMM formulation are negatively affected by the lack of optimization
+symmetry where arbitrary assignment of fixed and moving images could
+lead to different solutions despite the fact that the theoretical
+geodesic solution describes the same path forwards and backwards.
+This observation led to the symmetric
+formulation of Eq. (\ref{eq:lddmm}) found in \cite{avants2008}:
+\begin{align}
+  \inf_{\phi_1} \inf_{\phi_2} \left(
+                     \int_0^{0.5} \left( \|v_1(t)\|_L^2 + \|v_2(1-t)\|_L^2 \right) dt +
+                     \int_{\Omega} | I \circ \phi_1^{-1}(\mathbf{x},0.5)
+                           - J \circ \phi_2^{-1}(\mathbf{x},0.5) |^2 d\Omega
+              \right)
+\end{align}
+where
+\begin{align}
+  \frac{d \phi_i(\mathbf{x},t)}{dt} = v_i( \phi_i(\mathbf{x},t), t ),\,\, \phi_i(\mathbf{x},0) = \mathbf{Id}, \,\, i \in \{1,2\}
+\end{align}
+with extension to arbitrary similarity metric choice, i.e. the second
+term is replaced with
+\begin{align}
+\int_{\Omega} \Pi_{\sim}
+                          \left( I \circ \phi_1^{-1}(\mathbf{x},0.5),
+                           J \circ \phi_2^{-1}(\mathbf{x},0.5) \right) d\Omega
+\end{align}
+with a popular choice for $\Pi_{\sim}$ being a local neighborhood cross
+correlation \citep{avants2008,avants2011}.
+Note that
+$t$ is parameterized in opposite directions between $\phi_1$ and $\phi_2$.
+A diagrammatic
+illustration of the explicit symmetry associated with SyN is shown in
+Figure \ref{fig:syn}.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{SyN.pdf}
+  \caption{Illustration of the greedy SyN formulation.  Given images
+  $I_A$ and $I_B$, the symmetric set-up requires finding the two
+  transform pairs $\left(\phi_1,\phi_1^{-1}\right)$
+  $\left(\phi_2,\phi_2^{-1}\right)$ which map to/from
+  the respective images to the midway point. During
+  optimization, the update field at each iteration is
+  determined from the metric field gradient taken at
+  the midway point, i.e.
+  $\nabla \Pi_{\sim} \left(I\circ\phi_1(0.5),J\circ\phi_2(0.5)\right)$.
+  The full
+  forward and inverse transforms are found through
+  composition, i.e. $\phi=\phi_1 \circ \phi_2^{-1}$ and
+  $\phi^{-1}=\phi_2 \circ \phi_1^{-1}$.
+  }
+  \label{fig:syn}
+\end{figure}
+
+\subsubsection{St. Nava's Theory of Greed and Original SyN}
+
+Although presenting a rigorous framework for image registration solutions
+with desirable properties, the complexity of these diffeomorphic methodologies
+requires substantial computational resources.  For typical 3-D neuroimaging
+applications, the corresponding solutions require numerical integration over
+and storage of 4-D velocity fields at each iteration which is limiting for
+many common, contemporary processing infrastructures.
+
+Therefore, in addition to the full-scale
+symmetric SyN offering described in \cite{avants2008}, the authors therein
+provided a ``greedy'' alternative which has demonstrated superior performance
+in various applications \citep{avants2011,klein2009,murphy2011}%
+\footnote{
+SyN was also the standard registration for the MICCAI 2013
+Workshop on Segmentation:  Algorithms Theory, and Applications (SATA)
+and corresponding challenge (https://masi.vuse.vanderbilt.edu/workshop2013).
+}
+while simultaneously being capable of running with limited computational
+resources.  In this variant, the discrete time-parameterized
+velocity field samples are restricted to their respective
+endpoints, i.e. the $v_i(\mathbf{x},t)$ are sampled at $t \in \{0,0.5,1\}$
+implying simultaneous storage of only four transform vector fields
+$\phi_1(\mathbf{x})$,
+$\phi_1^{-1}(\mathbf{x})$, $\phi_2(\mathbf{x})$, and $\phi_2^{-1}(\mathbf{x})$
+(cf Figure \ref{fig:syn}).  The algorithmic steps are briefly sketched
+in Algorithm \ref{alg:syn}.
+
+\begin{algorithm}
+\caption{SyN algorithmic overview}
+\label{alg:syn}
+\begin{algorithmic}
+\State $\phi_i \leftarrow \mathbf{Id}$, $\phi_i^{-1} \leftarrow \mathbf{Id}$
+\Comment $i \in \{1,2\}$
+\State $n \leftarrow 1$
+\While {not converged}
+  \State $v_1^n \leftarrow \nabla \Pi_{\sim} \left(I\circ\phi_1^{n-1},J\circ\phi_2^{n-1}\right)$
+  \State $v_2^n \leftarrow \nabla \Pi_{\sim} \left(J\circ\phi_2^{n-1},I\circ\phi_1^{n-1}\right)$
+  \State $v_i^n \leftarrow S_v( v_i^n )$      \Comment $S_v$ is a smoothing operation on the update transform field
+  \State $\phi_i^n \leftarrow S_\phi( v_i^n \circ \phi_i^{n-1} )$      \Comment $S_\phi$ is a smoothing operation on the total transform field
+  \State $\left(\phi_i^n\right)^{-1} \leftarrow Inv\left(\phi_i^n, \left(\phi_i^{n-1}\right)^{-1}\right)$
+    \Comment Inverse field estimation described in \cite{avants2008}
+  \State $n \leftarrow n + 1$
+\EndWhile
+\State \Return $\phi \leftarrow \phi_1 \circ \phi_2^{-1}$, $\phi^{-1} \leftarrow \phi_2 \circ \phi_1^{-1}$
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Directly Manipulated Free-Form Deformation Diffeomorphic Analogs}
+
+Although several velocity field regularization
+operators have been proposed, many algorithmic instantiations
+default to Gaussian smoothing due to its simplicity both in implementation
+and complexity terms.
+Extending the work presented in \cite{tustison2009,tustison2012},
+the DMFFD approach is a viable and practical alternative
+for explicit regularization in diffeomorphic image registration.
+
+As we mentioned previously, it has been shown that the preconditioned
+gradient of the B-spline object is equivalent to a B-spline approximation
+smoothing operation on the spatial derivative of the similarity metric
+gradient \citep{tustison2009}. In the case of
+$d$-dimensional image registration, the time-dependent velocity field can be
+represented
+as a $(d + 1)$-dimensional B-spline object
+\begin{align}
+v(\mathbf{x}, t) = \sum_{i_1=1}^{X_1}\ldots\sum_{i_d=1}^{X_d}\sum_{i_t=1}^T v_{i_1,\ldots,i_d,i_t} B_{i_t}(t) \prod_{j=1}^d B_{i_j}(x_j)
+\end{align}
+where $v_{i_1,\ldots,i_d,i_t}$ is a $(d+1)$-dimensional control point lattice
+characterizing the velocity field and $B(\cdot)$ are the univariate B-spline
+basis functions separately modulating regularity in the solution for each parametric dimension.  The preconditioned gradient analog for updating the velocity
+field is as follows \citep{tustison2012}:
+$\delta v_{i_1,\ldots,i_d,i_t}$, given the similarity metric, $\Pi_\sim$,
+\begin{align}
+\label{eq:dmffd}
+  \delta v_{i_1,\ldots,i_d,i_t} &= \left( \sum_{c=1}^{N_{\Omega} \times N_t} \left( \frac{\partial \Pi_\sim}{\partial \mathbf{x}} \right)_c B_{i_t}(t^c)\prod_{j=1}^d B_{i_j}(x_j^c)  \right.
+\nonumber \\
+  &\cdot \left. \frac{B_{i_t}^2(t^c) \prod_{j=1}^d B_{i_j}^2 (x_j^c)}
+  {\sum_{k_1=1}^{r+1}\ldots\sum_{k_d=1}^{r+1} \sum_{k_t=1}^{r+1} B_{k_t}^2(t^c)
+  \prod_{j=1}^d B_{k_j}^2 (x_j^c)} \right) \nonumber \\
+  &\cdot\left({\sum_{c=1}^{N_{\Omega}\times N_t}B_{i_t}^2(t^c) \prod_{j=1}^d B_{i_j}^2 (x_j^c)} \right) ^{-1}
+\end{align}
+which is a slight modification of Eqn. (21) in \cite{tustison2009}
+which takes into account the temporal locations of the
+dense gradient field sampled in $t \in [0,1]$. $N_t$ and $N_\Omega$ are the number
+of time point samples and the number of voxels in the reference image domain, respectively.
+$r$ is the spline order in all dimensions%
+\footnote{
+In terms of implementation spline orders can be specified separately for each dimension but, for simplicity,
+we only specify a single spline order.
+}
+and $c$ indexes the spatio-temporal dense metric gradient sample.
+Additionally, in \cite{tustison2006} it was shown that one could associate each
+a sample, $\left( \frac{\partial \Pi}{\partial \mathbf{x}}\right)_c$ with a confidence
+weighting.  Thus, in order to enforce stationary boundaries, we
+assign image boundary metric gradients a value of zero with a corresponding
+large confidence value.
 
@@
 
+Although various methods exist for solving Eqns. (\ref{eq:ode}) and (\ref{eq:integral}),
+we use $4^{th}$-order Runge-Kutta, i.e.
+\begin{align}
+  \phi_{n+1} &= \phi_{n} + \frac{1}{6}\left( k_1 + 2k_2 + 2k_3 + k_4 \right) \\
+  t_{n+1} &= t_{n} + \Delta t \\
+  \phi_0 &= \phi(t_0)
+\end{align}
+where
+\begin{align}
+  k_1 &= v\left( \phi_{n}, t_{n} \right)\Delta t \\
+  k_2 &= v\left( \phi_{n} + \frac{k_1}{2}, t_{n} + \frac{\Delta t}{2} \right)\Delta t \\
+  k_3 &= v\left( \phi_{n} + \frac{k_2}{2}, t_{n} + \frac{\Delta t}{2} \right)\Delta t \\
+  k_4 &= v\left( \phi_{n} + k_3, t_{n} + \Delta t \right)\Delta t
+\end{align}
+which provides a more stable and reliable alternative than other numerical
+methods \citep{press2007}.
+
+\subsection{Evaluation Data}
+
+We take inspiration from the well-known
+Klein comparative study in which 14 image registration
+algorithms were evaluated based on performance
+on publicly available labeled brain data \citep{klein2009}.
+Based on the final rankings, the SyN algorithm is one
+of the top-performing registration algorithms.  Its relative
+performance coupled with its public availability has made
+it a de facto standard for algorithmic comparison.
+
+For our evaluation, we used the same data as that in the
+Klein study.  Specifically, we used the data sets denoted as:
+\begin{itemize}
+  \item MGH10
+  \item CUMC12
+  \item IBSR18
+  \item LPBA40
+\end{itemize}
+which are available for download from Arno Klein's website.%
+\footnote{
+http://mindboggle.info/papers/evaluation\_NeuroImage2009.php
+}
+We also use the labeled brain data provided at the
+MICCAI 2012 Grand Challenge and Workshop on Multi-Atlas
+Labeling%
+\footnote{
+https://masi.vuse.vanderbilt.edu/workshop2012
+} which we denote as MICCAI35.  This T1-weighted
+MRI data set consists of 35 subject MRI taken from the
+Oasis database%
+\footnote{
+http://www.oasis-brains.org
+}.  The corresponding labels were provided by Neuromorphometrics,
+Inc.%
+\footnote{
+http://Neuromorphometrics.com/
+}
+under academic subscription.
+
 \section{Results}
 %Results should be clear and concise.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{allCUMC.pdf}
+  \caption{}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{allIBSR.pdf}
+  \caption{Currently redoing on the cluster.}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{allLPBA.pdf}
+  \caption{}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{allMGH.pdf}
+  \caption{Currently redoing on the cluster.}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{allMICCAI.pdf}
+  \caption{}
+\end{figure}
 

### 2013-10-19 paper.tex
ENH:  Finished first draft.

@@
 % \usepackage{amsthm}
-
- \usepackage{makecell}
 
@@
 
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
+
 \copyrightyear{}
@@
 
-\title[Explicit B-spline regularization]{Explicit B-spline regularization in diffeomorphic image registration}
+\title[DMFFD Diffeomorphic Image Registration]{Explicit B-spline regularization in diffeomorphic image registration}
 \author[\firstAuthorLast ]{\Authors}
@@
 As alluded to earlier, B-spline approximation can also be used
-for regularizing vector fields in an analogous fashion as
+for regularizing time-varying vector fields in an analogous fashion as
 Gaussian convolution.  In this vein and similar to
@@
 represent the characteristic velocity fields.  However, we use the
-directly manipulated free-form deformation formulation to improve the solution convergence and actual results.  This also facilitates modeling
+directly manipulated free-form deformation formulation to improve the solution convergence.  This also facilitates modeling
 temporal periodicity and the enforcement of stationary boundaries.
-
-Both this latter work \citep{tustison2012a} and our earlier work
-\citep{tustison2009} demonstrate that our DMFFD framework is potentially
+Both this work and our earlier work
+\citep{tustison2009} demonstrate that the DMFFD framework is potentially
 applicable to the
@@
 the fitting process constructs a continuous object prior
-to any discrete sampling.  Further elaboration on these differences
-are discussed in subsequent sections.
+to any discrete sampling.  
+Interestingly, a similar comparison was made with respect to 
+Gaussian derivative estimation  \citep{bourma2007}.  Although 
+typically estimated using  truncated, discrete Gaussian 
+convolution, an alternative based on 
+B-spline approximation demonstrated superior performance with
+similar computational cost.  
+%Further elaboration on these differences
+%are discussed in subsequent sections.
 
 Variants of three popular diffeomorphic algorithms and their DMFFD analogs
-(LDDMM%
-\footnote{
-\begin{itemize}
-\item http://www.itk.org/Doxygen/html/classitk\_1\_1TimeVaryingVelocityFieldTransform.html
-\item http://www.itk.org/Doxygen/html/classitk\_1\_1TimeVaryingBSplineVelocityFieldTransform.html
-\end{itemize}
-},
-DARTEL%
-\footnote{
-\begin{itemize}
-\item http://www.itk.org/Doxygen/html/classitk\_1\_1GaussianExponentialDiffeomorphicTransform.html
-\item http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineExponentialDiffeomorphicTransform.html
-\end{itemize}
-},
- and SyN%
-\footnote{
-\begin{itemize}
-\item http://www.itk.org/Doxygen/html/classitk\_1\_1SyNImageRegistrationMethod.html
-\item http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineSyNImageRegistrationMethod.html
-\end{itemize}
-})
+(LDDMM,
+DARTEL,
+ and SyN)
 were implemented by the authors
@@
 are performed using the \verb#antsRegistration# program found
-in the ANTs repository (also originally developed by the authors \citep{tustison2012}).  This permits a direct algorithmic comparison
+in the ANTs repository (also originally developed by the authors).  This permits a direct algorithmic comparison
 as potential sources for implementation bias have been removed \citep{tustison2013}.  Additionally, in the spirit of open science,
@@
 $\phi_1^{-1}(\mathbf{x})$, $\phi_2(\mathbf{x})$, and $\phi_2^{-1}(\mathbf{x})$
-(cf Figure \ref{fig:syn}).  The algorithmic steps are briefly sketched
+(cf Figure \ref{fig:syn}).  Since the greedy SyN framework is the focus of the evaluation,
+the algorithmic steps are briefly sketched
 in Algorithm \ref{alg:syn}.
@@
 \begin{algorithm}
-\caption{SyN algorithmic overview}
+\caption{Greedy SyN algorithm}
 \label{alg:syn}
@@
 default to Gaussian smoothing due to its simplicity both in implementation
-and complexity terms.
-Extending the work presented in \cite{tustison2009,tustison2012},
-the DMFFD approach is a viable and practical alternative
-for explicit regularization in diffeomorphic image registration.
+and complexity terms.  A viable and practical alternative is the DMFFD 
+approach based on B-splines for explicit regularization of vector fields.
+As presented in \cite{tustison2009}, given the similarity metric
+$\Pi_{\sim}$, the $d$-dimensional update field (i.e. preconditioned
+gradient field), $\delta v_{i_1,\ldots,i_d}$, is calculated from
+\begin{align}
+\label{eq:dmffd}
+  \delta v_{i_1,\ldots,i_d} = \frac{ \sum_{c=1}^{N_{\Omega}} \left( \frac{\partial \Pi_\sim}{\partial \mathbf{x}} \right)_c \prod_{j=1}^d B_{i_j}(x_j^c)  
+  \cdot \frac{\prod_{j=1}^d B_{i_j}^2 (x_j^c)}
+  {\sum_{k_1=1}^{r+1}\ldots\sum_{k_d=1}^{r+1} 
+  \prod_{j=1}^d B_{k_j}^2 (x_j^c)} }
+  {\sum_{c=1}^{N_{\Omega}} \prod_{j=1}^d B_{i_j}^2 (x_j^c)}
+\end{align}
+where the set of $B(\cdot)$ are the univariate B-spline basis functions for
+separately modulating regularity in the solution for each parametric dimension,
+$N_\Omega$ is the number of voxels in the reference image domain,
+$r$ is the spline order in all dimensions,%
+\footnote{
+In terms of implementation spline orders can be specified separately for each dimension but, for simplicity,
+we only specify a single spline order.
+}
+and $\frac{\partial \Pi_\sim}{\partial \mathbf{x}}$ is the spatial 
+similarity metric gradient indexed by $c$.
 
-As we mentioned previously, it has been shown that the preconditioned
-gradient of the B-spline object is equivalent to a B-spline approximation
-smoothing operation on the spatial derivative of the similarity metric
-gradient \citep{tustison2009}. In the case of
-$d$-dimensional image registration, the time-dependent velocity field can be
-represented
+Similarly, in the case of $d$-dimensional time-parameterized 
+diffeomorphic image registration, 
+the time-dependent velocity field can be represented
 as a $(d + 1)$-dimensional B-spline object
@@
 where $v_{i_1,\ldots,i_d,i_t}$ is a $(d+1)$-dimensional control point lattice
-characterizing the velocity field and $B(\cdot)$ are the univariate B-spline
-basis functions separately modulating regularity in the solution for each parametric dimension.  The preconditioned gradient analog for updating the velocity
-field is as follows \citep{tustison2012}:
+characterizing the velocity field.  The preconditioned gradient analog of
+Equation (\ref{eq:dmffd}) for updating the time-varying velocity field control point lattice
+is as follows:
 $\delta v_{i_1,\ldots,i_d,i_t}$, given the similarity metric, $\Pi_\sim$,
 \begin{align}
-\label{eq:dmffd}
+\label{eq:tvdmffd}
   \delta v_{i_1,\ldots,i_d,i_t} &= \left( \sum_{c=1}^{N_{\Omega} \times N_t} \left( \frac{\partial \Pi_\sim}{\partial \mathbf{x}} \right)_c B_{i_t}(t^c)\prod_{j=1}^d B_{i_j}(x_j^c)  \right.
@@
 \end{align}
-which is a slight modification of Eqn. (21) in \cite{tustison2009}
 which takes into account the temporal locations of the
@@
 of time point samples and the number of voxels in the reference image domain, respectively.
-$r$ is the spline order in all dimensions%
 \footnote{
-In terms of implementation spline orders can be specified separately for each dimension but, for simplicity,
-we only specify a single spline order.
-}
-and $c$ indexes the spatio-temporal dense metric gradient sample.
 Additionally, in \cite{tustison2006} it was shown that one could associate each
 a sample, $\left( \frac{\partial \Pi}{\partial \mathbf{x}}\right)_c$ with a confidence
-weighting.  Thus, in order to enforce stationary boundaries, we
+weighting.  Thus, in order to enforce stationary boundaries for all described
+DMFFD-based diffeomorphic algorithms, we
 assign image boundary metric gradients a value of zero with a corresponding
 large confidence value.
+}
+
+For regularization of constant velocity fields, e.g. SyN or DARTEL, updating the 
+field is performed using Equation (\ref{eq:dmffd}).  In the case B-spline SyN, 
+this applies to the smoothing operators $S_{v}$ and $S_{\phi}$ in Algorithm \ref{alg:syn}
+although best performance (at least for the data described in this work) typically
+employs no smoothing on the total transform field, i.e. $S_\phi$ is such that $S_\phi( \phi_i^n ) = \phi_i^n$.
 
@@
 
-Although various methods exist for solving Eqns. (\ref{eq:ode}) and (\ref{eq:integral}),
-we use $4^{th}$-order Runge-Kutta, i.e.
-\begin{align}
-  \phi_{n+1} &= \phi_{n} + \frac{1}{6}\left( k_1 + 2k_2 + 2k_3 + k_4 \right) \\
-  t_{n+1} &= t_{n} + \Delta t \\
-  \phi_0 &= \phi(t_0)
-\end{align}
-where
-\begin{align}
-  k_1 &= v\left( \phi_{n}, t_{n} \right)\Delta t \\
-  k_2 &= v\left( \phi_{n} + \frac{k_1}{2}, t_{n} + \frac{\Delta t}{2} \right)\Delta t \\
-  k_3 &= v\left( \phi_{n} + \frac{k_2}{2}, t_{n} + \frac{\Delta t}{2} \right)\Delta t \\
-  k_4 &= v\left( \phi_{n} + k_3, t_{n} + \Delta t \right)\Delta t
-\end{align}
-which provides a more stable and reliable alternative than other numerical
-methods \citep{press2007}.
+Three diffeomorphic algorithms described previously utilizing Gaussian convolution smoothing and their DMFFD counterparts are available through the following classes in the ITK repository: 
+\begin{itemize}
+  \item {\bf LDDMM:}
+    \begin{itemize}
+      \item {\tt itk::TimeVaryingVelocityFieldTransform} 
+      \item {\tt itk::TimeVaryingVelocityFieldTransformParametersAdaptor}
+      \item {\tt itk::TimeVaryingVelocityFieldImageRegistrationMethodv4}
+      \item {\tt itk::GaussianSmoothingOnUpdateTimeVaryingVelocityFieldTransform}
+      \item {\tt itk::TimeVaryingBSplineVelocityFieldTransform} 
+      \item {\tt itk::TimeVaryingBSplineVelocityFieldImageRegistrationMethodv4}
+      \item {\tt itk::TimeVaryingBSplineVelocityFieldTransformParametersAdaptor}
+    \end{itemize}  
+  \item {\bf DARTEL:}
+    \begin{itemize}
+       \item{\tt itk::ConstantVelocityFieldTransform}
+       \item{\tt itk::ConstantVelocityFieldTransformParametersAdaptor}
+      \item {\tt itk::GaussianExponentialDiffeomorphicTransform} 
+      \item {\tt itk::GaussianExponentialDiffeomorphicTransformParametersAdaptor}
+      \item {\tt itk::BSplineExponentialDiffeomorphicTransform} 
+      \item {\tt itk::BSplineExponentialDiffeomorphicTransformParametersAdaptor}
+    \end{itemize}  
+  \item {\bf SyN:}
+    \begin{itemize}
+      \item {\tt itk::SyNImageRegistrationMethod} 
+      \item {\tt itk::BSplineSyNImageRegistrationMethod} 
+    \end{itemize}  
+\end{itemize}
+These and other classes (e.g. similarity metrics, optimization methods, and
+utility classes) were developed as part of the ITKv4 registration
+framework refactoring.  Further details can be found in the documentation
+provided within the classes themselves.%
+\footnote{
+http://www.itk.org/Doxygen/html/classes.html
+}
+
+The class {\tt itk::BSplineScatteredDataPointSetToImageFilter} underlies all DMFFD regularization which implements the 
+methods described in \cite{tustison2009}.  Although applicable to
+various scenarios (e.g. curve and surface estimation), it has
+been optimized for imaging applications and multi-threaded
+for fast processing on multi-core machines.
+Additionally, numerical integration for solving Equations (\ref{eq:ode}) and (\ref{eq:integral}) utilizes Runge-Kutta which
+provides a more stable alternative than other
+methods \citep{press2007}.  Implementation is provided in the class {\tt itk::TimeVaryingVelocityFieldIntegrationImageFilter}.
+
+A complete packaging of these classes has been made
+available as part of our Advanced Normalization Tools (ANTs) toolkit.%
+\footnote{
+http://stnava.github.io/ANTs/
+}
+The \verb#antsRegistration# program%
+\footnote{
+The long command and short command line help menus can be invoked by
+the commands `{\tt antsRegistrationantsRegistration --help}' and `{\tt antsRegistration -h}', respectively.
+}
+takes advantage of the enhanced ITKv4
+registration framework and was developed by the 
+authors to provide a robust and versatile solution for a wide variety of
+image registration applications. The basic conceptualization for use is that one 
+can
+set-up any number of registration ``stages'' with each stage being 
+characterized by a specified transform.  For example, a representative 
+command call is as follows:
+\vspace{2mm}
+\lstset{frame = htb,
+        framerule = 0.25pt,
+        float,
+        fontadjust,
+        backgroundcolor={\color{listlightgray}},
+        basicstyle = {\ttfamily\scriptsize},
+        keywordstyle = {\ttfamily\color{listkeyword}\textbf},
+        identifierstyle = {\ttfamily},
+        commentstyle = {\ttfamily\color{listcomment}\textit},
+        stringstyle = {\ttfamily},
+        showstringspaces = false,
+        showtabs = false
+        numbers = none,
+        numbersep = 6pt,
+        numberstyle={\ttfamily\color{listnumbers}},
+        tabsize = 2,
+        language=bash,
+        floatplacement=!h,
+        caption={Representative script containing {\tt antsRegistration} and {\tt antsApplyTransforms} command calls used for evaluation},
+        captionpos=b,
+        label=listing:script
+        }
+\begin{lstlisting}
+# Register the $fixed and $moving images with initial
+# alignment of the centers of intensity followed by 
+# the following three stages:
+#   rigid -> affine -> B-spline SyN
+
+antsRegistration --dimensionality 3 \
+                 --output ${prefix} \
+                 --use-histogram-matching 1 \
+                 --initial-moving-transform [${fixed},${moving},1] \
+                 --transform Rigid[0.1] \
+                 --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
+                 --convergence 1000x500x250 \
+                 --smoothing-sigmas 2.0x1.0x0.0mm \
+                 --shrink-factors 4x3x2 \
+                 --transform Affine[0.1] \
+                 --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
+                 --convergence 1000x500x250 \
+                 --smoothing-sigmas 2.0x1.0x0.0mm \
+                 --shrink-factors 4x3x2 \
+                 --transform BSplineSyN[0.1,10x10x10,0x0x0,3] \
+                 --metric CC[${fixed},${moving},1,4] \ 
+                 --convergence 100x70x50x20 \
+                 --smoothing-sigmas 2x1.0x0.5x0.0mm \
+                 --shrink-factors 6x4x2x1
+
+# Apply the resulting transforms (generic affine + 
+# B-spline SyN) to the moving labels.
+
+antsApplyTransforms --dimensionality 3 \
+                    --input ${moving_labels} \
+                    --reference-image ${fixed} \
+                    --output ${moving_warped_labels} \
+                    --n NearestNeighbor \
+                    --transform ${prefix}1Warp.nii.gz \
+                    --transform ${prefix}0GenericAffine.mat \
+                    --default-value 0
+\end{lstlisting}
+In this example, we first calculate an initial translation transform by aligning
+the centers of (intensity) mass. The resulting transform is then used as input 
+for determining an optimal rigid transform.  Serial propagation of the resulting
+composite transform continues until all optimal transforms have been determined.
+Optimization for each stage is determined by the specified general parameters 
+including: smoothing and downsampling of fixed and moving images, convergence
+criteria (including number of iterations perl resolution level) and metric (or
+metrics).  Note that any pair of images can be specified per stage.
+
+Although the
+resulting transforms for each stage can be written to disk as output, the 
+default output consists of a condensed set of transforms where compatible 
+transforms have been composed to a single transform.  For example, in the
+above command call, the initial translation, rigid, and affine transforms are
+combined into a single generic affine transform file with the results of the
+deformable transform consisting of discrete vector fields.  The output
+transform files can then be applied using the {\tt antsApplyTransforms}
+program which permits composition of any number of transform files with 
+various interpolation schemes.   For both programs interpolation is never
+performed more than once.
 
@@
 
-We take inspiration from the well-known
-Klein comparative study in which 14 image registration
-algorithms were evaluated based on performance
-on publicly available labeled brain data \citep{klein2009}.
-Based on the final rankings, the SyN algorithm is one
-of the top-performing registration algorithms.  Its relative
-performance coupled with its public availability has made
-it a de facto standard for algorithmic comparison.
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{templates.pdf}
+  \caption{Canonical views for each of the five cohort-specific templates generated using
+  the ANTs tools as described
+  in \cite{avants2010}.   The pseudo-geodesic transform between subjects is created from 
+  the composition of transforms to/from the relevant template.
+  }
+  \label{fig:templates}
+\end{figure}
 
+In the well-known Klein comparative study \citep{klein2009}, 14 image registration
+algorithms were evaluated based on performance on publicly available labeled brain data.
+%Based on the final rankings, the SyN algorithm was determined to be one
+%of the top-performing deformable registration algorithms.  Its relative
+%performance coupled with its public availability has made
+%it a de facto standard for algorithmic comparison.
 For our evaluation, we used the same data as that in the
@@
 }
-We also use the labeled brain data provided at the
+The number of subjects per cohort is provided in the denotation.
+Further details of these labeled brain data (e.g. labeling protocol,
+data sources) are given in \citep{klein2009}.
+We also include the labeled brain data provided at the
 MICCAI 2012 Grand Challenge and Workshop on Multi-Atlas
@@
 https://masi.vuse.vanderbilt.edu/workshop2012
-} which we denote as MICCAI35.  This T1-weighted
-MRI data set consists of 35 subject MRI taken from the
-Oasis database%
+} 
+which we denote as MAL35.  This T1-weighted
+MRI data set consists of 35 subject MRIs taken from the
+Oasis database.%
 \footnote{
 http://www.oasis-brains.org
-}.  The corresponding labels were provided by Neuromorphometrics,
+}  The corresponding labels were provided by Neuromorphometrics,
 Inc.%
@@
 
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\textwidth]{pseudo_geodesic.pdf}
+  \caption{Illustration of generating a pseudo-geodesic for any two subjects
+  within the MGH10 cohort.  Once the transforms between the
+  template and each subject are calculated, the mapping between any two
+  subjects is found by composition of forward and inverse transforms.  For
+  example, in the MGH data set, the pseudo-geodesic transform to
+  map Subject g4 to Subject g7 is found by composing the forward
+  transform from g4 to the template with the inverse transform from the
+  template to g7 (green dashed lines).
+  }
+  \label{fig:pseudo_geodesic}
+\end{figure}
+
+Comparative evaluation of the two SyN registration approaches was performed 
+within each cohort using a ``pseudo-geodesic'' approach.  Instead of
+registering every subject to every other subject within a data set,
+we generated the transforms from each subject to a cohort-specific
+shape/intensity template.  This was accomplished using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
+which is a multivariate implementation of the work described in
+\cite{avants2010}. Canonical views for each of the five templates
+used for this study are given in Figure \ref{fig:templates}.  
+Since calculation of the transform from each subject to the template also
+includes generation of the corresponding inverse transform, the total
+transformation from a given subject to any other is determined from the composition
+of transforms mapping through the template.  An example illustration of the
+geodesic approach is given in Figure \ref{fig:pseudo_geodesic}.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.95\textwidth]{labels.pdf}
+  \caption{Axial views of sample labelings for a member of each data set.  
+  The second row consists of the original labelings with the third row being
+  refined versions of those labelings using the MALF algorithm \cite{wang2012}.
+  These refinements provide more consistency between labelings and improved
+  comparative assessments between algorithms.
+  }
+  \label{fig:labels}
+\end{figure}
+
+Additionally, we refined the labelings for each subject of each cohort
+using the multi-atlas label fusion algorithm (MALF) developed by \cite{wang2012}
+which is also distributed with the ANTs repository.
+For a given subject within a data set, every other subject was mapped to 
+that subject using the pseudo-geodesic transform.  The set of transformed labelings 
+were then used to determine a consensus
+labeling for that subject.
+This was to minimize the obvious {\it observer dimensionality bias} where manual
+raters observe and label in a single dimension at a time.  This
+is most easily seen in the axial or sagittal views of the different cohorts 
+as labelings were done primarily in the coronal view (see Figure \ref{fig:labels}).  
+We include both sets of results.  This provides two sets of labels per subject
+for evaluation.
+
 \section{Results}
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{allCUMC.pdf}
-  \caption{}
+  \begin{tabular}{ccccc}
+  {} &
+  \multicolumn{2}{c}{\bf Original labeling} & 
+  \multicolumn{2}{c}{\bf MALF labeling} \\
+  \rotatebox{90}{\qquad\,\, CUMC12} &
+  \includegraphics[height=38mm]{versusCUMC.png} &
+  \includegraphics[height=38mm]{synComparisonCUMC.pdf} &
+  \includegraphics[height=38mm]{versusCUMCMALF.png} &
+  \includegraphics[height=38mm]{synComparisonCUMCMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad IBSR18} &
+  \includegraphics[height=38mm]{versusIBSR.png} &
+  \includegraphics[height=38mm]{synComparisonIBSR.pdf} &
+  \includegraphics[height=38mm]{versusIBSRMALF.png} &
+  \includegraphics[height=38mm]{synComparisonIBSRMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad LPBA40} &
+  \includegraphics[height=38mm]{versusLPBA.png} &
+  \includegraphics[height=38mm]{synComparisonLPBA.pdf} &
+  \includegraphics[height=38mm]{versusLPBAMALF.png} &
+  \includegraphics[height=38mm]{synComparisonLPBAMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad MGH10} &
+  \includegraphics[height=38mm]{versusMGH.png} &
+  \includegraphics[height=38mm]{synComparisonMGH.pdf} &
+  \includegraphics[height=38mm]{versusMGHMALF.png} &
+  \includegraphics[height=38mm]{synComparisonMGHMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad MAL35} &
+  \includegraphics[height=38mm]{versusMAL.png} &
+  \includegraphics[height=38mm]{synComparisonMAL.pdf} &
+  \includegraphics[height=38mm]{versusMALMALF.png} &
+  \includegraphics[height=38mm]{synComparisonMALMALF.pdf} \\
+  \end{tabular}
+  \caption{Dice results for both algorithms for 
+  each subject warped to every other subject using the pseudo-
+  geodesic transform.  Each 
+  row corresponds to one of the five data sets used for evaluation.
+  For each data set we include a plotting of all individual label 
+  Dice results by volume and a combined label box plotting.  The
+  left and right halves show the results for MALF-derived and 
+  original labelings, respectively.
+  }
+  \label{fig:dice}
 \end{figure}
+
+As mentioned previously, a template was constructed for each data
+set (cf Figure \ref{fig:templates}) from all cohort images.  Subsequently, 
+each image was registered to its corresponding template using either 
+SyN or B-spline SyN as described previously (prior linear registration 
+stages were identical between the two algorithms).  Note that parameter
+exploration is not a part of this work.  Rather, we rely on previously
+reported research \citep{klein2009,avants2011} and our own experience as
+authors/developers of the algorithm/software to select parameters which 
+demonstrate robust performance across data sets.  For both algorithms,
+the gradient step was 0.1 during each of the four multi-resolution levels
+with shrink factors of $\{6,4,2,1\}$ with Gaussian smoothing for each
+of those levels being $\mathcal{N}\left(0,\{9,4,1,0\}\right)$ in terms
+of voxels.  The number of iterations per level were $\{100,100,70,20\}$
+with a convergence threshold of $10^{-9}$ and window size of 15 iterations.
+
+All processing was performed using the computational cluster at the 
+University of Virginia.%
+\footnote{
+http://www.uvacse.virginia.edu
+}
+All perl scripts used to create the jobs for the cluster are
+included in the github account associated with this evaluation.  
+It should be noted that for the
+data sets used in this study, times for B-spline SyN are approximately 
+15-40\% greater than Gaussian-based SyN (single-threaded).
+
+The differences between the two algorithms consist of the Gaussian and 
+B-spline parameters governing the update field smoothing, $S_v$.  In our 
+experience, smoothing of the total field did not improve the results, at
+least for these data (which conforms with our experience extended to other
+data), so the total field smoothing, $S_{\phi}$ is 0 for
+both registration approaches.  Specifically, the chosen parameters
+for the SyN algorithm were:  $S_{\phi} = \mathcal{N}(0,0)$ and 
+$S_{v} = \mathcal{N}(0,3)$ in voxel terms.%
+\footnote{
+ Or, in equivalent {\tt antsRegistration} command line parlance, {\tt -t
+ SyN[0.1,3,0]}.
+ }.  Although our experience with B-spline SyN is much more limited, we 
+were able to choose comparable parameters based on a knot spacing
+for the update field of 26 mm at the base level which is reduced by
+a factor of two for each subsequent multiresolution level.  This 
+yields a final knot spacing of 3.25 mm.%
+\footnote{
+ It should be noted that the current {\tt antsRegistration} command line currently 
+ permits specification of the
+ B-spline total and update field smoothing in terms of total mesh size over
+ the entire image domain, i.e. {\tt -t BSplineSyN[0.1,10x10x8,0]}.  However, 
+ in order to minimize anisotropic smoothing, we determined for each subject 
+ the mesh size closest to producing a knot spacing of 26 mm which was typically
+ around 10 mesh elements per dimension.  We are currently planning on modifying
+ the command line to permit specification of a knot size (similar to the 
+ ANTs-based N4 bias correction algorithm \citep{tustison2010} which also uses the
+ same underlying B-spline computational machinery).
+ }
+As specified in the supplementary material associated with \cite{klein2009}, 
+this is similar to the 
+gradient smoothing parameter for the IRTK FFD algorithm
+which also used four multi-resolution levels with an initial knot spacing of
+20 mm for a final knot spacing of 2.5 mm.
+
+Quality of overlap  using the 
+Dice similarity metric 
+was determined from the transformed labels
+using the open source ITK
+implementation described in \cite{tustison2009a}.
+Both the original labels and MALF labels were warped
+to the fixed image for comparison using nearest neighbor
+interpolation.  A joint Dice metric value was calculated
+from the combined labels for each cohort for each of the
+two SyN methods.  These values are rendered in notched
+box plot format in Figure \ref{fig:dice}.  Non-overlapping notches
+indicate approximately statistically significantly different median values
+at the 95\% confidence level \citep{mcgill1978}.
+For all data sets, the B-spline SyN variant showed a small but statistically
+significant improvement in overall Dice values.
+In order to provide a more complete picture of performance differences,
+we also accounted for label volumetric considerations \cite{rohlfing2012}.  In Figure
+\ref{fig:dice} we plotted the Dice value difference between each
+SyN variant (B-spline SyN - SyN) for each label of each intra-subject
+registration pair within a data set versus the volume of the label
+in the fixed image.
 
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{allIBSR.pdf}
-  \caption{Currently redoing on the cluster.}
+    \includegraphics[width=0.95\textwidth]{jacobianRange.pdf}
+  \caption{Violin plots of the range of log Jacobian values 
+           (95th\% - the 5th\%)  
+           for all deformable
+           transforms from each subject to its corresponding
+           template.  B-spline SyN demonstrates a tendency 
+           to produce a much
+           greater range of log Jacobian values.
+           }
+  \label{fig:jacobian}
 \end{figure}
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.95\textwidth]{allLPBA.pdf}
-  \caption{}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.95\textwidth]{allMGH.pdf}
-  \caption{Currently redoing on the cluster.}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.95\textwidth]{allMICCAI.pdf}
-  \caption{}
-\end{figure}
+To further characterize the deformable transform differences, we 
+calculated the log of the Jacobian of the transformations from 
+each subject to the template and 
+tabulated statistical information within the brain region only. 
+A noticeable difference between the two algorithms was the
+respective range of values in log Jacobians.  We plotted the 
+(95th\% - the 5th\%) for each algorithm across all data sets
+ in Figure \ref{fig:jacobian}.  It is apparent 
+that B-spline SyN produces a much greater range of log Jacobian
+values during the course of optimization.
 
@@
 
-B-spline regularization is easily adapted into the diffeomorphic registration framework and performs comparably to analogous algorithms which we demonstrated in the case of SyN.
-Even more importantly, all source code and data used in this work is publicly available for the interested researcher to explore.
+A significant amount of research has been devoted to image registration
+algorithmic development.  Given their many salient characteristics particularly with
+respect to large deformation estimation constrained by topological 
+continuity, diffeomorphic registration
+approaches have been a particular focus in the neuroimaging community.
+However, many groups continue to find success with non-diffeomorphic
+FFD methods
+\citep[e.g.][]{rueckert1999,klein2010}.  Using our DMFFD framework, 
+B-spline regularization is easily adapted into the diffeomorphic registration 
+framework and performs comparably to analogous algorithms which we demonstrated 
+in this work for the case of the widely-used SyN.
+
+B-spline SyN produced slightly greater Dice values than the original SyN.
+Interestingly it was shown that the range of log Jacobian values tend to be
+significantly higher for the former than the latter.  Although possible 
+explanations include the approximation-versus-convolution distinction mentioned
+earlier, further exploration is planned.
+
+One of the advantages that has not been explored in this manuscript is 
+the use of B-spline SyN for small deformation estimation problems such 
+as in pulmonary or cardiac applications.  Such problems typically require
+greater regularization which implies larger discrete Gaussian kernels.
+A related issue is applications involving severely anisotropic data where
+the continuous nature of the DMFFD approach might help over Gaussian convolution.
+Ongoing work will continue to explore these issues.
+
+% What about sparsely sampled data and the advantages of b-splines
+
+Perhaps most important, in the spirt of open science, all source code
+used in this work is publicly available and well-vetted as part of the 
+ANTs initiative.
 

### 2013-10-19 Results/jacobianMeasures.R

@@ -1,0 +1,104 @@
+library( ggplot2 )
+library( reshape )
+
+stats <- list()
+
+# cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MAL' )
+cohortNames <- c( 'CUMC12', 'IBSR18', 'LPBA40', 'MGH10', 'MAL35' )
+
+stats[[1]] <- read.csv( "./logJacobianMeasuresCUMCTemplate.csv" )
+stats[[2]] <- read.csv( "./logJacobianMeasuresIBSRTemplate.csv" )
+stats[[3]] <- read.csv( "./logJacobianMeasuresLPBATemplate.csv" )
+stats[[4]] <- read.csv( "./logJacobianMeasuresMGHTemplate.csv" )
+stats[[5]] <- read.csv( "./logJacobianMeasuresMALTemplate.csv" )
+
+jacobianRange <- c()
+regularizationCohort <- c()
+factorLevels <- c()
+
+for( i in 1:length( cohortNames ) )
+  {
+  jacSyN5th <- stats[[i]]$X5th.[stats[[i]]$Regularization == 'syn']
+  jacBSplineSyN5th <- stats[[i]]$X5th.[stats[[i]]$Regularization == 'bsyn']
+
+  jacSyN95th <- stats[[i]]$X95th.[stats[[i]]$Regularization == 'syn']
+  jacBSplineSyN95th <- stats[[i]]$X95th.[stats[[i]]$Regularization == 'bsyn']
+
+  jacSyNRange <- jacSyN95th - jacSyN5th
+  jacBSplineSyNRange <- jacBSplineSyN95th - jacBSplineSyN5th
+
+  regCohort <- c( rep.int( paste0( "SyN\n(", cohortNames[i], ")" ), length( jacSyNRange ) ),
+                  rep.int( paste0( "BSplineSyN\n(", cohortNames[i], ")" ), length( jacBSplineSyNRange ) ) )
+
+  factorLevels <- append( factorLevels, c( paste0( "BSplineSyN\n(", cohortNames[i], ")" ), paste0( "SyN\n(", cohortNames[i], ")" ) ) )
+
+  regularizationCohort <- append( regularizationCohort, regCohort )
+  jacobianRange <- append( jacobianRange, c( jacSyNRange, jacBSplineSyNRange ) )
+  }
+
[…]

### 2013-10-19 Results/overlapMeasures.R

@@ -1,20 +1,34 @@
 library( ggplot2 )
+library( reshape )
 
 stats <- list()
 volumes <- list()
 
-cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MICCAI' )
 
-stats[[1]] <- read.csv( "OverlapMeasures/overlapMeasuresCUMC.csv" )
-volumes[[1]] <- read.csv( "OverlapMeasures/labelVolumesCUMC.csv" )
-stats[[2]] <- read.csv( "OverlapMeasures/overlapMeasuresIBSR.csv" )
-volumes[[2]] <- read.csv( "OverlapMeasures/labelVolumesIBSR.csv" )
-stats[[3]] <- read.csv( "OverlapMeasures/overlapMeasuresLPBA.csv" )
-volumes[[3]] <- read.csv( "OverlapMeasures/labelVolumesLPBA.csv" )
-stats[[4]] <- read.csv( "OverlapMeasures/overlapMeasuresMGH.csv" )
-volumes[[4]] <- read.csv( "OverlapMeasures/labelVolumesMGH.csv" )
-stats[[5]] <- read.csv( "OverlapMeasures/overlapMeasuresMICCAI.csv" )
-volumes[[5]] <- read.csv( "OverlapMeasures/labelVolumesMICCAI.csv" )
+
+cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MAL' )
+
+stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplateMALF.csv" )
+volumes[[1]] <- read.csv( "./labelVolumesCUMCTemplateMALF.csv" )
+stats[[2]] <- read.csv( "./overlapMeasuresIBSRTemplateMALF.csv" )
+volumes[[2]] <- read.csv( "./labelVolumesIBSRTemplateMALF.csv" )
+stats[[3]] <- read.csv( "./overlapMeasuresLPBATemplateMALF.csv" )
+volumes[[3]] <- read.csv( "./labelVolumesLPBATemplateMALF.csv" )
+stats[[4]] <- read.csv( "./overlapMeasuresMGHTemplateMALF.csv" )
+volumes[[4]] <- read.csv( "./labelVolumesMGHTemplateMALF.csv" )
+stats[[5]] <- read.csv( "./overlapMeasuresMALTemplateMALF.csv" )
[…]

### 2013-10-19 paper.tex
ENH:  adding some mnor edits.

@@
 topological properties and success in providing biologically plausible
-solutions to small and large deformation
+solutions to deformation and morphological
 estimation problems. Popular variants of diffeomorphic algorithms include those characterized by
@@
 Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework, the traditional free-form deformation method can be viewed as a variant in which explicit regularization is achieved through the B-spline basis functions.
+In the context of the original Demons' framework, a free-form deformation methodological variant, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through
+fast B-spline appromixation.
 This characterization can be used to provide alternative B-spline ``flavored'' diffeomorphic image registration solutions with several advantages which we describe in this work.
@@
 algorithmic development which continues contemporaneously.
-We highlight several key contributions which are particularly relevant to the work presented.
+We highlight several key historical contributions which are particularly relevant to the work presented.
 
@@
 smoothing via Gaussian convolution, a defining characteristic of Demons,
-of the update or displacement
+of the update or total displacement
 field is a greedy approximation for solving the partial differential equations governing
@@
 This analogy was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
-This preconditioning is used to mitigate the hemstitching effect induced by the ill-conditioned nature of the traditional gradient-based FFD formulation.
-Further details can be gleaned from the ITK implementation of this
-earlier work%
+This preconditioning is used to mitigate the hemstitching effect induced by the ill-conditioned nature of the traditional gradient-based FFD formulation.  We denote
+this new FFD variant as directly manipulated free-form deformation (DMFFD) and, 
+as part of the ITKv4 refactoring efforts, has been implemented for use with the new
+registration framework%
 \footnote{
@@
 the log-Euclidean framework of \cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using Exponential Lie algebra) uses a constant velocity field parameterization to provide a fast, diffeomorphic alternative \citep{ashburner2007}.
-Additionally, symmetrical considerations in the velocity field parameterization are discussed in \cite{avants2008} in the context of a cross correlation similarity metric which minimizes the bias of the resulting transformation when selecting the ``fixed'' and ``moving'' images.
+Additionally, symmetrical considerations in the velocity field parameterization are discussed in \cite{avants2008} in the context of a cross correlation similarity metric.
+By explicitly symmetrizing the LDDMM formulation, this Symmetric Normalization (SyN) minimizes the bias of the resulting transformation when selecting the ``fixed'' and ``moving'' images.
 A greedy version of this algorithm has proven successful in neuroimaging \citep{klein2009} and pulmonary \citep{murphy2011} applications.
@@
 a full LDDMM-style algorithm based on B-splines called
-{\em temporal free-form deformation} in which the
+{\it temporal free-form deformation} in which the
 time-varying velocity field
-is modeled using a 4-D B-spline object (3-D + time).  Numerical integration of the mapping propagated within the velocity field yields the transform between parameterized time points.
+is modeled using a 4-D B-spline object (3-D + time).  Numerical Eulerian integration of the mapping propagated within the velocity field yields the transform between parameterized time points.
 
@@
 for regularizing time-varying vector fields in an analogous fashion as
-Gaussian convolution.  In this vein and similar to
+Gaussian convolution.  In this vein, and similar to
 \cite{de-craene2011}, we reported in \cite{tustison2012a,tustison2012}
@@
 entire gamut of diffeomorphic registration algorithms and provides
-unique ``flavors'' of smoothing possibilities.  However, differences
-between the two approaches produce characteristically different
+unique ``flavors'' of smoothing possibilities.  However, 
+the two approaches produce characteristically different
 solutions.  In addition to
@@
 approach.  Also, whereas
-convolution operates entirely within the discrete space,
-the fitting process constructs a continuous object prior
-to any discrete sampling.  
+Gaussian convolution operates entirely within discretized space,
+the B-spline approximation routine constructs a continuous object prior
+to any voxelwise reconstruction of the sampled fields.  
 Interestingly, a similar comparison was made with respect to 
@@
 }
-Given the popularity and excellent performance of the symmetric normalization (SyN) algorithm, our focus in this work is
-its B-spline equivalent which we denote as ``B-spline SyN.''
+Given the popularity and excellent performance of the greedy variant of SyN algorithm, our evaluation focus in this work is
+its B-spline analog which we denote as ``B-spline SyN'' or ``DMFFD SyN.''
 Evaluation and comparisons of the respective algorithmic instantiations
@@
 where $v$ is a time-dependent smooth field (as dictated by the functional norm $L$), $v : \Omega \times t
-\rightarrow \mathrm{R}^d$ parameterized by $t \in [0,1]$.  Diffeomorphic mappings
+\rightarrow \mathrm{R}^d$.  Diffeomorphic mappings
 between parameterized time points $\{t_a,t_b\} \in [0,1]$
@@
   \inf_{\phi_1} \inf_{\phi_2} \left(
-                     \int_0^{0.5} \left( \|v_1(t)\|_L^2 + \|v_2(1-t)\|_L^2 \right) dt +
+                     \int_0^{0.5} \left( \|v_1(t)\|_L^2 + \|v_2(t)\|_L^2 \right) dt +
                      \int_{\Omega} | I \circ \phi_1^{-1}(\mathbf{x},0.5)
@@
 while simultaneously being capable of running with limited computational
-resources.  In this variant, the discrete time-parameterized
+resources.  This is possible since the discrete time-parameterized
 velocity field samples are restricted to their respective
@@
 default to Gaussian smoothing due to its simplicity both in implementation
-and complexity terms.  A viable and practical alternative is the DMFFD 
+and complexity terms.  As mentioned earlier, 
+a viable and practical alternative is the DMFFD 
 approach based on B-splines for explicit regularization of vector fields.
@@
 characterizing the velocity field.  The preconditioned gradient analog of
-Equation (\ref{eq:dmffd}) for updating the time-varying velocity field control point lattice
-is as follows:
-$\delta v_{i_1,\ldots,i_d,i_t}$, given the similarity metric, $\Pi_\sim$,
+Equation (\ref{eq:dmffd}) for updating the time-varying velocity field control point lattice is 
 \begin{align}
@@
 
-The class {\tt itk::BSplineScatteredDataPointSetToImageFilter} underlies all DMFFD regularization which implements the 
-methods described in \cite{tustison2009}.  Although applicable to
+The class {\tt itk::BSplineScatteredDataPointSetToImageFilter} underlies all DMFFD regularization which is an implementation of the 
+methods described in \cite{tustison2006}.  Although applicable to
 various scenarios (e.g. curve and surface estimation), it has
 been optimized for imaging applications and multi-threaded
-for fast processing on multi-core machines.
+for fast processing on suitable machines.
 Additionally, numerical integration for solving Equations (\ref{eq:ode}) and (\ref{eq:integral}) utilizes Runge-Kutta which
@@
 \footnote{
-The long command and short command line help menus can be invoked by
-the commands `{\tt antsRegistrationantsRegistration --help}' and `{\tt antsRegistration -h}', respectively.
+The long and short command line help menus can be invoked by
+the commands `{\tt antsRegistration --help}' and `{\tt antsRegistration -h}', respectively.
 }
@@
 including: smoothing and downsampling of fixed and moving images, convergence
-criteria (including number of iterations perl resolution level) and metric (or
+criteria (including number of iterations per resolution level) and metric (or
 metrics).  Note that any pair of images can be specified per stage.
@@
 %it a de facto standard for algorithmic comparison.
-For our evaluation, we used the same data as that in the
-Klein study.  Specifically, we used the data sets denoted as:
+For our evaluation, we used these same data.  
+Specifically, we used the data sets denoted as:
 \begin{itemize}
@@
 SyN or B-spline SyN as described previously (prior linear registration 
-stages were identical between the two algorithms).  Note that parameter
-exploration is not a part of this work.  Rather, we rely on previously
+stages were identical between the two algorithms).  As a brute-force parameter
+exploration is not a part of this work, we rely on previously
 reported research \citep{klein2009,avants2011} and our own experience as
@@
 It should be noted that for the
-data sets used in this study, times for B-spline SyN are approximately 
-15-40\% greater than Gaussian-based SyN (single-threaded).
+data sets used in this study, times for B-spline SyN were approximately 
+15-40\% greater than Gaussian-based SyN using single-threading and a dense
+metric gradient sampling.
 
@@
 experience, smoothing of the total field did not improve the results, at
-least for these data (which conforms with our experience extended to other
+least for these data (which conforms with our experience with other
 data), so the total field smoothing, $S_{\phi}$ is 0 for
@@
  SyN[0.1,3,0]}.
- }.  Although our experience with B-spline SyN is much more limited, we 
+ }  Although our experience with B-spline SyN is much more limited, we 
 were able to choose comparable parameters based on a knot spacing
@@
 \footnote{
- It should be noted that the current {\tt antsRegistration} command line currently 
+ The {\tt antsRegistration} command line currently 
  permits specification of the
- B-spline total and update field smoothing in terms of total mesh size over
+ B-spline total and update field smoothing only in terms of total mesh size over
  the entire image domain, i.e. {\tt -t BSplineSyN[0.1,10x10x8,0]}.  However, 
@@
 which also used four multi-resolution levels with an initial knot spacing of
-20 mm for a final knot spacing of 2.5 mm.
+20 mm per dimension for a final knot spacing of 2.5 mm.
 
@@
 In order to provide a more complete picture of performance differences,
-we also accounted for label volumetric considerations \cite{rohlfing2012}.  In Figure
+we also accounted for label volumetric considerations \citep{rohlfing2012}.  In Figure
 \ref{fig:dice} we plotted the Dice value difference between each
@@
 To further characterize the deformable transform differences, we 
-calculated the log of the Jacobian of the transformations from 
+calculated the log of the Jacobian determinant of the transformations from 
 each subject to the template and 
@@
  in Figure \ref{fig:jacobian}.  It is apparent 
-that B-spline SyN produces a much greater range of log Jacobian
+that B-spline SyN produces a much greater range of deformation
 values during the course of optimization.
@@
 the continuous nature of the DMFFD approach might help over Gaussian convolution.
+Additionally, we have not yet explored the use of sparsely sampled data 
+and the advantages of the proposed framework.
 Ongoing work will continue to explore these issues.
 
-% What about sparsely sampled data and the advantages of b-splines
-
-Perhaps most important, in the spirt of open science, all source code
-used in this work is publicly available and well-vetted as part of the 
-ANTs initiative.
+%Perhaps most important, in the spirt of open science, all source code
+%used in this work is publicly available and well-vetted as part of the 
+%ANTs initiative.
 

### 2013-10-25 paper.tex
ENH:  More edits.

@@
                  --shrink-factors 4x3x2 \
-                 --transform BSplineSyN[0.1,10x10x10,0x0x0,3] \
+                 --transform BSplineSyN[0.1,26,0x0x0,3] \
                  --metric CC[${fixed},${moving},1,4] \ 
@@
 labeling for that subject.
-This was to minimize the obvious {\it observer dimensionality bias} where manual
+This was to minimize the obvious {\it observer dimensionality artifacts} where manual
 raters observe and label in a single dimension at a time.  This
@@
 \footnote{
- The {\tt antsRegistration} command line currently 
- permits specification of the
- B-spline total and update field smoothing only in terms of total mesh size over
- the entire image domain, i.e. {\tt -t BSplineSyN[0.1,10x10x8,0]}.  However, 
- in order to minimize anisotropic smoothing, we determined for each subject 
- the mesh size closest to producing a knot spacing of 26 mm which was typically
- around 10 mesh elements per dimension.  We are currently planning on modifying
- the command line to permit specification of a knot size (similar to the 
- ANTs-based N4 bias correction algorithm \citep{tustison2010} which also uses the
- same underlying B-spline computational machinery).
+ Or, equivalently, {\tt -t
+ BSplineSyN[0.1,26,0]}.
+% The {\tt antsRegistration} command line currently 
+% permits specification of the
+% B-spline total and update field smoothing only in terms of total mesh size over
+% the entire image domain, i.e. {\tt -t BSplineSyN[0.1,10x10x8,0]}.  However, 
+% in order to minimize anisotropic smoothing, we determined for each subject 
+% the mesh size closest to producing a knot spacing of 26 mm which was typically
+% around 10 mesh elements per dimension.  We are currently planning on modifying
+% the command line to permit specification of a knot size (similar to the 
+% ANTs-based N4 bias correction algorithm \citep{tustison2010} which also uses the
+% same underlying B-spline computational machinery).
  }

### 2013-10-25 paper.tex
ENH: Keywords.

@@
 
-\title[DMFFD Diffeomorphic Image Registration]{Explicit B-spline regularization in diffeomorphic image registration}
+\title[DMFFD diffeomorphic image registration]{Explicit B-spline regularization in diffeomorphic image registration}
 \author[\firstAuthorLast ]{\Authors}
@@
 \keyFont{ \section{Keywords:}
-open science, best practices, reproducibility, software }
+Advanced Normalization Tools, diffeomorphisms, directly manipulated free-form deformation, Insight Toolkit, spatial normalization }
 \end{abstract}

### 2013-10-25 paper.tex
ENH: Getting stuff ready for submission.

@@
   \centering
-  \includegraphics[width=0.95\textwidth]{SyN.pdf}
+  \includegraphics[width=0.95\textwidth]{SyN.jpg}
   \caption{Illustration of the greedy SyN formulation.  Given images
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{templates.pdf}
+  \includegraphics[width=0.95\textwidth]{templates.jpg}
   \caption{Canonical views for each of the five cohort-specific templates generated using
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{pseudo_geodesic.pdf}
+  \includegraphics[width=0.75\textwidth]{pseudo_geodesic.jpg}
   \caption{Illustration of generating a pseudo-geodesic for any two subjects
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{labels.pdf}
+  \includegraphics[width=0.95\textwidth]{labels.jpg}
   \caption{Axial views of sample labelings for a member of each data set.  
@@
   \centering
-  \begin{tabular}{ccccc}
-  {} &
-  \multicolumn{2}{c}{\bf Original labeling} & 
-  \multicolumn{2}{c}{\bf MALF labeling} \\
-  \rotatebox{90}{\qquad\,\, CUMC12} &
-  \includegraphics[height=38mm]{versusCUMC.png} &
-  \includegraphics[height=38mm]{synComparisonCUMC.pdf} &
-  \includegraphics[height=38mm]{versusCUMCMALF.png} &
-  \includegraphics[height=38mm]{synComparisonCUMCMALF.pdf} \\
-  \rotatebox{90}{\qquad\quad IBSR18} &
-  \includegraphics[height=38mm]{versusIBSR.png} &
-  \includegraphics[height=38mm]{synComparisonIBSR.pdf} &
-  \includegraphics[height=38mm]{versusIBSRMALF.png} &
-  \includegraphics[height=38mm]{synComparisonIBSRMALF.pdf} \\
-  \rotatebox{90}{\qquad\quad LPBA40} &
-  \includegraphics[height=38mm]{versusLPBA.png} &
-  \includegraphics[height=38mm]{synComparisonLPBA.pdf} &
-  \includegraphics[height=38mm]{versusLPBAMALF.png} &
-  \includegraphics[height=38mm]{synComparisonLPBAMALF.pdf} \\
-  \rotatebox{90}{\qquad\quad MGH10} &
-  \includegraphics[height=38mm]{versusMGH.png} &
-  \includegraphics[height=38mm]{synComparisonMGH.pdf} &
-  \includegraphics[height=38mm]{versusMGHMALF.png} &
-  \includegraphics[height=38mm]{synComparisonMGHMALF.pdf} \\
-  \rotatebox{90}{\qquad\quad MAL35} &
-  \includegraphics[height=38mm]{versusMAL.png} &
-  \includegraphics[height=38mm]{synComparisonMAL.pdf} &
-  \includegraphics[height=38mm]{versusMALMALF.png} &
-  \includegraphics[height=38mm]{synComparisonMALMALF.pdf} \\
+  \begin{tabular}{c}
+  \includegraphics[width=0.95\textwidth]{evaluation.jpg}
   \end{tabular}
-  \caption{Dice results for both algorithms for 
-  each subject warped to every other subject using the pseudo-
-  geodesic transform.  Each 
-  row corresponds to one of the five data sets used for evaluation.
-  For each data set we include a plotting of all individual label 
-  Dice results by volume and a combined label box plotting.  The
-  left and right halves show the results for MALF-derived and 
-  original labelings, respectively.
-  }
-  \label{fig:dice}
 \end{figure}
+
+%\begin{figure}[htb]
+%  \centering
+%  \begin{tabular}{ccccc}
+%  {} &
+%  \multicolumn{2}{c}{\bf Original labeling} & 
+%  \multicolumn{2}{c}{\bf MALF labeling} \\
+%  \rotatebox{90}{\qquad\,\, CUMC12} &
+%  \includegraphics[height=38mm]{versusCUMC.png} &
+%  \includegraphics[height=38mm]{synComparisonCUMC.pdf} &
+%  \includegraphics[height=38mm]{versusCUMCMALF.png} &
+%  \includegraphics[height=38mm]{synComparisonCUMCMALF.pdf} \\
+%  \rotatebox{90}{\qquad\quad IBSR18} &
+%  \includegraphics[height=38mm]{versusIBSR.png} &
+%  \includegraphics[height=38mm]{synComparisonIBSR.pdf} &
+%  \includegraphics[height=38mm]{versusIBSRMALF.png} &
+%  \includegraphics[height=38mm]{synComparisonIBSRMALF.pdf} \\
+%  \rotatebox{90}{\qquad\quad LPBA40} &
+%  \includegraphics[height=38mm]{versusLPBA.png} &
+%  \includegraphics[height=38mm]{synComparisonLPBA.pdf} &
+%  \includegraphics[height=38mm]{versusLPBAMALF.png} &
+%  \includegraphics[height=38mm]{synComparisonLPBAMALF.pdf} \\
+%  \rotatebox{90}{\qquad\quad MGH10} &
+%  \includegraphics[height=38mm]{versusMGH.png} &
+%  \includegraphics[height=38mm]{synComparisonMGH.pdf} &
+%  \includegraphics[height=38mm]{versusMGHMALF.png} &
+%  \includegraphics[height=38mm]{synComparisonMGHMALF.pdf} \\
+%  \rotatebox{90}{\qquad\quad MAL35} &
+%  \includegraphics[height=38mm]{versusMAL.png} &
+%  \includegraphics[height=38mm]{synComparisonMAL.pdf} &
+%  \includegraphics[height=38mm]{versusMALMALF.png} &
+%  \includegraphics[height=38mm]{synComparisonMALMALF.pdf} \\
+%  \end{tabular}
+%  \caption{Dice results for both algorithms for 
+%  each subject warped to every other subject using the pseudo-
+%  geodesic transform.  Each 
+%  row corresponds to one of the five data sets used for evaluation.
+%  For each data set we include a plotting of all individual label 
+%  Dice results by volume and a combined label box plotting.  The
+%  left and right halves show the results for MALF-derived and 
+%  original labelings, respectively.
+%  }
+%  \label{fig:dice}
+%\end{figure}
 
@@
   \centering
-    \includegraphics[width=0.95\textwidth]{jacobianRange.pdf}
+    \includegraphics[width=0.95\textwidth]{jacobianRange.jpg}
   \caption{Violin plots of the range of log Jacobian values 

### 2013-10-25 paper.tex
ENH: Getting stuff ready for submission.

@@
   \end{tabular}
+  \caption{Dice results for both algorithms for 
+  each subject warped to every other subject using the pseudo-
+  geodesic transform.  Each 
+  row corresponds to one of the five data sets used for evaluation.
+  For each data set we include a plotting of all individual label 
+  Dice results by volume and a combined label box plotting.  The
+  left and right halves show the results for MALF-derived and 
+  original labelings, respectively.
+  }
+  \label{fig:dice}
 \end{figure}

### 2013-10-25 paper.tex
ENH:  edits.

@@
 time-varying and constant velocity fields, and symmetrical considerations.
-Prior information (i.e.
-regularization) is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
+Prior information in the form of regularization is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
 Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework, a free-form deformation methodological variant, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through
+In the context of the original Demons' framework, a free-form deformation methodological alternative, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through
 fast B-spline appromixation.
@@
 Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
-A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open data and open source evaluation tools.
+A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open labeled brain data and open source evaluation tools.
 \tiny

### 2013-10-25 paper.tex
ENH:  edits.

@@
 Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework, a free-form deformation methodological alternative, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through
-fast B-spline appromixation.
+In the context of the original Demons' framework the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through fast B-spline approximation.
 This characterization can be used to provide alternative B-spline ``flavored'' diffeomorphic image registration solutions with several advantages which we describe in this work.

### 2013-10-26 paper.tex
ENH:  BYU vs. Boise State.

@@
 Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through fast B-spline approximation.
-This characterization can be used to provide alternative B-spline ``flavored'' diffeomorphic image registration solutions with several advantages which we describe in this work.
+In the context of the original Demons' framework, the {\it directly manipulated free-form deformation} approach \citep{tustison2009}, can be viewed as a smoothing alternative in which explicit regularization is achieved through fast B-spline approximation.
+This characterization can be used to provide B-spline ``flavored'' diffeomorphic image registration solutions with several advantages.
 Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
@@
 algorithmic development which continues contemporaneously.
-We highlight several key historical contributions which are particularly relevant to the work presented.
+We highlight several key historical contributions which are particularly relevant to the work 
+subsequently presented.
 
@@
 contributions as \cite{szeliski1997,thevenaz1998,rueckert1999}.
-Continued development within this early spline-based paradigm produced additional innovations such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation constraints \citep[e.g.][]{rohlfing2003}, and notable open source implementations \citep[e.g.][]{ibanez2005,klein2010,shackleford2010}.
+Continued development within this early spline-based paradigm produced additional innovations such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation constraints \citep[e.g.][]{rohlfing2003}, and notable open source implementations \citep[e.g.][]{ibanez2005,klein2010,modat2010,shackleford2010}.
 
@@
 by \cite{christensen1996}, LDDMM yields the geodesic solution in the space of diffeomorphisms between two images. Since its introduction, LDDMM has
-inspired much innovation in the image registration literature.  Applying
+inspired much innovation in the image registration community.  Applying
 the log-Euclidean framework of \cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using Exponential Lie algebra) uses a constant velocity field parameterization to provide a fast, diffeomorphic alternative \citep{ashburner2007}.
@@
 lead to different solutions despite the fact that the theoretical
-geodesic solution describes the same path forwards and backwards.
+geodesic solution describes the same parameterized path forwards and backwards.
 This observation led to the symmetric
@@
 \begin{align}
-  \frac{d \phi_i(\mathbf{x},t)}{dt} = v_i( \phi_i(\mathbf{x},t), t ),\,\, \phi_i(\mathbf{x},0) = \mathbf{Id}, \,\, i \in \{1,2\}
+  \frac{d \phi_i(\mathbf{x},t)}{dt} = v_i( \phi_i(\mathbf{x},t), t ),\,\, \phi_i(\mathbf{x},0) = \mathbf{Id}, \,\, i \in \{1,2\}.
 \end{align}
-with extension to arbitrary similarity metric choice, i.e. the second
+With extension to arbitrary similarity metric choice, the second
 term is replaced with
@@
 and storage of 4-D velocity fields at each iteration which is limiting for
-many common, contemporary processing infrastructures.
+many common computational platforms.
 
@@
 while simultaneously being capable of running with limited computational
-resources.  This is possible since the discrete time-parameterized
-velocity field samples are restricted to their respective
+resources.  This is due to the restriction of the discrete, time-parameterized
+velocity field samples to their respective
 endpoints, i.e. the $v_i(\mathbf{x},t)$ are sampled at $t \in \{0,0.5,1\}$
@@
 \footnote{
-In terms of implementation spline orders can be specified separately for each dimension but, for simplicity,
+In terms of implementation, spline orders can be specified separately for each dimension but, for simplicity,
 we only specify a single spline order.
@@
 and $\frac{\partial \Pi_\sim}{\partial \mathbf{x}}$ is the spatial 
-similarity metric gradient indexed by $c$.
+similarity metric gradient at voxel $c$.
 
@@
 In this example, we first calculate an initial translation transform by aligning
-the centers of (intensity) mass. The resulting transform is then used as input 
+the centers of (intensity) mass (although alignment based on other features is possible). 
+The resulting transform is then used as input 
 for determining an optimal rigid transform.  Serial propagation of the resulting
@@
 criteria (including number of iterations per resolution level) and metric (or
-metrics).  Note that any pair of images can be specified per stage.
+metrics).  Note that any pair of images can be specified per metric per stage.
 
@@
 program which permits composition of any number of transform files with 
-various interpolation schemes.   For both programs interpolation is never
+different interpolation schemes.   For both programs interpolation is never
 performed more than once.
@@
 we generated the transforms from each subject to a cohort-specific
-shape/intensity template.  This was accomplished using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
+shape/intensity template.  These templates were built using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
 which is a multivariate implementation of the work described in
@@
 using the multi-atlas label fusion algorithm (MALF) developed by \cite{wang2012}
-which is also distributed with the ANTs repository.
+which is also distributed with ANTs.
 For a given subject within a data set, every other subject was mapped to 
@@
   Dice results by volume and a combined label box plotting.  The
-  left and right halves show the results for MALF-derived and 
-  original labelings, respectively.
+  left and right halves show the respective results for 
+  original and MALF-derived labelings.
   }
@@
 demonstrate robust performance across data sets.  For both algorithms,
-the gradient step was 0.1 during each of the four multi-resolution levels
-with shrink factors of $\{6,4,2,1\}$ with Gaussian smoothing for each
+the gradient step was 0.1 for each of the four multi-resolution levels
+with shrink factors of $\{6,4,2,1\}$ and Gaussian smoothing for each
 of those levels being $\mathcal{N}\left(0,\{9,4,1,0\}\right)$ in terms
@@
 }
-All perl scripts used to create the jobs for the cluster are
+The perl scripts used to create the jobs for the cluster are
 included in the github account associated with this evaluation.  
@@
 for the update field of 26 mm at the base level which is reduced by
-a factor of two for each subsequent multiresolution level.  This 
+a factor of two for each succeeding multiresolution level.  This 
 yields a final knot spacing of 3.25 mm.%
@@
  }
-As specified in the supplementary material associated with \cite{klein2009}, 
-this is similar to the 
+For comparison, after selecting these smoothing parameters we discovered 
+in the supplementary material of \cite{klein2009} the similarity to the 
 gradient smoothing parameter for the IRTK FFD algorithm
@@
 registration pair within a data set versus the volume of the label
-in the fixed image.
+in the fixed image.  Values above and below the dashed line indicate better 
+regional performance for B-spline SyN and SyN, respectively.
 
@@
 respective range of values in log Jacobians.  We plotted the 
-(95th\% - the 5th\%) for each algorithm across all data sets
+(95th\% - 5th\%) for each algorithm across all data sets
  in Figure \ref{fig:jacobian}.  It is apparent 
-that B-spline SyN produces a much greater range of deformation
-values during the course of optimization.
+that B-spline SyN results exhibit a much greater range of deformation.
 
@@
 Avoid extensive citations and discussion of published literature.
+
+B-spline SyN produced slightly greater Dice values than the original SyN.  Although
+actual differences are relatively small, they are statistically significant. By 
+implementing both algorithms in the the same code base, we are not only able to 
+eliminate non-regularization components of the registration but we are also able
+to eliminate implementation differences.  Thus, performance disparity can be isolated
+to smoothing choice.  However, even with this restricted focus there are various
+reasons for the evaluation outcome.  These include the approximation-versus-convolution 
+distinction mentioned earlier for the two regularization approaches (which could
+also explain the reason why the range of log Jacobian values tend to be
+significantly higher for B-spline SyN).  Also, the fact that the regularization 
+for B-spline SyN is theoretically continuous whereas the truncated Gaussian 
+convolution is only a discrete approximation could be a potential factor.  Exploration
+of other possibilities are planned for future work.
+
+One of the advantages that has not been explored in this manuscript is 
+the use of B-spline SyN for small deformation estimation problems such 
+as in pulmonary or cardiac applications.  Such problems typically require
+greater regularization which implies larger discrete kernels for 
+Gaussian convolution.
+A related issue is applications involving severely anisotropic data where
+the continuous nature of the DMFFD approach might help over Gaussian convolution.
+Additionally, we have not yet explored the use of sparsely sampled data 
+and the advantages of the proposed framework.
+Ongoing work will continue to explore these issues.
 
@@
 B-spline regularization is easily adapted into the diffeomorphic registration 
-framework and performs comparably to analogous algorithms which we demonstrated 
+framework and performs well compared to analogous algorithms which we demonstrated 
 in this work for the case of the widely-used SyN.
-
-B-spline SyN produced slightly greater Dice values than the original SyN.
-Interestingly it was shown that the range of log Jacobian values tend to be
-significantly higher for the former than the latter.  Although possible 
-explanations include the approximation-versus-convolution distinction mentioned
-earlier, further exploration is planned.
-
-One of the advantages that has not been explored in this manuscript is 
-the use of B-spline SyN for small deformation estimation problems such 
-as in pulmonary or cardiac applications.  Such problems typically require
-greater regularization which implies larger discrete Gaussian kernels.
-A related issue is applications involving severely anisotropic data where
-the continuous nature of the DMFFD approach might help over Gaussian convolution.
-Additionally, we have not yet explored the use of sparsely sampled data 
-and the advantages of the proposed framework.
-Ongoing work will continue to explore these issues.
 

### 2013-10-26 paper.tex
ENH:  Jambi

@@
                  --shrink-factors 4x3x2 \
-                 --transform BSplineSyN[0.1,26,0x0x0,3] \
+                 --transform BSplineSyN[0.1,26,0,3] \
                  --metric CC[${fixed},${moving},1,4] \ 

### 2013-10-28 paper.tex
ENH:  Walking dead.

@@
   \caption{Violin plots of the range of log Jacobian values 
-           (95th\% - the 5th\%)  
+           (95th\% - 5th\%)  
            for all deformable

### 2013-10-28 Results/overlapMeasures.R

@@ -8,27 +8,27 @@ volumes <- list()
 
 cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MAL' )
 
-stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplateMALF.csv" )
-volumes[[1]] <- read.csv( "./labelVolumesCUMCTemplateMALF.csv" )
-stats[[2]] <- read.csv( "./overlapMeasuresIBSRTemplateMALF.csv" )
-volumes[[2]] <- read.csv( "./labelVolumesIBSRTemplateMALF.csv" )
-stats[[3]] <- read.csv( "./overlapMeasuresLPBATemplateMALF.csv" )
-volumes[[3]] <- read.csv( "./labelVolumesLPBATemplateMALF.csv" )
-stats[[4]] <- read.csv( "./overlapMeasuresMGHTemplateMALF.csv" )
-volumes[[4]] <- read.csv( "./labelVolumesMGHTemplateMALF.csv" )
-stats[[5]] <- read.csv( "./overlapMeasuresMALTemplateMALF.csv" )
-volumes[[5]] <- read.csv( "./labelVolumesMALTemplateMALF.csv" )
+# stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplateMALF.csv" )
+# volumes[[1]] <- read.csv( "./labelVolumesCUMCTemplateMALF.csv" )
+# stats[[2]] <- read.csv( "./overlapMeasuresIBSRTemplateMALF.csv" )
+# volumes[[2]] <- read.csv( "./labelVolumesIBSRTemplateMALF.csv" )
+# stats[[3]] <- read.csv( "./overlapMeasuresLPBATemplateMALF.csv" )
+# volumes[[3]] <- read.csv( "./labelVolumesLPBATemplateMALF.csv" )
+# stats[[4]] <- read.csv( "./overlapMeasuresMGHTemplateMALF.csv" )
+# volumes[[4]] <- read.csv( "./labelVolumesMGHTemplateMALF.csv" )
+# stats[[5]] <- read.csv( "./overlapMeasuresMALTemplateMALF.csv" )
+# volumes[[5]] <- read.csv( "./labelVolumesMALTemplateMALF.csv" )
 
-# stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplate.csv" )
[…]

### 2013-10-28 paper.tex
WIP: minor ba edits

--- 2013-10-26 paper.tex
@@
 
+\selectlanguage{eng}
 \begin{document}
@@
 
-\title[DMFFD diffeomorphic image registration]{Explicit B-spline regularization in diffeomorphic image registration}
+\title[DMFFD Diffeomorphic Image Registration]{Explicit B-spline regularization in diffeomorphic image registration}
 \author[\firstAuthorLast ]{\Authors}
@@
 \begin{abstract}
-Important methodological developments in the evolution of image registration algorithms include those
-in which the correspondence relationship is characterized by diffeomorphisms.
-The  popularity of these approaches is due largely to their
+Diffeomorphic maps are central to image registration due largely to their
 topological properties and success in providing biologically plausible
@@
 time-varying and constant velocity fields, and symmetrical considerations.
-Prior information in the form of regularization is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
+Prior information (i.e.
+regularization) is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
 Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework, the {\it directly manipulated free-form deformation} approach \citep{tustison2009}, can be viewed as a smoothing alternative in which explicit regularization is achieved through fast B-spline approximation.
-This characterization can be used to provide B-spline ``flavored'' diffeomorphic image registration solutions with several advantages.
+In the context of the original Demons' framework, a free-form deformation methodological variant, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through
+fast B-spline appromixation.
+This characterization can be used to provide alternative B-spline ``flavored'' diffeomorphic image registration solutions with several advantages which we describe in this work.
 Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
-A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open labeled brain data and open source evaluation tools.
+A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open data and open source evaluation tools.
 \tiny
 \keyFont{ \section{Keywords:}
-Advanced Normalization Tools, diffeomorphisms, directly manipulated free-form deformation, Insight Toolkit, spatial normalization }
+open science, best practices, reproducibility, software }
 \end{abstract}
@@
 
-Establishment of anatomical and functional correspondence
-is a crucial step towards gaining insight into certain biological
-sciences.  Neuroscience research efforts, such as characterizing
-brain morphology, require accurate and robust methods for
-producing such mappings.  The
-extensive literature detailing methodology is evidence of the rich history of
-algorithmic development which continues contemporaneously.
-We highlight several key historical contributions which are particularly relevant to the work 
-subsequently presented.
-
-Free-form deformation (FFD) image registration, characterized by  regularization based on the B-spline basis functions, has several
-advantages including algorithmic simplicity,
-good performance, desirable properties, and several available
-implementations.  Current research was
-preceded by related work for
-geometric modeling \citep{sederberg1986} and originated with such important
-contributions as \cite{szeliski1997,thevenaz1998,rueckert1999}.
-Continued development within this early spline-based paradigm produced additional innovations such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation constraints \citep[e.g.][]{rohlfing2003}, and notable open source implementations \citep[e.g.][]{ibanez2005,klein2010,modat2010,shackleford2010}.
+Establishment of anatomical and functional correspondence is a crucial
+step towards gaining insight into biological processes.
+Neuroscience research efforts, such as characterizing brain
+morphology, require accurate and robust methods for producing such
+mappings.  The extensive literature detailing methodology is evidence
+of the rich history of algorithmic development which continues
+contemporaneously.  We highlight several key historical contributions
+which are particularly relevant to the work presented.
+
+Free-form deformation (FFD) image registration, characterized by
+regularization based on the B-spline basis functions, has several
+advantages including algorithmic simplicity, good performance,
+desirable properties, and several available implementations.  Current
+research was preceded by related work for geometric modeling
+\citep{sederberg1986} and originated with such important contributions
+as \cite{szeliski1997,thevenaz1998,rueckert1999}.  Continued
+development within this early spline-based paradigm produced
+additional innovations such as integrated similarity metrics
+\citep[e.g.][]{mattes2003}, additional transformation constraints
+\citep[e.g.][]{rohlfing2003}, and notable open source implementations
+\citep[e.g.][]{ibanez2005,klein2010,shackleford2010}.
 
 Parallel to this branch of algorithmic progress are the informally
-denoted ``dense transforms''
-perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
-Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are detailed in
-the works of \cite{bro-nielsen1996} and \cite{pennec1999} who observe that
-smoothing via Gaussian convolution, a defining characteristic of Demons,
-of the update or total displacement
-field is a greedy approximation for solving the partial differential equations governing
-the physics of an elastic or fluid deformation, respectively.  However, the use of such
-approximations entails that physical properties, such as topological
-regularity, are no longer guaranteed.
-
-It is interesting to note that within this context, traditional FFD algorithms
-can be viewed as a type of fluid-like Demons approach
-where, rather than projecting the update field to the space
-of regularized fields using Gaussian convolution, gradient fields
-are projected to a smooth space characterized by the B-spline
-basis functions.
-This analogy was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
-This preconditioning is used to mitigate the hemstitching effect induced by the ill-conditioned nature of the traditional gradient-based FFD formulation.  We denote
-this new FFD variant as directly manipulated free-form deformation (DMFFD) and, 
-as part of the ITKv4 refactoring efforts, has been implemented for use with the new
+denoted ``dense transforms'' perhaps best exemplified by Thirion's
+seminal contribution \citep{thirion1998}.  Relationships with earlier
+elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996}
+registration methods are detailed in the works of
+\cite{bro-nielsen1996} and \cite{pennec1999} who observe that
+smoothing via Gaussian convolution, a defining characteristic of
+Demons, of the update or total displacement field is a greedy
+approximation for solving the partial differential equations governing
+the physics of an elastic or fluid deformation, respectively.
+However, the use of such approximations entails that physical
+properties, such as topological regularity, are no longer guaranteed.
+
+It is interesting to note that within this context, traditional FFD
+algorithms can be viewed as a type of fluid-like Demons approach
+where, rather than projecting the update field to the space of
+regularized fields using Gaussian convolution, gradient fields are
+projected to a smooth space characterized by the B-spline basis
+functions.  This analogy was hinted at in our earlier work
+\citep{tustison2009} where we showed that fitting the update field to
+a B-spline object using a fast approximation routine
+\citep{tustison2006} is equivalent to a preconditioning of the
+standard gradient used in gradient descent-based FFD optimization.
+This preconditioning is used to mitigate the hemstitching effect
+induced by the ill-conditioned nature of the traditional
+gradient-based FFD formulation.  We denote this new FFD variant as
+directly manipulated free-form deformation (DMFFD) and, as part of the
+ITKv4 refactoring efforts, has been implemented for use with the new
 registration framework%
@@
 
-Continuing from the work of \cite{christensen1996} and subsequent exploration into the mathematical formalisms of diffeomorphisms \citep[e.g.][]{dupuis1998},
-the well-known Large Deformation Diffeomorphic Metric Mapping (LDDMM) algorithm
-was proposed in \cite{beg2005}.  In contrast to the mapping produced
-by \cite{christensen1996}, LDDMM yields the geodesic solution in the space of diffeomorphisms between two images. Since its introduction, LDDMM has
-inspired much innovation in the image registration community.  Applying
-the log-Euclidean framework of \cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using Exponential Lie algebra) uses a constant velocity field parameterization to provide a fast, diffeomorphic alternative \citep{ashburner2007}.
-Additionally, symmetrical considerations in the velocity field parameterization are discussed in \cite{avants2008} in the context of a cross correlation similarity metric.
-By explicitly symmetrizing the LDDMM formulation, this Symmetric Normalization (SyN) minimizes the bias of the resulting transformation when selecting the ``fixed'' and ``moving'' images.
-A greedy version of this algorithm has proven successful in neuroimaging \citep{klein2009} and pulmonary \citep{murphy2011} applications.
+Continuing from the work of \cite{christensen1996} and subsequent
+exploration into the mathematical formalisms of diffeomorphisms
+\citep[e.g.][]{dupuis1998}, the well-known Large Deformation
+Diffeomorphic Metric Mapping (LDDMM) algorithm was proposed in
+\cite{beg2005}.  In contrast to the mapping produced by
+\cite{christensen1996}, LDDMM yields the geodesic solution in the
+space of diffeomorphisms between two images. Since its introduction,
+LDDMM has inspired much innovation in the image registration
+literature.  Applying the log-Euclidean framework of
+\cite{arsigny2006}, DARTEL (Diffeomorphic Anatomic Registration using
+Exponential Lie algebra) uses a constant velocity field
+parameterization to provide a fast, diffeomorphic alternative
+\citep{ashburner2007}. Additionally, symmetrical considerations in the
+velocity field parameterization are discussed in \cite{avants2008} in
+the context of a cross correlation similarity metric.  By explicitly
+symmetrizing the LDDMM formulation, this Symmetric Normalization (SyN)
+minimizes the bias of the resulting transformation when selecting the
+``fixed'' and ``moving'' images.  A greedy version of this algorithm
+has proven successful in neuroimaging \citep{klein2009} and pulmonary
+\citep{murphy2011} applications as well as in multi-atlas label fusion \cite{wang2012}.
 
@@
 
-Although many extensions of LDDMM rely
-on some form of Gaussian convolution for
-regularization \citep[e.g.][]{risser2011}, there has been significant interest in constraining
-FFD approaches to the space of diffeomorphisms.
-An early attempt reported in \cite{rueckert2006} enforced
-diffeomorphic transforms
-by concatenating multiple FFD transforms, each of which is constrained
-to describe a one-to-one mapping.
-\cite{modat2011} incorporated the log-Euclidean
-framework for enforcing diffeomorphic transformations and ensuring invertibility.  Similarly, the work of \cite{de-craene2011} provided
-a full LDDMM-style algorithm based on B-splines called
-{\it temporal free-form deformation} in which the
-time-varying velocity field
-is modeled using a 4-D B-spline object (3-D + time).  Numerical Eulerian integration of the mapping propagated within the velocity field yields the transform between parameterized time points.
-
-As alluded to earlier, B-spline approximation can also be used
-for regularizing time-varying vector fields in an analogous fashion as
+Although many extensions of LDDMM rely on some form of Gaussian
+convolution for regularization \citep[e.g.][]{risser2011}, there has
+been significant interest in constraining FFD approaches to the space
+of diffeomorphisms.  An early attempt reported in \cite{rueckert2006}
+enforced diffeomorphic transforms by concatenating multiple FFD
+transforms, each of which is constrained to describe a one-to-one
+mapping.  \cite{modat2011} incorporated the log-Euclidean framework
+for enforcing diffeomorphic transformations and ensuring
+invertibility.  Similarly, the work of \cite{de-craene2011} provided a
+full LDDMM-style algorithm based on B-splines called {\it temporal
+free-form deformation} in which the time-varying velocity field is
+modeled using a 4-D B-spline object (3-D + time).  Numerical Eulerian
+integration of the mapping propagated within the velocity field yields
+the transform between parameterized time points.
+
+As alluded to earlier, B-spline approximation can also be used for
+regularizing time-varying vector fields in an analogous fashion as
 Gaussian convolution.  In this vein, and similar to
 \cite{de-craene2011}, we reported in \cite{tustison2012a,tustison2012}
-the use of an $n$-D + time B-spline object to
-represent the characteristic velocity fields.  However, we use the
-directly manipulated free-form deformation formulation to improve the solution convergence.  This also facilitates modeling
-temporal periodicity and the enforcement of stationary boundaries.
-Both this work and our earlier work
-\citep{tustison2009} demonstrate that the DMFFD framework is potentially
-applicable to the
-entire gamut of diffeomorphic registration algorithms and provides
-unique ``flavors'' of smoothing possibilities.  However, 
-the two approaches produce characteristically different
-solutions.  In addition to
-smoothing kernel differences, Gaussian
-convolution tends to ``flatten'' the signal in contrast to an
-approximation or fitting of the signal provided by the DMFFD
-approach.  Also, whereas
-Gaussian convolution operates entirely within discretized space,
-the B-spline approximation routine constructs a continuous object prior
-to any voxelwise reconstruction of the sampled fields.  
-Interestingly, a similar comparison was made with respect to 
-Gaussian derivative estimation  \citep{bourma2007}.  Although 
-typically estimated using  truncated, discrete Gaussian 
-convolution, an alternative based on 
-B-spline approximation demonstrated superior performance with
-similar computational cost.  
+the use of an $n$-D + time B-spline object to represent the
+characteristic velocity fields.  However, we use the directly
+manipulated free-form deformation formulation to improve the solution
+convergence.  This also facilitates modeling temporal periodicity and
+the enforcement of stationary boundaries.  Both this work and our
+earlier work \citep{tustison2009} demonstrate that the DMFFD framework
+is potentially applicable to the entire gamut of diffeomorphic
+registration algorithms and provides unique ``flavors'' of smoothing
+possibilities.  However, the two approaches produce characteristically
+different solutions.  In addition to smoothing kernel differences,
+Gaussian convolution tends to ``flatten'' the signal in contrast to an
+approximation or fitting of the signal provided by the DMFFD approach.
+Also, whereas Gaussian convolution operates entirely within
+discretized space, the B-spline approximation routine constructs a
+continuous object prior to any voxelwise reconstruction of the sampled
+fields.  Interestingly, a similar comparison was made with respect to
+Gaussian derivative estimation \citep{bourma2007}.  Although typically
+estimated using truncated, discrete Gaussian convolution, an
+alternative based on B-spline approximation demonstrated superior
+performance with similar computational cost.
 %Further elaboration on these differences
@@
 
-Variants of three popular diffeomorphic algorithms and their DMFFD analogs
-(LDDMM,
-DARTEL,
- and SyN)
-were implemented by the authors
-as part of the recent refactoring of the open source
-Insight Toolkit (ITK) although related work had been
-previously implemented within the popular Advanced
-Normalization Tools (ANTs).%
+Variants of three popular diffeomorphic algorithms and their DMFFD
+analogs (LDDMM, DARTEL, and SyN) were implemented by the authors as
+part of the recent refactoring of the open source Insight Toolkit
+(ITK) although related work had been previously implemented within the
+popular Advanced Normalization Tools (ANTs).%
 \footnote{
@@
 }
-Given the popularity and excellent performance of the greedy variant of SyN algorithm, our evaluation focus in this work is
-its B-spline analog which we denote as ``B-spline SyN'' or ``DMFFD SyN.''
-Evaluation and comparisons of the respective algorithmic instantiations
-are performed using the \verb#antsRegistration# program found
-in the ANTs repository (also originally developed by the authors).  This permits a direct algorithmic comparison
-as potential sources for implementation bias have been removed \citep{tustison2013}.  Additionally, in the spirit of open science,
-all text, figures, and
-scripts to reproduce the results contained in this work are publicly available online.%
+Given the popularity and excellent performance of the greedy variant
+of SyN, our evaluation focus in this work is its B-spline
+analog which we denote as ``B-spline SyN'' or ``DMFFD SyN.''
+Evaluations of the respective algorithmic
+instantiations are performed using the \verb#antsRegistration# program
+found in the ANTs repository (also originally developed by the
+authors).  This permits a direct algorithmic comparison as potential
+sources for implementation bias have been reduced
+\citep{tustison2013}.  Additionally, in the spirit of open science,
+all text, figures, and scripts to reproduce the results contained in
+this work are publicly available online.%
 \footnote{
@@
 lead to different solutions despite the fact that the theoretical
-geodesic solution describes the same parameterized path forwards and backwards.
+geodesic solution describes the same path forwards and backwards.
 This observation led to the symmetric
@@
 \begin{align}
-  \frac{d \phi_i(\mathbf{x},t)}{dt} = v_i( \phi_i(\mathbf{x},t), t ),\,\, \phi_i(\mathbf{x},0) = \mathbf{Id}, \,\, i \in \{1,2\}.
+  \frac{d \phi_i(\mathbf{x},t)}{dt} = v_i( \phi_i(\mathbf{x},t), t ),\,\, \phi_i(\mathbf{x},0) = \mathbf{Id}, \,\, i \in \{1,2\}
 \end{align}
-With extension to arbitrary similarity metric choice, the second
+with extension to arbitrary similarity metric choice, i.e. the second
 term is replaced with
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{SyN.jpg}
+  \includegraphics[width=0.95\textwidth]{SyN.pdf}
   \caption{Illustration of the greedy SyN formulation.  Given images
@@
 and storage of 4-D velocity fields at each iteration which is limiting for
-many common computational platforms.
+many common, contemporary processing infrastructures.
 
@@
 while simultaneously being capable of running with limited computational
-resources.  This is due to the restriction of the discrete, time-parameterized
-velocity field samples to their respective
+resources.  This is possible since the discrete time-parameterized
+velocity field samples are restricted to their respective
 endpoints, i.e. the $v_i(\mathbf{x},t)$ are sampled at $t \in \{0,0.5,1\}$
@@
 $\phi_1^{-1}(\mathbf{x})$, $\phi_2(\mathbf{x})$, and $\phi_2^{-1}(\mathbf{x})$
-(cf Figure \ref{fig:syn}).  Since the greedy SyN framework is the focus of the evaluation,
+(cf Figure \ref{fig:syn}).  Furthermore, the forward and inverse
+mappings are guaranteed to be consistent within the discrete domain i.e. $\|
+\phi_i^{-1}(\phi_i) - \mathbf{Id} \|^2 < \epsilon$.   
+Since the greedy SyN framework is the focus of the evaluation,
 the algorithmic steps are briefly sketched
@@
 \footnote{
-In terms of implementation, spline orders can be specified separately for each dimension but, for simplicity,
+In terms of implementation spline orders can be specified separately for each dimension but, for simplicity,
 we only specify a single spline order.
@@
 and $\frac{\partial \Pi_\sim}{\partial \mathbf{x}}$ is the spatial 
-similarity metric gradient at voxel $c$.
+similarity metric gradient indexed by $c$.
 
@@
 In this example, we first calculate an initial translation transform by aligning
-the centers of (intensity) mass (although alignment based on other features is possible). 
-The resulting transform is then used as input 
+the centers of (intensity) mass. The resulting transform is then used as input 
 for determining an optimal rigid transform.  Serial propagation of the resulting
@@
 criteria (including number of iterations per resolution level) and metric (or
-metrics).  Note that any pair of images can be specified per metric per stage.
+metrics).  Note that any pair of images can be specified per stage.
 
@@
 program which permits composition of any number of transform files with 
-different interpolation schemes.   For both programs interpolation is never
+various interpolation schemes.   For both programs interpolation is never
 performed more than once.
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{templates.jpg}
+  \includegraphics[width=0.95\textwidth]{templates.pdf}
   \caption{Canonical views for each of the five cohort-specific templates generated using
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{pseudo_geodesic.jpg}
+  \includegraphics[width=0.75\textwidth]{pseudo_geodesic.pdf}
   \caption{Illustration of generating a pseudo-geodesic for any two subjects
@@
 we generated the transforms from each subject to a cohort-specific
-shape/intensity template.  These templates were built using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
+shape/intensity template.  This was accomplished using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
 which is a multivariate implementation of the work described in
@@
   \centering
-  \includegraphics[width=0.95\textwidth]{labels.jpg}
+  \includegraphics[width=0.95\textwidth]{labels.pdf}
   \caption{Axial views of sample labelings for a member of each data set.  
@@
 using the multi-atlas label fusion algorithm (MALF) developed by \cite{wang2012}
-which is also distributed with ANTs.
+which is also distributed with the ANTs repository.
 For a given subject within a data set, every other subject was mapped to 
@@
   \centering
-  \begin{tabular}{c}
-  \includegraphics[width=0.95\textwidth]{evaluation.jpg}
+  \begin{tabular}{ccccc}
+  {} &
+  \multicolumn{2}{c}{\bf Original labeling} & 
+  \multicolumn{2}{c}{\bf MALF labeling} \\
+  \rotatebox{90}{\qquad\,\, CUMC12} &
+  \includegraphics[height=38mm]{versusCUMC.png} &
+  \includegraphics[height=38mm]{synComparisonCUMC.pdf} &
+  \includegraphics[height=38mm]{versusCUMCMALF.png} &
+  \includegraphics[height=38mm]{synComparisonCUMCMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad IBSR18} &
+  \includegraphics[height=38mm]{versusIBSR.png} &
+  \includegraphics[height=38mm]{synComparisonIBSR.pdf} &
+  \includegraphics[height=38mm]{versusIBSRMALF.png} &
+  \includegraphics[height=38mm]{synComparisonIBSRMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad LPBA40} &
+  \includegraphics[height=38mm]{versusLPBA.png} &
+  \includegraphics[height=38mm]{synComparisonLPBA.pdf} &
+  \includegraphics[height=38mm]{versusLPBAMALF.png} &
+  \includegraphics[height=38mm]{synComparisonLPBAMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad MGH10} &
+  \includegraphics[height=38mm]{versusMGH.png} &
+  \includegraphics[height=38mm]{synComparisonMGH.pdf} &
+  \includegraphics[height=38mm]{versusMGHMALF.png} &
+  \includegraphics[height=38mm]{synComparisonMGHMALF.pdf} \\
+  \rotatebox{90}{\qquad\quad MAL35} &
+  \includegraphics[height=38mm]{versusMAL.png} &
+  \includegraphics[height=38mm]{synComparisonMAL.pdf} &
+  \includegraphics[height=38mm]{versusMALMALF.png} &
+  \includegraphics[height=38mm]{synComparisonMALMALF.pdf} \\
   \end{tabular}
@@
   Dice results by volume and a combined label box plotting.  The
-  left and right halves show the respective results for 
-  original and MALF-derived labelings.
+  left and right halves show the results for MALF-derived and 
+  original labelings, respectively.
   }
@@
 \end{figure}
-
-%\begin{figure}[htb]
-%  \centering
-%  \begin{tabular}{ccccc}
-%  {} &
-%  \multicolumn{2}{c}{\bf Original labeling} & 
-%  \multicolumn{2}{c}{\bf MALF labeling} \\
-%  \rotatebox{90}{\qquad\,\, CUMC12} &
-%  \includegraphics[height=38mm]{versusCUMC.png} &
-%  \includegraphics[height=38mm]{synComparisonCUMC.pdf} &
-%  \includegraphics[height=38mm]{versusCUMCMALF.png} &
-%  \includegraphics[height=38mm]{synComparisonCUMCMALF.pdf} \\
-%  \rotatebox{90}{\qquad\quad IBSR18} &
-%  \includegraphics[height=38mm]{versusIBSR.png} &
-%  \includegraphics[height=38mm]{synComparisonIBSR.pdf} &
-%  \includegraphics[height=38mm]{versusIBSRMALF.png} &
-%  \includegraphics[height=38mm]{synComparisonIBSRMALF.pdf} \\
-%  \rotatebox{90}{\qquad\quad LPBA40} &
-%  \includegraphics[height=38mm]{versusLPBA.png} &
-%  \includegraphics[height=38mm]{synComparisonLPBA.pdf} &
-%  \includegraphics[height=38mm]{versusLPBAMALF.png} &
-%  \includegraphics[height=38mm]{synComparisonLPBAMALF.pdf} \\
-%  \rotatebox{90}{\qquad\quad MGH10} &
-%  \includegraphics[height=38mm]{versusMGH.png} &
-%  \includegraphics[height=38mm]{synComparisonMGH.pdf} &
-%  \includegraphics[height=38mm]{versusMGHMALF.png} &
-%  \includegraphics[height=38mm]{synComparisonMGHMALF.pdf} \\
-%  \rotatebox{90}{\qquad\quad MAL35} &
-%  \includegraphics[height=38mm]{versusMAL.png} &
-%  \includegraphics[height=38mm]{synComparisonMAL.pdf} &
-%  \includegraphics[height=38mm]{versusMALMALF.png} &
-%  \includegraphics[height=38mm]{synComparisonMALMALF.pdf} \\
-%  \end{tabular}
-%  \caption{Dice results for both algorithms for 
-%  each subject warped to every other subject using the pseudo-
-%  geodesic transform.  Each 
-%  row corresponds to one of the five data sets used for evaluation.
-%  For each data set we include a plotting of all individual label 
-%  Dice results by volume and a combined label box plotting.  The
-%  left and right halves show the results for MALF-derived and 
-%  original labelings, respectively.
-%  }
-%  \label{fig:dice}
-%\end{figure}
 
@@
 demonstrate robust performance across data sets.  For both algorithms,
-the gradient step was 0.1 for each of the four multi-resolution levels
-with shrink factors of $\{6,4,2,1\}$ and Gaussian smoothing for each
+the gradient step was 0.1 during each of the four multi-resolution levels
+with shrink factors of $\{6,4,2,1\}$ with Gaussian smoothing for each
 of those levels being $\mathcal{N}\left(0,\{9,4,1,0\}\right)$ in terms
@@
 }
-The perl scripts used to create the jobs for the cluster are
+All perl scripts used to create the jobs for the cluster are
 included in the github account associated with this evaluation.  
@@
 for the update field of 26 mm at the base level which is reduced by
-a factor of two for each succeeding multiresolution level.  This 
+a factor of two for each subsequent multiresolution level.  This 
 yields a final knot spacing of 3.25 mm.%
@@
  }
-For comparison, after selecting these smoothing parameters we discovered 
-in the supplementary material of \cite{klein2009} the similarity to the 
+As specified in the supplementary material associated with \cite{klein2009}, 
+this is similar to the 
 gradient smoothing parameter for the IRTK FFD algorithm
@@
 registration pair within a data set versus the volume of the label
-in the fixed image.  Values above and below the dashed line indicate better 
-regional performance for B-spline SyN and SyN, respectively.
+in the fixed image.
 
@@
   \centering
-    \includegraphics[width=0.95\textwidth]{jacobianRange.jpg}
+    \includegraphics[width=0.95\textwidth]{jacobianRange.pdf}
   \caption{Violin plots of the range of log Jacobian values 
@@
 respective range of values in log Jacobians.  We plotted the 
-(95th\% - 5th\%) for each algorithm across all data sets
+(95th\% - the 5th\%) for each algorithm across all data sets
  in Figure \ref{fig:jacobian}.  It is apparent 
-that B-spline SyN results exhibit a much greater range of deformation.
+that B-spline SyN produces a much greater range of deformation
+values during the course of optimization.
 
@@
 Avoid extensive citations and discussion of published literature.
-
-B-spline SyN produced slightly greater Dice values than the original SyN.  Although
-actual differences are relatively small, they are statistically significant. By 
-implementing both algorithms in the the same code base, we are not only able to 
-eliminate non-regularization components of the registration but we are also able
-to eliminate implementation differences.  Thus, performance disparity can be isolated
-to smoothing choice.  However, even with this restricted focus there are various
-reasons for the evaluation outcome.  These include the approximation-versus-convolution 
-distinction mentioned earlier for the two regularization approaches (which could
-also explain the reason why the range of log Jacobian values tend to be
-significantly higher for B-spline SyN).  Also, the fact that the regularization 
-for B-spline SyN is theoretically continuous whereas the truncated Gaussian 
-convolution is only a discrete approximation could be a potential factor.  Exploration
-of other possibilities are planned for future work.
-
-One of the advantages that has not been explored in this manuscript is 
-the use of B-spline SyN for small deformation estimation problems such 
-as in pulmonary or cardiac applications.  Such problems typically require
-greater regularization which implies larger discrete kernels for 
-Gaussian convolution.
-A related issue is applications involving severely anisotropic data where
-the continuous nature of the DMFFD approach might help over Gaussian convolution.
-Additionally, we have not yet explored the use of sparsely sampled data 
-and the advantages of the proposed framework.
-Ongoing work will continue to explore these issues.
 
@@
 B-spline regularization is easily adapted into the diffeomorphic registration 
-framework and performs well compared to analogous algorithms which we demonstrated 
+framework and performs comparably to analogous algorithms which we demonstrated 
 in this work for the case of the widely-used SyN.
+
+B-spline SyN produced slightly greater Dice values than the original SyN.
+Interestingly it was shown that the range of log Jacobian values tend to be
+significantly higher for the former than the latter.  Although possible 
+explanations include the approximation-versus-convolution distinction mentioned
+earlier, further exploration is planned.
+
+One of the advantages that has not been explored in this manuscript is 
+the use of B-spline SyN for small deformation estimation problems such 
+as in pulmonary or cardiac applications.  Such problems typically require
+greater regularization which implies larger discrete Gaussian kernels.
+A related issue is applications involving severely anisotropic data where
+the continuous nature of the DMFFD approach might help over Gaussian convolution.
+Additionally, we have not yet explored the use of sparsely sampled data 
+and the advantages of the proposed framework.
+Ongoing work will continue to explore these issues.
 

### 2013-10-28 paper.tex
ENH:  Preparing for merge.

@@
 
-\selectlanguage{eng}
+%\selectlanguage{eng}
 \begin{document}
@@
 solutions to deformation and morphological
-estimation problems. Popular variants of diffeomorphic algorithms include those characterized by
+estimation problems. Popular diffeomorphic algorithms include those characterized by
 time-varying and constant velocity fields, and symmetrical considerations.
-Prior information (i.e.
-regularization) is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
-Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework, a free-form deformation methodological variant, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}, can be viewed as an alternative smoothing approach in which explicit regularization is achieved through
-fast B-spline appromixation.
-This characterization can be used to provide alternative B-spline ``flavored'' diffeomorphic image registration solutions with several advantages which we describe in this work.
-Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
-A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open data and open source evaluation tools.
+Prior information in the form of regularization is used to enforce transform plausibility 
+taking the form of physics-based constraints or through some approximation thereof, e.g. 
+Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).  In 
+the context of the original Demons' framework, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}
+can be viewed as a smoothing alternative  in which explicit regularization is achieved through 
+fast B-spline approximation.
+This characterization can be used to provide B-spline ``flavored'' diffeomorphic image 
+registration solutions with several advantages.  Implementation 
+is open source and available through the Insight Toolkit and our Advanced Normalization Tools
+(ANTs) repository.  A thorough comparative evaluation with the well-known SyN algorithm 
+\citep{avants2008}, implemented within the same framework, and its B-spline analog is 
+performed using open labeled brain data and open source evaluation tools.
 \tiny
 \keyFont{ \section{Keywords:}
-open science, best practices, reproducibility, software }
+Advanced Normalization Tools, diffeomorphisms, directly manipulated free-form deformation, Insight Toolkit, spatial normalization }
 \end{abstract}
@@
 \citep[e.g.][]{rohlfing2003}, and notable open source implementations
-\citep[e.g.][]{ibanez2005,klein2010,shackleford2010}.
+\citep[e.g.][]{ibanez2005,klein2010,modat2010,shackleford2010}.
 
@@
 induced by the ill-conditioned nature of the traditional
-gradient-based FFD formulation.  We denote this new FFD variant as
+gradient-based FFD formulation.  We denoted this new FFD variant as
 directly manipulated free-form deformation (DMFFD) and, as part of the
@@
 and storage of 4-D velocity fields at each iteration which is limiting for
-many common, contemporary processing infrastructures.
+many common processing infrastructures.
 
 Therefore, in addition to the full-scale
-symmetric SyN offering described in \cite{avants2008}, the authors therein
+SyN offering described in \cite{avants2008}, the authors therein
 provided a ``greedy'' alternative which has demonstrated superior performance
-in various applications \citep{avants2011,klein2009,murphy2011}%
+in differents applications \citep{avants2011,klein2009,murphy2011}%
 \footnote{
@@
 DMFFD-based diffeomorphic algorithms, we
-assign image boundary metric gradients a value of zero with a corresponding
+assign to the image boundary metric gradients the null vector with a corresponding
 large confidence value.
@@
 
+B-spline SyN produced slightly greater Dice values than the original SyN. 
+Although actual differences are relatively small, they are statistically 
+significant. By implementing both algorithms in the the same code base, 
+we are not only able to eliminate non-regularization components of the 
+registration but we are also able to eliminate implementation differences. 
+Thus, performance disparity can be isolated to smoothing choice. However, 
+even with this restricted focus there are various reasons for the evaluation 
+outcome. These include the approximation-versus-convolution distinction 
+mentioned earlier for the two regularization approaches (which could also 
+explain the reason why the range of log Jacobian values tend to be significantly 
+higher for B-spline SyN). Also, the fact that the regularization for 
+B-spline SyN is theoretically continuous whereas the truncated Gaussian 
+convolution is only a discrete approximation could be a potential factor. 
+Exploration of other possibilities are planned for future work.
+
+One of the advantages that has not been explored in this manuscript is 
+the use of B-spline SyN for small deformation estimation problems such 
+as in pulmonary or cardiac applications.  Such problems typically require
+greater regularization which implies larger discrete Gaussian kernels.
+A related issue is applications involving severely anisotropic data where
+the continuous nature of the DMFFD approach might help over Gaussian convolution.
+Additionally, we have not yet explored the use of sparsely sampled data 
+and the advantages of the proposed framework.
+Ongoing work will continue to explore these issues.
+
 A significant amount of research has been devoted to image registration
@@
 
-B-spline SyN produced slightly greater Dice values than the original SyN.
-Interestingly it was shown that the range of log Jacobian values tend to be
-significantly higher for the former than the latter.  Although possible 
-explanations include the approximation-versus-convolution distinction mentioned
-earlier, further exploration is planned.
-
-One of the advantages that has not been explored in this manuscript is 
-the use of B-spline SyN for small deformation estimation problems such 
-as in pulmonary or cardiac applications.  Such problems typically require
-greater regularization which implies larger discrete Gaussian kernels.
-A related issue is applications involving severely anisotropic data where
-the continuous nature of the DMFFD approach might help over Gaussian convolution.
-Additionally, we have not yet explored the use of sparsely sampled data 
-and the advantages of the proposed framework.
-Ongoing work will continue to explore these issues.
-
 %Perhaps most important, in the spirt of open science, all source code

### 2013-10-28 paper.tex

--- 2013-10-28 paper.tex
@@
 
+%\selectlanguage{eng}
 \begin{document}
@@
 \begin{abstract}
-Important methodological developments in the evolution of image registration algorithms include those
-in which the correspondence relationship is characterized by diffeomorphisms.
-The  popularity of these approaches is due largely to their
+Diffeomorphic maps are central to image registration due largely to their
 topological properties and success in providing biologically plausible
 solutions to deformation and morphological
-estimation problems. Popular variants of diffeomorphic algorithms include those characterized by
+estimation problems. Popular diffeomorphic algorithms include those characterized by
 time-varying and constant velocity fields, and symmetrical considerations.
-Prior information in the form of regularization is used to enforce transform plausibility taking the form of physics-based constraints or through some approximation thereof, e.g.
-Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).
-In the context of the original Demons' framework, the {\it directly manipulated free-form deformation} approach \citep{tustison2009}, can be viewed as a smoothing alternative in which explicit regularization is achieved through fast B-spline approximation.
-This characterization can be used to provide B-spline ``flavored'' diffeomorphic image registration solutions with several advantages.
-Implementation is open source and available through the Insight Toolkit and our Advanced Normalization Tools (ANTs) repository.
-A thorough comparative evaluation with the well-known SyN algorithm \citep{avants2008}, implemented within the same framework, and its B-spline analog is performed using open labeled brain data and open source evaluation tools.
+Prior information in the form of regularization is used to enforce transform plausibility 
+taking the form of physics-based constraints or through some approximation thereof, e.g. 
+Gaussian smoothing of the vector fields (a la Thirion's Demons \citep{thirion1998}).  In 
+the context of the original Demons' framework, the so-called {\it directly manipulated free-form deformation} \citep{tustison2009}
+can be viewed as a smoothing alternative  in which explicit regularization is achieved through 
+fast B-spline approximation.
+This characterization can be used to provide B-spline ``flavored'' diffeomorphic image 
+registration solutions with several advantages.  Implementation 
+is open source and available through the Insight Toolkit and our Advanced Normalization Tools
+(ANTs) repository.  A thorough comparative evaluation with the well-known SyN algorithm 
+\citep{avants2008}, implemented within the same framework, and its B-spline analog is 
+performed using open labeled brain data and open source evaluation tools.
 \tiny
@@
 
-Establishment of anatomical and functional correspondence
-is a crucial step towards gaining insight into certain biological
-sciences.  Neuroscience research efforts, such as characterizing
-brain morphology, require accurate and robust methods for
-producing such mappings.  The
-extensive literature detailing methodology is evidence of the rich history of
-algorithmic development which continues contemporaneously.
-We highlight several key historical contributions which are particularly relevant to the work 
-subsequently presented.
-
-Free-form deformation (FFD) image registration, characterized by  regularization based on the B-spline basis functions, has several
-advantages including algorithmic simplicity,
-good performance, desirable properties, and several available
-implementations.  Current research was
-preceded by related work for
-geometric modeling \citep{sederberg1986} and originated with such important
-contributions as \cite{szeliski1997,thevenaz1998,rueckert1999}.
-Continued development within this early spline-based paradigm produced additional innovations such as integrated similarity metrics \citep[e.g.][]{mattes2003}, additional transformation constraints \citep[e.g.][]{rohlfing2003}, and notable open source implementations \citep[e.g.][]{ibanez2005,klein2010,modat2010,shackleford2010}.
+Establishment of anatomical and functional correspondence is a crucial
+step towards gaining insight into biological processes.
+Neuroscience research efforts, such as characterizing brain
+morphology, require accurate and robust methods for producing such
+mappings.  The extensive literature detailing methodology is evidence
+of the rich history of algorithmic development which continues
+contemporaneously.  We highlight several key historical contributions
+which are particularly relevant to the work presented.
+
+Free-form deformation (FFD) image registration, characterized by
+regularization based on the B-spline basis functions, has several
+advantages including algorithmic simplicity, good performance,
+desirable properties, and several available implementations.  Current
+research was preceded by related work for geometric modeling
+\citep{sederberg1986} and originated with such important contributions
+as \cite{szeliski1997,thevenaz1998,rueckert1999}.  Continued
+development within this early spline-based paradigm produced
+additional innovations such as integrated similarity metrics
+\citep[e.g.][]{mattes2003}, additional transformation constraints
+\citep[e.g.][]{rohlfing2003}, and notable open source implementations
+\citep[e.g.][]{ibanez2005,klein2010,modat2010,shackleford2010}.
 
 Parallel to this branch of algorithmic progress are the informally
-denoted ``dense transforms''
-perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
-Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are detailed in
-the works of \cite{bro-nielsen1996} and \cite{pennec1999} who observe that
-smoothing via Gaussian convolution, a defining characteristic of Demons,
-of the update or total displacement
-field is a greedy approximation for solving the partial differential equations governing
-the physics of an elastic or fluid deformation, respectively.  However, the use of such
-approximations entails that physical properties, such as topological
-regularity, are no longer guaranteed.
-
-It is interesting to note that within this context, traditional FFD algorithms
-can be viewed as a type of fluid-like Demons approach
-where, rather than projecting the update field to the space
-of regularized fields using Gaussian convolution, gradient fields
-are projected to a smooth space characterized by the B-spline
-basis functions.
-This analogy was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.
-This preconditioning is used to mitigate the hemstitching effect induced by the ill-conditioned nature of the traditional gradient-based FFD formulation.  We denote
-this new FFD variant as directly manipulated free-form deformation (DMFFD) and, 
-as part of the ITKv4 refactoring efforts, has been implemented for use with the new
+denoted ``dense transforms'' perhaps best exemplified by Thirion's
+seminal contribution \citep{thirion1998}.  Relationships with earlier
+elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996}
+registration methods are detailed in the works of
+\cite{bro-nielsen1996} and \cite{pennec1999} who observe that
+smoothing via Gaussian convolution, a defining characteristic of
+Demons, of the update or total displacement field is a greedy
+approximation for solving the partial differential equations governing
+the physics of an elastic or fluid deformation, respectively.
+However, the use of such approximations entails that physical
+properties, such as topological regularity, are no longer guaranteed.
+
+It is interesting to note that within this context, traditional FFD
+algorithms can be viewed as a type of fluid-like Demons approach
+where, rather than projecting the update field to the space of
+regularized fields using Gaussian convolution, gradient fields are
+projected to a smooth space characterized by the B-spline basis
+functions.  This analogy was hinted at in our earlier work
+\citep{tustison2009} where we showed that fitting the update field to
+a B-spline object using a fast approximation routine
+\citep{tustison2006} is equivalent to a preconditioning of the
+standard gradient used in gradient descent-based FFD optimization.
+This preconditioning is used to mitigate the hemstitching effect
+induced by the ill-conditioned nature of the traditional
+gradient-based FFD formulation.  We denoted this new FFD variant as
+directly manipulated free-form deformation (DMFFD) and, as part of the
+ITKv4 refactoring efforts, has been implemented for use with the new
 registration framework%
@@
 Therefore, in addition to the full-scale
-symmetric SyN offering described in \cite{avants2008}, the authors therein
+SyN offering described in \cite{avants2008}, the authors therein
 provided a ``greedy'' alternative which has demonstrated superior performance
-in various applications \citep{avants2011,klein2009,murphy2011}%
+in differents applications \citep{avants2011,klein2009,murphy2011}%
 \footnote{
@@
 
-B-spline SyN produced slightly greater Dice values than the original SyN.  Although
-actual differences are relatively small, they are statistically significant. By 
-implementing both algorithms in the the same code base, we are not only able to 
-eliminate non-regularization components of the registration but we are also able
-to eliminate implementation differences.  Thus, performance disparity can be isolated
-to smoothing choice.  However, even with this restricted focus there are various
-reasons for the evaluation outcome.  These include the approximation-versus-convolution 
-distinction mentioned earlier for the two regularization approaches (which could
-also explain the reason why the range of log Jacobian values tend to be
-significantly higher for B-spline SyN).  Also, the fact that the regularization 
-for B-spline SyN is theoretically continuous whereas the truncated Gaussian 
-convolution is only a discrete approximation could be a potential factor.  Exploration
-of other possibilities are planned for future work.
+B-spline SyN produced slightly greater Dice values than the original SyN. 
+Although actual differences are relatively small, they are statistically 
+significant. By implementing both algorithms in the the same code base, 
+we are not only able to eliminate non-regularization components of the 
+registration but we are also able to eliminate implementation differences. 
+Thus, performance disparity can be isolated to smoothing choice. However, 
+even with this restricted focus there are various reasons for the evaluation 
+outcome. These include the approximation-versus-convolution distinction 
+mentioned earlier for the two regularization approaches (which could also 
+explain the reason why the range of log Jacobian values tend to be significantly 
+higher for B-spline SyN). Also, the fact that the regularization for 
+B-spline SyN is theoretically continuous whereas the truncated Gaussian 
+convolution is only a discrete approximation could be a potential factor. 
+Exploration of other possibilities are planned for future work.
 
@@
 
+A significant amount of research has been devoted to image registration
+algorithmic development.  Given their many salient characteristics particularly with
+respect to large deformation estimation constrained by topological 
+continuity, diffeomorphic registration
+approaches have been a particular focus in the neuroimaging community.
+However, many groups continue to find success with non-diffeomorphic
+FFD methods
+\citep[e.g.][]{rueckert1999,klein2010}.  Using our DMFFD framework, 
+B-spline regularization is easily adapted into the diffeomorphic registration 
+framework and performs comparably to analogous algorithms which we demonstrated 
+in this work for the case of the widely-used SyN.
+
 %Perhaps most important, in the spirt of open science, all source code

### 2013-10-28 paper.tex
ENH:  Submission ready.

@@
 \begin{abstract}
-Diffeomorphic maps are central to image registration due largely to their
+Diffeomorphic mappings are central to image registration due largely to their
 topological properties and success in providing biologically plausible
 solutions to deformation and morphological
-estimation problems. Popular diffeomorphic algorithms include those characterized by
+estimation problems. Popular diffeomorphic image registration algorithms include those characterized by
 time-varying and constant velocity fields, and symmetrical considerations.
@@
 
-Variants of three popular diffeomorphic algorithms and their DMFFD
+Three popular diffeomorphic algorithms and their DMFFD
 analogs (LDDMM, DARTEL, and SyN) were implemented by the authors as
@@
 criteria (including number of iterations per resolution level) and metric (or
-metrics).  Note that any pair of images can be specified per metric per stage.
+metrics).  Any pair of images can be specified per metric per stage.
 
@@
 included in the github account associated with this evaluation.  
-It should be noted that for the
+For the
 data sets used in this study, times for B-spline SyN were approximately 
@@
 convolution is only a discrete approximation could be a potential factor. 
-Exploration of other possibilities are planned for future work.
-
-One of the advantages that has not been explored in this manuscript is 
+%Exploration of other possibilities are planned for future work.
+
+Relative to Klein's study \citep{klein2009}, it should be noted that 
+the SyN implementation used in that study is found in the {\tt ANTS} 
+program which is a precursor of the {\tt antsRegistration} program
+described earlier.  Although the theoretical aspects 
+are the same, there are substantial implementation differences between 
+the two programs.  In addition, several parameters varied between the two
+studies which translated into a more aggressive metric and 
+gradient step in addition to fewer levels and iterations.%
+\footnote{
+Specifically, as provided in the supplementary material of \cite{klein2009},
+the command line parameters were:
+{\tt ANTS 3 -m PR[\$\{fixed\},\$\{moving\},1,2] -o \$\{output\} -r Gauss[2,0] 
+-t SyN[0.5] -i 30x99x11 --use-Histogram-Matching}.  
+}
+In this study we took a more conservative approach based on our 
+continued development and experience resulting in parameters which have proven
+useful in our cortical thickness pipeline (encapsulated in 
+the ANTs script {\tt antsCorticalThickness.sh}) and our
+experience with the MICCAI 2013 SATA challenge data.  We also 
+ran our own internal experiments with Gaussian SyN (both {\tt ANTS} and 
+{\tt antsRegistration}) using the same conservative parameters 
+on the MAL35 data for which the latter demonstrated slightly improved
+performance over the former.
+
+One of the advantages that has not been explored in this work is 
 the use of B-spline SyN for small deformation estimation problems such 
@@
 
-A significant amount of research has been devoted to image registration
-algorithmic development.  Given their many salient characteristics particularly with
-respect to large deformation estimation constrained by topological 
-continuity, diffeomorphic registration
-approaches have been a particular focus in the neuroimaging community.
-However, many groups continue to find success with non-diffeomorphic
-FFD methods
-\citep[e.g.][]{rueckert1999,klein2010}.  Using our DMFFD framework, 
-B-spline regularization is easily adapted into the diffeomorphic registration 
-framework and performs comparably to analogous algorithms which we demonstrated 
-in this work for the case of the widely-used SyN.
-
 %Perhaps most important, in the spirt of open science, all source code

### 2013-11-01 paper.tex
BUG:  Missed parentheses.
BUG:  Another missed parentheses.

@@
 has proven successful in neuroimaging \citep{klein2009} and pulmonary
-\citep{murphy2011} applications as well as in multi-atlas label fusion \cite{wang2012}.
+\citep{murphy2011} applications as well as in multi-atlas label fusion \citep{wang2012}.
 
@@
   The second row consists of the original labelings with the third row being
-  refined versions of those labelings using the MALF algorithm \cite{wang2012}.
+  refined versions of those labelings using the MALF algorithm \citep{wang2012}.
   These refinements provide more consistency between labelings and improved

### 2013-11-24 paper.tex
ENH:  Resubmission.

@@
 
-\def\journal{Neuroscience}%%% write here for which journal %%%
+\def\journal{Neuroinformatics}%%% write here for which journal %%%
 \def\DOI{}
@@
 regularization based on the B-spline basis functions, has several
-advantages including algorithmic simplicity, good performance,
-desirable properties, and several available implementations.  Current
+advantages including algorithmic simplicity, good performance, and
+{\color{blue} guaranteed parametric continuity}.  Current
 research was preceded by related work for geometric modeling
@@
 induced by the ill-conditioned nature of the traditional
-gradient-based FFD formulation.  We denoted this new FFD variant as
-directly manipulated free-form deformation (DMFFD) and, as part of the
+gradient-based FFD formulation.%
+\footnote{
+{\color{blue}{This phenomenon is explained in greater detail in \cite{tustison2009}
+but, briefly, the distribution of the uniform B-spline basis functions
+spanning the image domain produces long ravines in the image registration
+energy topography.  These topographical features cause an
+inefficient back-and-forth traversal in classic steepest descent optimization, also known as ``hemstitching.''}}
+}
+We denoted this new FFD variant as
+{\it directly manipulated free-form deformation} (DMFFD) and, as part of the
 ITKv4 refactoring efforts, has been implemented for use with the new
@@
 is potentially applicable to the entire gamut of diffeomorphic
-registration algorithms and provides unique ``flavors'' of smoothing
-possibilities.  However, the two approaches produce characteristically
+registration algorithms and provides {\color{blue} alternative smoothing 
+possibilities with different continuity properties (e.g. $C^2$
+versus $C^3$)}.
+
+However, the two approaches produce characteristically
 different solutions.  In addition to smoothing kernel differences,
@@
 provided a ``greedy'' alternative which has demonstrated superior performance
-in differents applications \citep{avants2011,klein2009,murphy2011}%
+in different applications \citep{avants2011,klein2009,murphy2011}%
 \footnote{
@@
 \Comment $i \in \{1,2\}$
+\ForAll {image resolution levels} 
 \State $n \leftarrow 1$
@@
 \EndWhile
+  \State upsample current $\phi_i$ and $\phi_i^{-1}$ to next resolution level 
+  \Comment $i \in \{1,2\}$
+\EndFor
 \State \Return $\phi \leftarrow \phi_1 \circ \phi_2^{-1}$, $\phi^{-1} \leftarrow \phi_2 \circ \phi_1^{-1}$
@@
 default to Gaussian smoothing due to its simplicity both in implementation
-and complexity terms.  As mentioned earlier, 
-a viable and practical alternative is the DMFFD 
+and complexity terms.  A viable and practical alternative is the DMFFD 
 approach based on B-splines for explicit regularization of vector fields.
-As presented in \cite{tustison2009}, given the similarity metric
+
+Given the similarity metric
 $\Pi_{\sim}$, the $d$-dimensional update field (i.e. preconditioned
-gradient field), $\delta v_{i_1,\ldots,i_d}$, is calculated from
+gradient field), $\delta v_{i_1,\ldots,i_d}$, is given by
 \begin{align}
@@
 and $\frac{\partial \Pi_\sim}{\partial \mathbf{x}}$ is the spatial 
-similarity metric gradient at voxel $c$.
+similarity metric gradient at voxel $c$.%
+\footnote{
+{\color{blue}{For comparison, FFD image registration uses the
+calculated gradient which, for a single control point is evaluated using (in 
+discretized form assuming no other explicit regularization):
+\begin{align}
+\label{eq:ffd}
+  \delta v_{i_1,\ldots,i_d} = \sum_{c=1}^{N_{\Omega}} 
+  \left( \frac{\partial \Pi_\sim}{\partial \mathbf{x}} \right)_c 
+  \prod_{j=1}^d B_{i_j}(x_j^c).
+\end{align}
+Intuitively, this can be understood as a distribution of the spatial similarity metric
+gradient to the set of control points according to the basis function weighting which, due to
+the local property of B-splines, is only non-zero in a localized region surrounding each 
+sample point.  The problematic issue with standard FFD image registration is that 
+the parametric domain of 
+the peripheral B-spline basis functions extends beyond the image boundaries resulting in
+a relatively lower weighting contribution effect.  As a corrective, Equation (\ref{eq:dmffd})
+normalizes the control point gradient contribution by the actual B-spline weighting overlap
+with the image domain.
+
+Note that this modification is first discussed and derived in \cite{hsu1992,lee1997} in the context of fitting scattered data for a computer graphics audience which was an extension and improvement over Sederberg's original FFD manipulation computer graphics technique \citep{sederberg1986}.
+Whereas the latter (i.e.
+FFD) technique performs geometric deformation of the graphics object embedded in a B-spline object via manipulation of the control points, the former (i.e.
+DMFFD) technique permits direct manipulation of the actual
+object with the control point values being updated indirectly.  The corresponding fast algorithm for scattered data approximation using DMFFD was proposed in \cite{lee1997}.
+
+Based on the needs of one of our colleagues, the first author (N.T.) implemented and
+generalized the work of \cite{lee1997}  which was reported in \cite{tustison2006} and
+was N.T.'s very first ITK contribution.  
+A couple years later the second author (B.A.) requested an FFD implementation for 
+ANTs which N.T. naively assumed to be equivalent to applying \cite{tustison2006}
+to smoothing the metric gradient (a la Demons).  The subsequent realization 
+that this assumption was incorrect and figuring out why the algorithm still worked 
+despite it not being a correct FFD implementation resulted in \cite{tustison2009}.
+}}
+}
 
@@
 utility classes) were developed as part of the ITKv4 registration
-framework refactoring.  Further details can be found in the documentation
+framework refactoring.  {\color{blue}{Much of the original ITK image registration infrastructure 
+was left intact including the so-called ``sparse'' transforms such as various
+rigid (versor, Euclidean) and other linear transforms.  The transform classes
+contributed by our group were, including those listed above, were meant to 
+augment what already existed.  All the transforms listed above are derived from
+the {\tt itk::DisplacementFieldTransform} class which permits specification of 
+a transform described by a sampled displacement field.  The derived classes are
+then modified according to the different transform constraints. Other types
+of classes are used to coordinate the image registration process.  The {\tt itk::ImageRegistrationMethodv4} is the base interface for performing all 
+image registration steps.  Smoothing and resampling for multi-resolution
+image registration is performed in this class as is the calling of the
+selected optimizer.  Output consists of a single optimized transform.
+Multiple instantiations of this class in series, in conjunction with the 
+{\tt itk::CompositeTransform} class, are used to optimize a composition of transforms.
+Some of the diffeomorphic approaches do not easily fit into this
+generalized optimization framework necessitating specialized method classes such as
+those listed above.  The adaptors%
+\footnote{
+{\color{blue}{We gratefully acknowledge Marius Staring's contribution in providing the broad
+design of the adaptors.}}
+}
+are used to modify the parameters between resolution levels during the course
+of transform optimization within the methods classes.  For example, the 
+resolution of the displacement field transforms follows that of the image 
+resolution and the updating is handled by the corresponding transform adaptors.
+}}
+Further details can be found in the documentation
 provided within the classes themselves.%
@@
 In this example, we first calculate an initial translation transform by aligning
-the centers of (intensity) mass (although alignment based on other features is possible). 
+the centers of (intensity) mass (although alignment based on other features is possible).%
+\footnote{
+{\color{blue}{This step corresponds to the {\tt --initial-moving-transform} option.  Generally, the
+user can specify an ITK transform for initialization or can perform an initial
+translation based on either the geometric center of the images, the center of the
+image intensities, or the origin of the images.  See the help menu `{\tt antsRegistration --help}' for more details.
+}}
+} 
 The resulting transform is then used as input 
@@
 criteria (including number of iterations per resolution level) and metric (or
-metrics).  Any pair of images can be specified per metric per stage.
+metrics).  Any pair of images can be specified per metric per stage.%
+\footnote{
+{
+\color{blue}{
+To help the reader who wishes to explore {\tt antsRegistration} with specific use of the
+parameters used for this study and specified above, we have created a script with a simplified interface and placed it in the {\tt Scripts/} subdirectory of the ANTs repository.
+This script, called {\tt antsRegistrationSyN.sh} takes a fixed and moving image and performs a comprehensive (i.e.
+rigid $\rightarrow$ affine $\rightarrow$ deformable) registration using SyN (`{\tt -t s}') or B-spline SyN (`{\tt -t b}').
+}}
+}
 
@@
 \begin{itemize}
-  \item MGH10 
   \item CUMC12
@@
   \item LPBA40
+  \item MGH10 
 \end{itemize}
@@
 }
+
 The number of subjects per cohort is provided in the denotation.
-Further details of these labeled brain data (e.g. labeling protocol,
+{\color{blue}{Table 1 summarizes core information about the data sets used.}}
+Further details of these first four labeled brain data (e.g. labeling protocol,
 data sources) are given in \citep{klein2009}.
@@
 under academic subscription.
+
+\begin{table*}[htb]
+  \caption{
+    {\color{blue}{ Brief overview of the data sets used for the SyN vs. B-spline SyN comparison.}}
+    }
+  \centering
+  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l c c  c l}
+  \toprule
+  \multicolumn{1}{c}{Cohort} & Resolution & Number of labels & \multicolumn{1}{c}{Preprocessing} \\
+  \midrule
+  CUMC12 & $[0.86,0.86,1.5]$ & 131 & \parbox[t]{6.5cm}{rotated to ``cardinal'' pose}\\ 
+  IBSR18 & $[1,1.5,1]$ & 96 & \parbox[t]{6.5cm}{aligned to Talaraich, bias corr.}\\ 
+  LPBA40 & $[0.86,1.5,0.86]$ & 56  & \parbox[t]{6.5cm}{registered to MNI305, bias corr.}\\
+  MGH10 & $[1,1,1]$ & 106 & \parbox[t]{6.5cm}{affine-registered to MNI152, bias corr.}\\ 
+  MAL35 & $[1,1,1]$ & 138 & \parbox[t]{6.5cm}{average of 3-4 acquisitions, bias corr.}\\
+  \bottomrule
+  \end{tabular*}
+  \label{table:demo}
+\end{table*}
 
@@
 we generated the transforms from each subject to a cohort-specific
-shape/intensity template.  These templates were built using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
+shape/intensity template.  {\color{blue}{Not only does this reduce the computational
+time required for finding the pairwise transforms between subjects but 
+prior work has demonstrated  improvement in registration with this approach\
+over direct pairwise registration \citep{klein2010a}.  Since the two algorithms
+have been implemented within the same framework, all registration parameters are 
+identical (i.e. linear registration stage parameters, winsorizing values, etc.) 
+except for the parameters governing the smoothing of the gradient field. 
+}}
+
+The cohort templates were built using the ANTs script {\tt antsMultivariateTemplateConstruction.sh}
 which is a multivariate implementation of the work described in
@@
 that subject using the pseudo-geodesic transform.  The set of transformed labelings 
-were then used to determine a consensus
-labeling for that subject.
+were then used to determine a consensus labeling for that subject.
 This was to minimize the obvious {\it observer dimensionality artifacts} where manual
@@
 We include both sets of results.  This provides two sets of labels per subject
-for evaluation.
+for evaluation.%
+\footnote{
+{\color{blue}{The MALF labels are available at http://figshare.com/account/projects/196.}}
+}
 
@@
   \begin{tabular}{c}
-  \includegraphics[width=0.95\textwidth]{evaluation.jpg}
+  \includegraphics[width=0.95\textwidth]{evaluation2.jpg}
   \end{tabular}
@@
   left and right halves show the respective results for 
-  original and MALF-derived labelings.
+  original and MALF-derived labelings.  The black dashed regression line
+  ($y \sim x^3$) illustrates how performance difference varies with label volume for
+  each cohort.
   }
@@
 of voxels.  The number of iterations per level were $\{100,100,70,20\}$
-with a convergence threshold of $10^{-9}$ and window size of 15 iterations.
-
-All processing was performed using the computational cluster at the 
-University of Virginia.%
+with a convergence threshold of $10^{-9}$ and window size of 15 iterations.%
+\footnote{
+{\color{blue}{
+As part of the ITKv4 refactoring, we developed several convergence
+monitoring classes (cf the base class {\tt itk::ConvergenceMonitoringFunction}).
+Instead of testing for convergence between two successive time points, we use a 
+windowed monitoring function which keeps track of a sequence, or window, of 
+energy values and determines the convergence from the slope of a line fitted to 
+the series.
+}}
+}
+
+{\color{blue}{
+All processing was performed using the linux cluster at the 
+University of Virginia%
 \footnote{
@@
 }
+using the PBS Pro queuing system for managing resources.
+}}
 The perl scripts used to create the jobs for the cluster are
 included in the github account associated with this evaluation.  
-For the
-data sets used in this study, times for B-spline SyN were approximately 
+For the data sets used in this study, times for B-spline SyN were approximately 
 15-40\% greater than Gaussian-based SyN using single-threading and a dense
 metric gradient sampling.  
-
-The differences between the two algorithms consist of the Gaussian and 
-B-spline parameters governing the update field smoothing, $S_v$.  In our 
+{\color{blue}{
+Timing data for specific data sets are given in Table \ref{table:timing}
+Timing includes both rigid and affine transform  optimizations.  
+\begin{table*}[htb]
+  \caption{
+    {\color{blue}{Timing (in hours) per registration for both SyN algorithms across data sets.}}
+    }
+  \centering
+  \begin{tabular*}{0.5\textwidth}{@{\extracolsep{\fill}} l c c}
+  \toprule
+  \multicolumn{1}{c}{Cohort} & SyN & B-spline SyN \\
+  \midrule
+  CUMC12 & $8.6  \pm  2.3$ & $9.6 \pm 2.1$ \\ 
+  IBSR18 & $8.3  \pm  2.7$ & $10.7 \pm 2.8$ \\
+  LPBA40 & $6.6 \pm 0.7$ & $9.5 \pm 1.4$ \\
+  MGH10 & $4.7  \pm  1.5$ & $5.8 \pm 1.2$ \\
+  MAL35 & $10.6 \pm 1.9$ & $14.2 \pm 2.0$ \\
+  \bottomrule
+  \end{tabular*}
+  \label{table:timing}
+\end{table*}
+}}
+
+The {\color{blue}{only}} difference between the two {\color{blue}{registration 
+settings}} consist of the Gaussian and 
+B-spline parameters governing the update field smoothing, $S_v$.  
+In our 
 experience, smoothing of the total field did not improve the results, at
@@
 that B-spline SyN results exhibit a much greater range of deformation.
+{\color{blue}{
+Qualitative differences are shown in Figure \ref{fig:qualitative}.  In order to ensure
+randomness to minimize presentation bias in illustrating qualitative results 
+(and given the relative poor performance in humans as random number
+generators \citep{wagenaar1972}), we used R to generate uniform random numbers for
+both subject and axial slice selection.  We then used the log Jacobian
+images to locate regions of maximal difference between the SyN and B-spline SyN
+results.  From Figure \ref{fig:qualitative} it is quite apparent that the results
+are very similar which is to be expected considering the almost identical algorithmic
+make-up between the two approaches (e.g. similarity metric, implementation, linear 
+transforms).  However, there are subtle differences particularly in the cortex which 
+help explain both the relative difference in Jacobian and Dice distribution.
+}}
+
+\begin{figure}[htb]
+  \centering
+    \includegraphics[width=0.95\textwidth]{qualitative.jpg}
+  \caption{Randomly selected axial slices showing qualitative differences between
+  SyN and B-spline SyN.  Note that, at least for these slices, the cortical 
+  differences are the most prominent. 
+  }
+  \label{fig:qualitative}
+\end{figure}
 
@@
 Although actual differences are relatively small, they are statistically 
-significant. By implementing both algorithms in the the same code base, 
+significant. By implementing both algorithms in the same code base, 
 we are not only able to eliminate non-regularization components of the 
@@
 convolution is only a discrete approximation could be a potential factor.
+
+{\color{blue}{
+Additional observations of interest concern the differences in results
+between the MALF and original labelings for all data sets.  Not only was
+there an overall increase in performance for both algorithms with the 
+MALF labels for all cohorts, but there was
+also an increase in performance disparity relative to the variance in the
+resulting Dice values.  A possible explanation for this, and one that seems
+quite plausible, is that the MALF labelings are derived from registering
+a cohort to the target subject and then performing a consensus labeling \cite{wang2012}.  
+Both these steps are heavily reliant on image intensity information (in
+fact, both use a form of correlation as the similarity measure for
+optimization).  Since the MALF ``correction'' tends to group labeled regions 
+according to the same metric used for establishing anatomical correspondences,
+alignment of these labeled regions seems much more likely which would result 
+in higher Dice metrics.  A related effect for labeled data in general is 
+being currently investigated by the authors.
+From the label volume versus Dice difference plots, there is no immediately 
+discernible pattern of performance variation with region size.  However,
+a possible confound is that region definitions vary between cohorts.  Although
+we did not look at region versus performance difference variation within cohorts,
+such inquiry is certainly possible as we have made the resulting csv files
+available with the github repository associated with this work.
+}}
+
 %Exploration of other possibilities are planned for future work.
@@
 
+{\color{blue}{
+Despite the thorough evaluation with multiple data sets, we readily 
+acknowledge the limitations of this study including a very focused application,
+i.e. healthy brains of a single modality, and absence of a thorough 
+exploration of parameter selection and sensitivity.  Although such 
+work might be beneficial (e.g. by aiding other researchers in parameter
+selection), characterizing parameter permutations 
+of potential interest would expand the current work far beyond its 
+intended scope.  However, we note that in addition to this work,
+SyN has also been previously evaluated in \cite{klein2009} and 
+\cite{avants2011} and, based on additional experience and application,
+the parameter set was modified each time but still yielded excellent
+performance providing evidence for flexibility in parameter selection.
+Outside of a range of parameters based on
+sound engineering principles, experience, and intimate knowledge of both the 
+corresponding algorithms and software, determination of optimal generic
+parameters even for a specific application is difficult.  In fact, the
+``No Free Lunch Theorem'' \cite{wolpert1997} 
+emphasizes the importance of prior knowledge in tuning optimization 
+algorithms for a particular application.
+}}
+
 One of the advantages that has not been explored in this work is 
@@
 Gaussian convolution.
-A related issue is applications involving severely anisotropic data where
-the continuous nature of the DMFFD approach might help over Gaussian convolution.
-Additionally, we have not yet explored the use of sparsely sampled data 
-and the advantages of the proposed framework.
+A related issue concerns applications involving severely anisotropic data where
+the continuous nature of the DMFFD approach might help over Gaussian 
+convolution.  {\color{blue}{Also, we emphasize that underlying the DMFFD approaches 
+is a fitting routine for sparse and scattered data which offers added flexibility
+over smoothing using discrete convolution where
+the latter implies regularly placed  data on a rectilinear grid for conventional
+implementations.  This advantage could translate into faster running times if
+only select points are used to drive the registration or make possible more
+complex registration scenarios
+involving data arranged continuously within a finite domain \cite[e.g.][]{tustison2011}.
+Finally, the possibility of varying data confidence values, as introduced in \cite{tustison2006} with DMFFD-based routines
+would permit incorporation of spatial preferential weighting (i.e. additional 
+prior knowledge) during optimization.
+}} 
 Ongoing work will continue to explore these issues.
@@
 
-\bibliographystyle{frontiersinSCNS}
-%\bibliographystyle{plain}
+%\bibliographystyle{frontiersinSCNS}
+\bibliographystyle{plain}
 \bibliography{references}

### 2013-11-24 Results/overlapMeasures.R

@@ -4,32 +4,29 @@ library( reshape )
 stats <- list()
 volumes <- list()
 
-
-
 cohortNames <- c( 'CUMC', 'IBSR', 'LPBA', 'MGH', 'MAL' )
 
-# stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplateMALF.csv" )
-# volumes[[1]] <- read.csv( "./labelVolumesCUMCTemplateMALF.csv" )
-# stats[[2]] <- read.csv( "./overlapMeasuresIBSRTemplateMALF.csv" )
-# volumes[[2]] <- read.csv( "./labelVolumesIBSRTemplateMALF.csv" )
-# stats[[3]] <- read.csv( "./overlapMeasuresLPBATemplateMALF.csv" )
-# volumes[[3]] <- read.csv( "./labelVolumesLPBATemplateMALF.csv" )
-# stats[[4]] <- read.csv( "./overlapMeasuresMGHTemplateMALF.csv" )
-# volumes[[4]] <- read.csv( "./labelVolumesMGHTemplateMALF.csv" )
-# stats[[5]] <- read.csv( "./overlapMeasuresMALTemplateMALF.csv" )
-# volumes[[5]] <- read.csv( "./labelVolumesMALTemplateMALF.csv" )
+stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplateMALF.csv" )
+volumes[[1]] <- read.csv( "./labelVolumesCUMCTemplateMALF.csv" )
+stats[[2]] <- read.csv( "./overlapMeasuresIBSRTemplateMALF.csv" )
+volumes[[2]] <- read.csv( "./labelVolumesIBSRTemplateMALF.csv" )
+stats[[3]] <- read.csv( "./overlapMeasuresLPBATemplateMALF.csv" )
+volumes[[3]] <- read.csv( "./labelVolumesLPBATemplateMALF.csv" )
+stats[[4]] <- read.csv( "./overlapMeasuresMGHTemplateMALF.csv" )
+volumes[[4]] <- read.csv( "./labelVolumesMGHTemplateMALF.csv" )
+stats[[5]] <- read.csv( "./overlapMeasuresMALTemplateMALF.csv" )
+volumes[[5]] <- read.csv( "./labelVolumesMALTemplateMALF.csv" )
 
-stats[[1]] <- read.csv( "./overlapMeasuresCUMCTemplate.csv" )
[…]

### 2013-11-24 Results/plotConvergence.R

@@ -75,6 +75,9 @@ for( f in 1:length( logFiles ) )
   timeSyN <- convergenceTable$ELAPSED_TIME[convergenceTable$REGULARIZATION == 'syn'] / 3600
   timeBSplineSyN <- convergenceTable$ELAPSED_TIME[convergenceTable$REGULARIZATION == 'bsyn'] / 3600
 
+  cat( "  SyN time:  ", mean( timeSyN ), ' +/- ', sd( timeSyN ), "\n" );
+  cat( "  BSyN time:  ", mean( timeBSplineSyN ), ' +/- ', sd( timeBSplineSyN ), "\n" );
+
   plotLabels <- c( rep.int( "SyN", length( timeSyN ) ), rep.int( "BSplineSyN", length( timeBSplineSyN ) ) )
 
   timeData <- data.frame( Type = plotLabels, Time = c( timeSyN, timeBSplineSyN ) )

### 2013-12-08 paper.tex
ENH:  Final edits?

@@
 advantages including algorithmic simplicity, good performance, and
-{\color{blue} guaranteed parametric continuity}.  Current
+guaranteed parametric continuity.  Current
 research was preceded by related work for geometric modeling
@@
 \footnote{
-{\color{blue}{This phenomenon is explained in greater detail in \cite{tustison2009}
+This phenomenon is explained in greater detail in \cite{tustison2009}
 but, briefly, the distribution of the uniform B-spline basis functions
@@
 energy topography.  These topographical features cause an
-inefficient back-and-forth traversal in classic steepest descent optimization, also known as ``hemstitching.''}}
+inefficient back-and-forth traversal in classic steepest descent optimization, also known as ``hemstitching.''
 }
@@
 is potentially applicable to the entire gamut of diffeomorphic
-registration algorithms and provides {\color{blue} alternative smoothing 
+registration algorithms and provides alternative smoothing 
 possibilities with different continuity properties (e.g. $C^2$
-versus $C^3$)}.
+versus $C^3$).
 
@@
 \footnote{
-{\color{blue}{For comparison, FFD image registration uses the
+For comparison, FFD image registration uses the
 calculated gradient which, for a single control point is evaluated using (in 
@@
 that this assumption was incorrect and figuring out why the algorithm still worked 
-despite it not being a correct FFD implementation resulted in \cite{tustison2009}.
-}}
+despite it not being a ``correct'' FFD implementation resulted in \cite{tustison2009}.
 }
@@
 utility classes) were developed as part of the ITKv4 registration
-framework refactoring.  {\color{blue}{Much of the original ITK image registration infrastructure 
+framework refactoring.  Much of the original ITK image registration infrastructure 
 was left intact including the so-called ``sparse'' transforms such as various
@@
 \footnote{
-{\color{blue}{We gratefully acknowledge Marius Staring's contribution in providing the broad
-design of the adaptors.}}
+We gratefully acknowledge Marius Staring's contribution in providing the broad
+design of the adaptors.
 }
@@
 resolution and the updating is handled by the corresponding transform adaptors.
-}}
 Further details can be found in the documentation
@@
                  --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
-                 --convergence 1000x500x250 \
-                 --smoothing-sigmas 2.0x1.0x0.0mm \
-                 --shrink-factors 4x3x2 \
+                 --convergence 1000x500x250x100 \
+                 --smoothing-sigmas 3x2x1x0 \
+                 --shrink-factors 8x4x2x1 \
                  --transform Affine[0.1] \
                  --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
-                 --convergence 1000x500x250 \
-                 --smoothing-sigmas 2.0x1.0x0.0mm \
-                 --shrink-factors 4x3x2 \
+                 --convergence 1000x500x250x100 \
+                 --smoothing-sigmas 3x2x1x0 \
+                 --shrink-factors 8x4x2x1 \
                  --transform BSplineSyN[0.1,26,0,3] \
@@
                  --convergence 100x70x50x20 \
-                 --smoothing-sigmas 2x1.0x0.5x0.0mm \
+                 --smoothing-sigmas 3x2x1x0 \
                  --shrink-factors 6x4x2x1
@@
 \footnote{
-{\color{blue}{This step corresponds to the {\tt --initial-moving-transform} option.  Generally, the
+This step corresponds to the {\tt --initial-moving-transform} option.  Generally, the
 user can specify an ITK transform for initialization or can perform an initial
@@
 image intensities, or the origin of the images.  See the help menu `{\tt antsRegistration --help}' for more details.
-}}
 } 
@@
 \footnote{
-{
-\color{blue}{
 To help the reader who wishes to explore {\tt antsRegistration} with specific use of the
@@
 rigid $\rightarrow$ affine $\rightarrow$ deformable) registration using SyN (`{\tt -t s}') or B-spline SyN (`{\tt -t b}').
-}}
 }
@@
 The number of subjects per cohort is provided in the denotation.
-{\color{blue}{Table 1 summarizes core information about the data sets used.}}
+Table 1 summarizes core information about the data sets used.
 Further details of these first four labeled brain data (e.g. labeling protocol,
@@
   \caption{
-    {\color{blue}{ Brief overview of the data sets used for the SyN vs. B-spline SyN comparison.}}
+    Brief overview of the data sets used for the SyN vs. B-spline SyN comparison.
     }
@@
 we generated the transforms from each subject to a cohort-specific
-shape/intensity template.  {\color{blue}{Not only does this reduce the computational
+shape/intensity template.  Not only does this reduce the computational
 time required for finding the pairwise transforms between subjects but 
@@
 except for the parameters governing the smoothing of the gradient field.
-}}
 
@@
 \footnote{
-{\color{blue}{The MALF labels are available at http://figshare.com/account/projects/196.}}
+The MALF labels are available at http://figshare.com/account/projects/196.
 }
@@
 \footnote{
-{\color{blue}{
 As part of the ITKv4 refactoring, we developed several convergence
@@
 the series.
-}}
-}
-
-{\color{blue}{
+}
+
 All processing was performed using the linux cluster at the 
@@
 using the PBS Pro queuing system for managing resources.
-}}
 The perl scripts used to create the jobs for the cluster are
@@
 metric gradient sampling.  
-{\color{blue}{
 Timing data for specific data sets are given in Table \ref{table:timing}
@@
   \caption{
-    {\color{blue}{Timing (in hours) per registration for both SyN algorithms across data sets.}}
+    Timing (in hours) per registration for both SyN algorithms across data sets.
     }
@@
 \end{table*}
-}}
-
-The {\color{blue}{only}} difference between the two {\color{blue}{registration 
-settings}} consist of the Gaussian and 
+
+The only difference between the two registration 
+settings consists of the Gaussian and 
 B-spline parameters governing the update field smoothing, $S_v$.  
@@
 that B-spline SyN results exhibit a much greater range of deformation.
-{\color{blue}{
 Qualitative differences are shown in Figure \ref{fig:qualitative}.  In order to ensure
@@
 help explain both the relative difference in Jacobian and Dice distribution.
-}}
 
@@
   \caption{Randomly selected axial slices showing qualitative differences between
-  SyN and B-spline SyN.  Note that, at least for these slices, the cortical 
-  differences are the most prominent. 
+  SyN and B-spline SyN.  Crosshairs marks regions of maximal Jacobian difference. 
   }
@@
 
-{\color{blue}{
 Additional observations of interest concern the differences in results
@@
 available with the github repository associated with this work.
-}}
 
@@
 experience with the MICCAI 2013 SATA challenge data.  We also 
-ran our own internal experiments with Gaussian SyN (both {\tt ANTS} and 
+ran our own internal experiments with Gaussian-based SyN (both {\tt ANTS} and 
 {\tt antsRegistration}) using the same conservative parameters 
@@
 
-{\color{blue}{
 Despite the thorough evaluation with multiple data sets, we readily 
@@
 of potential interest would expand the current work far beyond its 
-intended scope.  However, we note that in addition to this work,
+intended scope.  However, in addition to this work,
 SyN has also been previously evaluated in \cite{klein2009} and 
@@
 algorithms for a particular application.
-}}
 
@@
 the continuous nature of the DMFFD approach might help over Gaussian 
-convolution.  {\color{blue}{Also, we emphasize that underlying the DMFFD approaches 
+convolution.  Also, we emphasize that underlying the DMFFD approaches 
 is a fitting routine for sparse and scattered data which offers added flexibility
@@
 prior knowledge) during optimization.
-}} 
 Ongoing work will continue to explore these issues.

### 2013-12-09 paper.tex
BUG:  word change.
ENH:  Final edits, indeed.

@@
 symmetrizing the LDDMM formulation, this Symmetric Normalization (SyN)
-minimizes the bias of the resulting transformation when selecting the
+approach minimizes the bias of the resulting transformation when selecting the
 ``fixed'' and ``moving'' images.  A greedy version of this algorithm
@@
 
-However, the two approaches produce characteristically
+The two regularization approaches (i.e. Gaussian convolution vs. DMFFD), 
+however,  produce characteristically
 different solutions.  In addition to smoothing kernel differences,
@@
   \caption{Randomly selected axial slices showing qualitative differences between
-  SyN and B-spline SyN.  Crosshairs marks regions of maximal Jacobian difference. 
+  SyN and B-spline SyN.  Crosshairs indicate regions of maximal Jacobian difference. 
   }
